---
title: "Untitled"
author: "Ross Cunning"
date: "2026-04-28"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
measurements <- read_csv("data/processed/turbidity.csv")

ggplot(measurements, aes(x = date, y = turbidity_ntu)) +
  geom_point(aes(color = sample_group), alpha = 0.6) +
  facet_wrap(~depth_position)

ggplot(measurements, aes(x = sample_depth_ft)) +
  facet_wrap(~depth_position) +
  geom_histogram()




diffs <- measurements %>%
  filter(!is.na(sample_number)) %>%
  select(date, sample_group, sample_number, depth_position, turbidity_ntu) %>%
  pivot_wider(names_from = sample_group, values_from = turbidity_ntu) %>%
  mutate(diff = `compliance` - `background`)



  

ggplot(diffs, aes(x = date, y = diff)) +
  geom_point(aes(shape = factor(sample_number))) +
  facet_wrap(~depth_position) +
  geom_smooth(method = "loess", span = 0.05)


ggplot(filter(diffs, depth_position != "Bottom"), aes(x = date, y = diff)) +
  geom_point(aes(shape = factor(sample_number))) +
  #geom_smooth(method = "loess", span = 0.019) +
  geom_line()


```

## Turbidity threshold screening

The project has two turbidity thresholds that are useful to screen directly
from the parsed monitoring data:

1. Construction shutdown: compliance turbidity exceeds background by more than
   29 NTU.
2. Benthic monitoring trigger: the densest portion of the plume extends within
   a 10 m radius of the same location for three consecutive days and exceeds
   10 NTU at each reading.

The second threshold has a spatial component. The parsed PDF table currently
contains sampling group, sample number, depth, date, and time, but not station
coordinates or plume-location geometry. The code below therefore screens the
`>10 NTU` portion directly and leaves a clearly marked placeholder for adding
the 10 m same-location logic when location data are available.

```{r turbidity-threshold-setup}
shutdown_threshold_ntu <- 29
benthic_screen_threshold_ntu <- 10
benthic_trigger_days <- 3
benthic_trigger_radius_m <- 10

turbidity_screen <- diffs %>%
  filter(
    #age_status == "sampled",
    #observed_depth_row,
    #!is.na(turbidity_ntu),
    #!is.na(difference_from_background)
  ) %>%
  mutate(
    shutdown_exceedance = diff > shutdown_threshold_ntu,
    benthic_threshold_exceedance = diff > benthic_screen_threshold_ntu,
    threshold_class = case_when(
      shutdown_exceedance ~ ">29 NTU above background",
      benthic_threshold_exceedance ~ ">10 to 29 NTU above background",
      TRUE ~ "<=10 NTU above background"
    ),
    threshold_class = factor(
      threshold_class,
      levels = c(
        "<=10 NTU above background",
        ">10 to 29 NTU above background",
        ">29 NTU above background"
      )
    )
  )
```

### Exceedance summaries

These tables list individual exceedance records and summarize counts by date,
sample group, and depth. They are intended as the first audit trail for dates
when shutdown or benthic-screen thresholds would have been crossed.

```{r turbidity-exceedance-tables}
shutdown_exceedances <- turbidity_screen %>%
  filter(shutdown_exceedance)

benthic_threshold_exceedances <- turbidity_screen %>%
  filter(benthic_threshold_exceedance)

turbidity_daily_summary <- turbidity_screen %>%
  summarize(
    n_readings = n(),
    n_benthic_threshold_exceedances = sum(benthic_threshold_exceedance, na.rm = TRUE),
    n_shutdown_exceedances = sum(shutdown_exceedance, na.rm = TRUE),
    max_difference_from_background = max(diff, na.rm = TRUE),
    .by = c(date, depth_position)
  ) %>%
  arrange(date, depth_position)

shutdown_exceedances
benthic_threshold_exceedances
turbidity_daily_summary %>% filter(!is.infinite(max_difference_from_background))
```

### Threshold plots

These plots show raw turbidity differences through time relative to the two
monitoring thresholds. Points are not averaged before plotting, so high
individual readings remain visible.

```{r turbidity-threshold-plots}
ggplot(
  turbidity_screen,
  aes(x = date, y = diff)
) +
  geom_hline(
    yintercept = benthic_screen_threshold_ntu,
    linetype = "dashed",
    color = "goldenrod3"
  ) +
  geom_hline(
    yintercept = shutdown_threshold_ntu,
    linetype = "dashed",
    color = "firebrick"
  ) +
  geom_point(aes(color = threshold_class), alpha = 0.75) +
  facet_wrap(~depth_position) +
  scale_color_manual(
    values = c(
      "<=10 NTU above background" = "grey45",
      ">10 to 29 NTU above background" = "goldenrod3",
      ">29 NTU above background" = "firebrick"
    )
  ) +
  labs(
    x = NULL,
    y = "Turbidity above background (NTU)",
    color = NULL
  )
```

### Consecutive-day benthic trigger scaffold

This chunk identifies three-or-more consecutive dates with at least one
`>10 NTU above background` reading. This is not a complete benthic-trigger
determination until a plume location or monitoring-station field is added and
used to enforce the 10 m same-location requirement.

```{r benthic-trigger-scaffold}
benthic_daily_screen <- turbidity_screen %>%
  filter(benthic_threshold_exceedance) %>%
  mutate(
    # TODO: Replace this placeholder with a station/plume-location identifier,
    # or derive location groups from coordinates using a 10 m radius.
    plume_location_id = "location_not_available"
  ) %>%
  summarize(
    n_exceeding_readings = n(),
    max_difference_from_background = max(diff, na.rm = TRUE),
    .by = c(plume_location_id, date)
  ) %>%
  arrange(plume_location_id, date) %>%
  group_by(plume_location_id) %>%
  mutate(
    consecutive_run = cumsum(date != lag(date, default = first(date)) + days(1))
  ) %>%
  group_by(plume_location_id, consecutive_run) %>%
  mutate(
    consecutive_exceedance_days = n_distinct(date),
    preliminary_benthic_trigger = consecutive_exceedance_days >= benthic_trigger_days
  ) %>%
  ungroup()

preliminary_benthic_trigger_periods <- benthic_daily_screen %>%
  filter(preliminary_benthic_trigger) %>%
  summarize(
    start_date = min(date),
    end_date = max(date),
    consecutive_exceedance_days = n_distinct(date),
    max_difference_from_background = max(max_difference_from_background, na.rm = TRUE),
    total_exceeding_readings = sum(n_exceeding_readings),
    .by = c(plume_location_id, consecutive_run)
  ) %>%
  arrange(start_date)

benthic_daily_screen
preliminary_benthic_trigger_periods
```
