---
title: "Sedimentation Monitoring"
author: "Ross Cunning"
date: "2026-04-29"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)

project_dir <- if (basename(getwd()) == "code") dirname(getwd()) else getwd()
knitr::opts_knit$set(root.dir = project_dir)
```

## Import

```{r paths}
sediment_dir <- "data/sedimentation_monitoring"

gps_file <- file.path(sediment_dir, "PE_SB_Station_Transect_GPS.xlsx")

sediment_files <- list.files(
  sediment_dir,
  pattern = "xlsx$",
  recursive = TRUE,
  full.names = TRUE
) %>%
  discard(~ basename(.x) == basename(gps_file))
```

```{r helpers}
excel_date <- function(x) {
  if (inherits(x, "Date")) return(x)
  if (inherits(x, "POSIXt")) return(as.Date(x))

  suppressWarnings(as.Date(as.numeric(x), origin = "1899-12-30"))
}

survey_number <- function(x) {
  if_else(
    x == "Baseline",
    0L,
    suppressWarnings(as.integer(str_extract(x, "\\d+")))
  )
}

normalize_transect_id <- function(x) {
  x %>%
    str_squish() %>%
    str_replace_all("\\s+", "_")
}

model_sedimentation_class <- function(transect) {
  case_when(
    transect %in% c("ST22", "ST31") ~ "1.0 cm predicted (dashed red)",
    transect %in% c("ST28", "ST5") ~ "1.5 mm predicted (solid red)",
    TRUE ~ "0 cm predicted (outside)"
  ) %>%
    factor(
      levels = c(
        "1.0 cm predicted (dashed red)",
        "1.5 mm predicted (solid red)",
        "0 cm predicted (outside)"
      ),
      ordered = TRUE
    )
}

model_sedimentation_predictions <- tibble(
  model_sedimentation_class = factor(
    c(
      "1.0 cm predicted (dashed red)",
      "1.5 mm predicted (solid red)",
      "0 cm predicted (outside)"
    ),
    levels = c(
      "1.0 cm predicted (dashed red)",
      "1.5 mm predicted (solid red)",
      "0 cm predicted (outside)"
    ),
    ordered = TRUE
  ),
  predicted_total_sedimentation_cm = c(1, 0.15, 0)
)

threshold_class <- function(x) {
  case_when(
    is.na(x) ~ NA_character_,
    x > 10 ~ ">10 cm",
    x > 5 ~ "5-10 cm",
    x > 1 ~ "1-5 cm",
    x > 0.5 ~ ">5 mm to 1 cm",
    x > 0.15 ~ ">1.5 to 5 mm",
    x >= 0 ~ "0 to 1.5 mm",
    TRUE ~ "below baseline"
  ) %>%
    factor(
      levels = c(
        "below baseline",
        "0 to 1.5 mm",
        ">1.5 to 5 mm",
        ">5 mm to 1 cm",
        "1-5 cm",
        "5-10 cm",
        ">10 cm"
      )
    )
}

read_sediment_raw <- function(path) {
  read_excel(path, sheet = "SedimentDepth_RawData", guess_max = 20000) %>%
    clean_names() %>%
    mutate(
      source_file = path,
      source_file_name = basename(path),
      source_modified = file.info(path)$mtime,
      monitoring_phase = case_when(
        str_detect(path, "1_Pre-Construction") ~ "pre_construction",
        str_detect(path, "2_During-Construction") ~ "during_construction",
        TRUE ~ "unknown"
      ),
      monitoring_type = case_when(
        str_detect(path, "Weekly Monitoring") ~ "weekly",
        str_detect(path, "Exceedance Triggered Monitoring") ~ "exceedance_triggered",
        TRUE ~ "compiled"
      ),
      date = excel_date(date),
      survey = str_squish(survey),
      survey_n = survey_number(survey),
      transect_type = str_squish(transect_type),
      transect = normalize_transect_id(transect),
      model_sedimentation_class = model_sedimentation_class(transect),
      meter = as.integer(meter),
      expected_permanent_length_m = case_when(
        transect_type == "Permanent" & transect %in% c("ST1", "ST38") ~ 150L,
        transect_type == "Permanent" ~ 50L,
        TRUE ~ NA_integer_
      ),
      across(starts_with("sed_depth_cm_"), as.numeric)
    )
}
```

## GPS Lookup

```{r gps}
hardbottom_gps <- read_excel(gps_file, sheet = "Hardbottom") %>%
  clean_names() %>%
  transmute(
    habitat = "hardbottom",
    transect = normalize_transect_id(str_remove(station_transect_id, "\\s+\\d+m$")),
    meter = as.integer(str_extract(station_transect_id, "\\d+(?=m$)")),
    gps_note = notes,
    lat_dd,
    lon_dd
  ) %>%
  filter(!is.na(transect), !is.na(meter))

seagrass_gps <- read_excel(gps_file, sheet = "Seagrass") %>%
  clean_names() %>%
  transmute(
    habitat = "seagrass",
    transect = normalize_transect_id(name),
    meter = NA_integer_,
    gps_note = location,
    lat_dd,
    lon_dd
  )

transect_meter_gps <- hardbottom_gps %>%
  filter(n() > 1, .by = transect) %>%
  arrange(transect, meter) %>%
  group_by(transect) %>%
  group_modify(~ {
    meter_seq <- seq(min(.x$meter), max(.x$meter))

    tibble(
      meter = meter_seq,
      lat_dd = approx(.x$meter, .x$lat_dd, xout = meter_seq, rule = 2)$y,
      lon_dd = approx(.x$meter, .x$lon_dd, xout = meter_seq, rule = 2)$y
    )
  }) %>%
  ungroup()
```

## Raw Sediment Data

The monitoring workbooks are cumulative, so later files repeat earlier surveys.
After reading all workbooks, this keeps the most recent source for each
survey/transect/meter record while retaining source metadata for traceability.

```{r sediment-raw}
sediment_raw_sources <- map_dfr(sediment_files, read_sediment_raw) %>%
  mutate(
    source_max_survey_n = max(survey_n, na.rm = TRUE),
    n_replicates_present = rowSums(!is.na(pick(starts_with("sed_depth_cm_")))),
    .by = source_file
  )

sediment_raw <- sediment_raw_sources %>%
  arrange(desc(source_max_survey_n), desc(source_modified), source_file_name) %>%
  distinct(
    date,
    survey,
    survey_n,
    transect_type,
    transect,
    meter,
    sed_depth_cm_1,
    sed_depth_cm_2,
    sed_depth_cm_3,
    .keep_all = TRUE
  ) %>%
  arrange(survey_n, transect, meter)

sediment_replicates <- sediment_raw %>%
  pivot_longer(
    starts_with("sed_depth_cm_"),
    names_to = "replicate",
    values_to = "sediment_depth_cm",
    values_drop_na = TRUE
  ) %>%
  mutate(
    replicate = as.integer(str_extract(replicate, "\\d+$")),
    sediment_depth_class = threshold_class(sediment_depth_cm)
  ) %>%
  left_join(transect_meter_gps, by = c("transect", "meter"))
```

## Baseline Differences

Sedimentation impact is defined in the Biological Monitoring Plan as an increase
in mean sediment depth of 1 cm or more above mean baseline sediment depth. This
chunk creates both raw replicate-level deltas and meter-level mean deltas so
individual high measurements remain visible before any transect averaging.

```{r baseline-differences}
baseline_by_meter <- sediment_raw %>%
  filter(survey == "Baseline") %>%
  transmute(
    transect_type,
    transect,
    meter,
    baseline_mean_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    ),
    baseline_max_cm = do.call(
      pmax,
      c(pick(starts_with("sed_depth_cm_")), list(na.rm = TRUE))
    )
  )

sediment_replicates_processed <- sediment_replicates %>%
  left_join(
    baseline_by_meter,
    by = c("transect_type", "transect", "meter")
  ) %>%
  mutate(
    sediment_delta_from_baseline_cm = sediment_depth_cm - baseline_mean_cm,
    sediment_delta_class = threshold_class(sediment_delta_from_baseline_cm),
    raw_gt_1cm_above_baseline = sediment_delta_from_baseline_cm >= 1
  )

sediment_meter_processed <- sediment_raw %>%
  mutate(
    mean_sediment_depth_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    ),
    max_sediment_depth_cm = do.call(
      pmax,
      c(pick(starts_with("sed_depth_cm_")), list(na.rm = TRUE))
    ),
    n_replicates = rowSums(!is.na(pick(starts_with("sed_depth_cm_"))))
  ) %>%
  select(
    date,
    survey,
    survey_n,
    transect_type,
    transect,
    model_sedimentation_class,
    meter,
    expected_permanent_length_m,
    mean_sediment_depth_cm,
    max_sediment_depth_cm,
    n_replicates,
    monitoring_phase,
    monitoring_type,
    source_file,
    source_file_name
  ) %>%
  left_join(baseline_by_meter, by = c("transect_type", "transect", "meter")) %>%
  left_join(transect_meter_gps, by = c("transect", "meter")) %>%
  mutate(
    mean_delta_from_baseline_cm = mean_sediment_depth_cm - baseline_mean_cm,
    max_delta_from_baseline_cm = max_sediment_depth_cm - baseline_mean_cm,
    mean_delta_class = threshold_class(mean_delta_from_baseline_cm),
    max_delta_class = threshold_class(max_delta_from_baseline_cm),
    sedimentation_impact = mean_delta_from_baseline_cm >= 1,
    main_50m_segment = transect %in% c("ST1", "ST38") & between(meter, 0, 50)
  )
```

## Transect Summaries

```{r transect-summaries}
sediment_transect_processed <- sediment_meter_processed %>%
  summarize(
    n_meters = n(),
    transect_mean_depth_cm = mean(mean_sediment_depth_cm, na.rm = TRUE),
    transect_mean_baseline_cm = mean(baseline_mean_cm, na.rm = TRUE),
    transect_mean_delta_cm = transect_mean_depth_cm - transect_mean_baseline_cm,
    max_meter_mean_delta_cm = max(mean_delta_from_baseline_cm, na.rm = TRUE),
    max_raw_delta_cm = max(max_delta_from_baseline_cm, na.rm = TRUE),
    n_meters_mean_gt_1cm = sum(mean_delta_from_baseline_cm >= 1, na.rm = TRUE),
    pct_meters_mean_gt_1cm = n_meters_mean_gt_1cm / n_meters,
    sedimentation_impact = transect_mean_delta_cm >= 1,
    main_50m_sedimentation_impact = any(
      main_50m_segment & mean_delta_from_baseline_cm >= 1,
      na.rm = TRUE
    ),
    .by = c(
      date,
      survey,
      survey_n,
      transect_type,
      transect,
      model_sedimentation_class
    )
  ) %>%
  arrange(survey_n, model_sedimentation_class, transect)
```

## Quick Visual Checks

The raw sediment sheets contain 19 transects total. Six adjacent transects are
Baseline-only, so construction-only plots show the 13 permanent transects.

```{r raw-coverage-plot, fig.width=11, fig.height=6}
sediment_raw %>%
  distinct(transect_type, transect, survey_n, survey) %>%
  mutate(
    transect = fct_reorder(transect, survey_n, .fun = max),
    survey = fct_reorder(survey, survey_n)
  ) %>%
  ggplot(aes(x = survey, y = transect, fill = transect_type)) +
  geom_tile(color = "white", linewidth = 0.2) +
  scale_fill_manual(
    values = c("Adjacent" = "#7f8c8d", "Permanent" = "#2a9d8f")
  ) +
  labs(
    x = "Survey event",
    y = "Transect",
    fill = "Transect type",
    title = "Raw survey coverage by transect and week"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r raw-baseline-depth-plot, fig.width=11, fig.height=7}
sediment_replicates %>%
  filter(survey == "Baseline") %>%
  ggplot(aes(x = meter, y = sediment_depth_cm)) +
  geom_point(aes(color = sediment_depth_class), alpha = 0.55, size = 0.7) +
  facet_wrap(~ model_sedimentation_class + transect, scales = "free_x") +
  scale_color_manual(
    values = c(
      "below baseline" = "grey90",
      "0 to 1.5 mm" = "#8ecae6",
      ">1.5 to 5 mm" = "#ffb703",
      ">5 mm to 1 cm" = "#fb8500",
      "1-5 cm" = "#d00000",
      "5-10 cm" = "#9d0208",
      ">10 cm" = "#370617"
    ),
    drop = FALSE
  ) +
  labs(
    x = "Transect meter",
    y = "Raw sediment depth (cm)",
    color = "Raw depth class",
    title = "Baseline raw sediment depth by transect"
  ) +
  theme_classic()
```

```{r construction-delta-plot, fig.width=11, fig.height=7}
sediment_meter_processed %>%
  filter(survey != "Baseline", transect_type == "Permanent") %>%
  mutate(
    plot_max_meter_m = if_else(
      transect %in% c("ST1", "ST38"),
      50L,
      expected_permanent_length_m
    )
  ) %>%
  {
    meter_grid <- distinct(
      .,
      date,
      survey,
      survey_n,
      transect_type,
      transect,
      model_sedimentation_class,
      expected_permanent_length_m,
      plot_max_meter_m
    ) %>%
      mutate(meter = map(plot_max_meter_m, ~ seq(0L, .x))) %>%
      unnest(meter)

    right_join(
      .,
      meter_grid,
      by = c(
        "date",
        "survey",
        "survey_n",
        "transect_type",
        "transect",
        "model_sedimentation_class",
        "expected_permanent_length_m",
        "plot_max_meter_m",
        "meter"
      )
    )
  } %>%
  mutate(
    transect = fct_reorder(
      transect,
      as.integer(model_sedimentation_class) * 1000 + parse_number(transect),
      .fun = min
    )
  ) %>%
  ggplot(aes(x = meter, y = survey_n, fill = mean_delta_from_baseline_cm)) +
  geom_tile() +
  facet_wrap(~ model_sedimentation_class + transect, scales = "free_x") +
  scale_fill_gradient2(
    low = "#08306b",
    mid = "gray90",
    high = "#7f0000",
    midpoint = 0,
    trans = scales::pseudo_log_trans(sigma = 0.15),
    na.value = "grey95",
    name = "Mean delta from baseline (cm)"
  ) +
  labs(
    x = "Transect meter",
    y = "Construction survey week number",
    title = "Construction sediment delta from baseline by transect"
  ) +
  theme_classic()
```

## Reef-scale Construction Delta Summaries

This summarizes construction-period sediment deltas by survey and model
sedimentation class. Percentages are calculated among measured meter positions;
unsampled expected meter positions are retained in the tables for coverage
tracking but omitted from the threshold-composition figure.

```{r reef-delta-summary, fig.width=10, fig.height=6}
threshold_levels <- levels(threshold_class(0))
construction_reef_delta_grid <- sediment_meter_processed %>%
  filter(survey != "Baseline", transect_type == "Permanent") %>%
  mutate(
    analysis_max_meter_m = if_else(
      transect %in% c("ST1", "ST38"),
      50L,
      expected_permanent_length_m
    )
  ) %>%
  {
    meter_grid <- distinct(
      .,
      date,
      survey,
      survey_n,
      transect_type,
      transect,
      model_sedimentation_class,
      expected_permanent_length_m,
      analysis_max_meter_m
    ) %>%
      mutate(meter = map(analysis_max_meter_m, ~ seq(0L, .x))) %>%
      unnest(meter)

    right_join(
      .,
      meter_grid,
      by = c(
        "date",
        "survey",
        "survey_n",
        "transect_type",
        "transect",
        "model_sedimentation_class",
        "expected_permanent_length_m",
        "analysis_max_meter_m",
        "meter"
      )
    )
  } %>%
  mutate(
    mean_delta_class = factor(
      if_else(
        is.na(mean_delta_class),
        "unmeasured",
        as.character(mean_delta_class)
      ),
      levels = c(threshold_levels, "unmeasured")
    )
  )

reef_delta_denominators <- construction_reef_delta_grid %>%
  summarize(
    survey_start_date = min(date, na.rm = TRUE),
    survey_end_date = max(date, na.rm = TRUE),
    survey_dates = paste(sort(unique(date)), collapse = ", "),
    expected_meter_positions = n(),
    measured_meter_positions = sum(!is.na(mean_delta_from_baseline_cm)),
    pct_meter_positions_measured = measured_meter_positions / expected_meter_positions,
    .by = c(survey, survey_n, model_sedimentation_class)
  )

reef_delta_summary_by_week_class <- construction_reef_delta_grid %>%
  filter(!is.na(mean_delta_from_baseline_cm)) %>%
  summarize(
    mean_delta_cm = mean(mean_delta_from_baseline_cm),
    p01_delta_cm = quantile(mean_delta_from_baseline_cm, 0.01),
    p10_delta_cm = quantile(mean_delta_from_baseline_cm, 0.10),
    p25_delta_cm = quantile(mean_delta_from_baseline_cm, 0.25),
    median_delta_cm = median(mean_delta_from_baseline_cm),
    p75_delta_cm = quantile(mean_delta_from_baseline_cm, 0.75),
    p90_delta_cm = quantile(mean_delta_from_baseline_cm, 0.90),
    p99_delta_cm = quantile(mean_delta_from_baseline_cm, 0.99),
    min_delta_cm = min(mean_delta_from_baseline_cm),
    max_delta_cm = max(mean_delta_from_baseline_cm),
    range_delta_cm = max_delta_cm - min_delta_cm,
    .by = c(survey, survey_n, model_sedimentation_class)
  ) %>%
  left_join(
    reef_delta_denominators,
    by = c("survey", "survey_n", "model_sedimentation_class")
  ) %>%
  arrange(survey_n, model_sedimentation_class)
```

```{r reef-delta-summary-figures, fig.width=11, fig.height=7}
reef_delta_summary_plot <- reef_delta_summary_by_week_class %>%
  select(
    survey_start_date,
    survey_n,
    model_sedimentation_class,
    p01_delta_cm,
    p10_delta_cm,
    p25_delta_cm,
    median_delta_cm,
    p75_delta_cm,
    p90_delta_cm,
    p99_delta_cm
  ) %>%
  pivot_longer(
    ends_with("_delta_cm"),
    names_to = "delta_percentile",
    values_to = "delta_cm"
  ) %>%
  mutate(
    delta_percentile = recode(
      delta_percentile,
      p01_delta_cm = "1st percentile",
      p10_delta_cm = "10th percentile",
      p25_delta_cm = "25th percentile (25% at or below)",
      median_delta_cm = "Median",
      p75_delta_cm = "75th percentile (25% at or above)",
      p90_delta_cm = "90th percentile",
      p99_delta_cm = "99th percentile"
    ),
    delta_percentile = factor(
      delta_percentile,
      levels = c(
        "1st percentile",
        "10th percentile",
        "25th percentile (25% at or below)",
        "Median",
        "75th percentile (25% at or above)",
        "90th percentile",
        "99th percentile"
      )
    )
  ) %>%
  arrange(model_sedimentation_class, delta_percentile, survey_start_date) %>%
  mutate(
    monitoring_gap_days = as.numeric(survey_start_date - lag(survey_start_date)),
    plot_segment = cumsum(coalesce(monitoring_gap_days > 14, FALSE)),
    .by = c(model_sedimentation_class, delta_percentile)
  ) %>%
  ggplot(aes(x = survey_n, y = delta_cm, color = delta_percentile)) +
  geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
  geom_hline(
    data = model_sedimentation_predictions,
    aes(yintercept = predicted_total_sedimentation_cm),
    linetype = "dashed",
    color = "black",
    linewidth = 0.4
  ) +
  geom_line(aes(group = interaction(delta_percentile, plot_segment)), linewidth = 0.7) +
  facet_wrap(~ model_sedimentation_class) +
  scale_color_manual(
    values = c(
      "1st percentile" = "#2166ac",
      "10th percentile" = "#67a9cf",
      "25th percentile (25% at or below)" = "#d1e5f0",
      "Median" = "#2a9d8f",
      "75th percentile (25% at or above)" = "#fb8500",
      "90th percentile" = "#d00000",
      "99th percentile" = "#370617"
    )
  ) +
  labs(
    x = "Construction survey",
    y = "Sediment delta from baseline (cm)",
    color = "Measured reef percentile",
    title = "Construction sediment delta percentiles by model sedimentation class"
  ) +
  theme_classic()

reef_delta_summary_plot
```

## Net Positive Exceedance Tests

These tests compare the measured area with positive sedimentation above a
threshold to the measured area with negative sedimentation below the equivalent
threshold. The `outside` class is outside the model-predicted sedimentation
footprint, but it is not treated as an unimpacted reference because it may still
receive dredging-related sediment. The bootstrap resamples transects within
each survey and model sedimentation class, so confidence intervals reflect
transect-level clustering rather than treating every meter as fully independent.

Interpretation: this is an area-balance screening metric. Positive values mean
more measured reef area exceeded a positive sedimentation threshold than fell
below the same magnitude negative threshold. It is best for asking whether
sediment accumulation above specific, interpretable thresholds is spatially more
common than equivalent sediment loss.

```{r net-positive-exceedance-tests, fig.width=11, fig.height=7}
set.seed(20260429)

net_exceedance_bootstrap_n <- 1000

net_exceedance_thresholds <- tibble(
  threshold_cm = c(0.15, 0.5, 1),
  threshold_label = factor(
    c("0.15 cm (1.5 mm)", "0.5 cm", "1.0 cm"),
    levels = c("0.15 cm (1.5 mm)", "0.5 cm", "1.0 cm")
  )
)

bootstrap_net_exceedance <- function(data, threshold_cm, n_boot) {
  transects <- sort(unique(data$transect))

  map_dfr(seq_len(n_boot), function(boot_i) {
    boot_delta <- sample(transects, length(transects), replace = TRUE) %>%
      map(~ data$mean_delta_from_baseline_cm[data$transect == .x]) %>%
      list_c()

    tibble(
      boot_i = boot_i,
      positive_area = mean(boot_delta >= threshold_cm),
      negative_area = mean(boot_delta <= -threshold_cm),
      net_positive_area = positive_area - negative_area
    )
  })
}

construction_measured_deltas <- construction_reef_delta_grid %>%
  filter(!is.na(mean_delta_from_baseline_cm))

net_exceedance_observed <- construction_measured_deltas %>%
  cross_join(net_exceedance_thresholds) %>%
  summarize(
    survey_start_date = min(date, na.rm = TRUE),
    survey_end_date = max(date, na.rm = TRUE),
    survey_dates = paste(sort(unique(date)), collapse = ", "),
    measured_meter_positions = n(),
    n_transects = n_distinct(transect),
    positive_meter_positions = sum(mean_delta_from_baseline_cm >= threshold_cm),
    negative_meter_positions = sum(mean_delta_from_baseline_cm <= -threshold_cm),
    positive_area = positive_meter_positions / measured_meter_positions,
    negative_area = negative_meter_positions / measured_meter_positions,
    net_positive_area = positive_area - negative_area,
    .by = c(
      survey,
      survey_n,
      model_sedimentation_class,
      threshold_cm,
      threshold_label
    )
  )

net_exceedance_bootstrap <- construction_measured_deltas %>%
  cross_join(net_exceedance_thresholds) %>%
  group_by(survey, survey_n, model_sedimentation_class, threshold_cm, threshold_label) %>%
  group_modify(
    ~ bootstrap_net_exceedance(
      .x,
      threshold_cm = .y$threshold_cm,
      n_boot = net_exceedance_bootstrap_n
    )
  ) %>%
  ungroup()

net_exceedance_ci <- net_exceedance_bootstrap %>%
  summarize(
    net_positive_area_lcl = quantile(net_positive_area, 0.025),
    net_positive_area_ucl = quantile(net_positive_area, 0.975),
    .by = c(
      survey,
      survey_n,
      model_sedimentation_class,
      threshold_cm,
      threshold_label
    )
  )

net_exceedance_by_week_class <- net_exceedance_observed %>%
  left_join(
    net_exceedance_ci,
    by = c(
      "survey",
      "survey_n",
      "model_sedimentation_class",
      "threshold_cm",
      "threshold_label"
    )
  ) %>%
  mutate(
    ci_note = if_else(
      n_transects < 3,
      "CI not estimated: fewer than 3 transects",
      "Transect bootstrap CI"
    ),
    across(
      c(net_positive_area_lcl, net_positive_area_ucl),
      ~ if_else(n_transects < 3, NA_real_, .x)
    ),
    evidence_positive_asymmetry = net_positive_area_lcl > 0
  ) %>%
  arrange(survey_n, model_sedimentation_class, threshold_cm)

```

```{r net-positive-exceedance-figures, fig.width=11, fig.height=7}
net_exceedance_threshold_plot_data <- net_exceedance_by_week_class %>%
  filter(threshold_cm %in% c(0.15, 1.0)) %>%
  arrange(model_sedimentation_class, threshold_label, survey_start_date) %>%
  mutate(
    monitoring_gap_days = as.numeric(survey_start_date - lag(survey_start_date)),
    plot_segment = cumsum(coalesce(monitoring_gap_days > 14, FALSE)),
    .by = c(model_sedimentation_class, threshold_label)
  )

positive_asymmetry_annotations <- net_exceedance_threshold_plot_data %>%
  filter(evidence_positive_asymmetry) %>%
  mutate(annotation_y = net_positive_area_ucl)

net_exceedance_threshold_plot <- net_exceedance_threshold_plot_data %>%
  ggplot(aes(x = survey_n, y = net_positive_area, color = threshold_label)) +
  geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
  geom_ribbon(
    data = ~ filter(.x, !is.na(net_positive_area_lcl), !is.na(net_positive_area_ucl)),
    aes(
      ymin = net_positive_area_lcl,
      ymax = net_positive_area_ucl,
      fill = threshold_label,
      group = interaction(threshold_label, plot_segment)
    ),
    alpha = 0.14,
    color = NA
  ) +
  geom_line(
    aes(group = interaction(threshold_label, plot_segment)),
    linewidth = 0.8
  ) +
  geom_text(
    data = positive_asymmetry_annotations,
    aes(x = survey_n, y = annotation_y, label = "*"),
    color = "black",
    vjust = -0.35,
    size = 4,
    show.legend = FALSE
  ) +
  facet_wrap(~ model_sedimentation_class) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(
    values = c(
      "0.15 cm (1.5 mm)" = "#fb8500",
      "1.0 cm" = "#d00000"
    )
  ) +
  scale_fill_manual(
    values = c(
      "0.15 cm (1.5 mm)" = "#fb8500",
      "1.0 cm" = "#d00000"
    )
  ) +
  labs(
    x = "Construction survey",
    y = "Net measured reef area",
    color = "Delta threshold",
    fill = "Delta threshold",
    title = "Net positive exceedance area",
    subtitle = paste(
      "Area above +threshold minus area below -threshold;",
      "positive values indicate more widespread sediment gain than loss"
    ),
    caption = paste(
      "* Bootstrap CI for net positive area is entirely above zero.",
      "CIs are only shown for survey/class groups with at least 3 transects;",
      "the 1.0 cm and 1.5 mm predicted classes each have 2 transects."
    )
  ) +
  theme_classic()

net_exceedance_threshold_plot
```

## Focused Attribution Analyses

The net-positive exceedance summaries are useful screening tools, but they are
not a complete attribution analysis. The following sections explore complementary
ways to ask whether construction-period deltas show directional sediment
accumulation. These are deliberately exploratory: each method answers a slightly
different question and has different assumptions.

The net-positive exceedance plot asks an area question: how much measured reef
area is above a positive threshold after subtracting area below the equivalent
negative threshold? The week-specific tail-balance plot below asks a different
magnitude question: when sediment gains occur in the upper tail of the weekly
distribution, are those gains larger than the corresponding sediment-loss tail?
The simpler tail-asymmetry diagnostic was removed because it repeated this tail
comparison without the bootstrap confidence intervals.

### Week-specific Tail-balance Tests

This approach keeps construction survey as a time series while explicitly
accounting for opposite negative deltas. For each survey and predicted
sedimentation class, it compares matched positive and negative tails:
`p75 - abs(p25)`, `p90 - abs(p10)`, and `p95 - abs(p05)`. Positive values mean
the sediment-gain tail is larger than the sediment-loss tail for that week. The
bootstrap resamples transects, and confidence intervals are not estimated when
fewer than three transects are available in a survey/class group.

Interpretation: this is a magnitude-asymmetry diagnostic. It is less directly
spatial than the net-exceedance plot, but it can show whether the largest
positive sediment deltas are more extreme than the largest negative deltas, even
when the affected area is limited.

```{r week-specific-tail-balance-tests, fig.width=11, fig.height=7}
week_tail_balance_bootstrap_n <- 100

summarize_tail_balance <- function(delta) {
  delta_quantiles <- quantile(
    delta,
    probs = c(0.05, 0.10, 0.25, 0.75, 0.90, 0.95),
    names = TRUE
  )

  tibble(
    tail_contrast = factor(
      c("75th minus |25th|", "90th minus |10th|", "95th minus |5th|"),
      levels = c("75th minus |25th|", "90th minus |10th|", "95th minus |5th|")
    ),
    lower_percentile = c(0.25, 0.10, 0.05),
    upper_percentile = c(0.75, 0.90, 0.95),
    lower_delta_cm = as.numeric(delta_quantiles[c("25%", "10%", "5%")]),
    upper_delta_cm = as.numeric(delta_quantiles[c("75%", "90%", "95%")]),
    upper_tail_excess_cm = upper_delta_cm - abs(lower_delta_cm)
  )
}

bootstrap_tail_balance <- function(data, n_boot) {
  transects <- sort(unique(data$transect))

  map_dfr(seq_len(n_boot), function(boot_i) {
    boot_delta <- sample(transects, length(transects), replace = TRUE) %>%
      map(~ data$mean_delta_from_baseline_cm[data$transect == .x]) %>%
      list_c()

    summarize_tail_balance(boot_delta) %>%
      mutate(boot_i = boot_i, .before = 1)
  })
}

week_tail_balance_observed <- construction_measured_deltas %>%
  summarize(
    survey_start_date = min(date, na.rm = TRUE),
    survey_end_date = max(date, na.rm = TRUE),
    survey_dates = paste(sort(unique(date)), collapse = ", "),
    measured_meter_positions = n(),
    n_transects = n_distinct(transect),
    tail_balance = list(summarize_tail_balance(mean_delta_from_baseline_cm)),
    .by = c(survey, survey_n, model_sedimentation_class)
  ) %>%
  unnest(tail_balance)

week_tail_balance_bootstrap <- construction_measured_deltas %>%
  group_by(survey, survey_n, model_sedimentation_class) %>%
  group_modify(~ bootstrap_tail_balance(.x, week_tail_balance_bootstrap_n)) %>%
  ungroup()

week_tail_balance_ci <- week_tail_balance_bootstrap %>%
  summarize(
    upper_tail_excess_lcl = quantile(upper_tail_excess_cm, 0.025),
    upper_tail_excess_ucl = quantile(upper_tail_excess_cm, 0.975),
    .by = c(survey, survey_n, model_sedimentation_class, tail_contrast)
  )

week_tail_balance_tests <- week_tail_balance_observed %>%
  left_join(
    week_tail_balance_ci,
    by = c("survey", "survey_n", "model_sedimentation_class", "tail_contrast")
  ) %>%
  mutate(
    ci_note = if_else(
      n_transects < 3,
      "CI not estimated: fewer than 3 transects",
      "Transect bootstrap CI"
    ),
    across(
      c(upper_tail_excess_lcl, upper_tail_excess_ucl),
      ~ if_else(n_transects < 3, NA_real_, .x)
    ),
    evidence_positive_tail_asymmetry = upper_tail_excess_lcl > 0
  ) %>%
  arrange(survey_n, model_sedimentation_class, tail_contrast)

week_tail_balance_plot_data <- week_tail_balance_tests %>%
  arrange(model_sedimentation_class, tail_contrast, survey_start_date) %>%
  mutate(
    monitoring_gap_days = as.numeric(survey_start_date - lag(survey_start_date)),
    plot_segment = cumsum(coalesce(monitoring_gap_days > 14, FALSE)),
    .by = c(model_sedimentation_class, tail_contrast)
  )

week_tail_balance_annotations <- week_tail_balance_plot_data %>%
  filter(evidence_positive_tail_asymmetry) %>%
  mutate(annotation_y = upper_tail_excess_ucl)

week_tail_balance_plot <- week_tail_balance_plot_data %>%
  ggplot(aes(x = survey_n, y = upper_tail_excess_cm, color = tail_contrast)) +
  geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
  geom_ribbon(
    data = ~ filter(.x, !is.na(upper_tail_excess_lcl), !is.na(upper_tail_excess_ucl)),
    aes(
      ymin = upper_tail_excess_lcl,
      ymax = upper_tail_excess_ucl,
      fill = tail_contrast,
      group = interaction(tail_contrast, plot_segment)
    ),
    alpha = 0.12,
    color = NA
  ) +
  geom_line(
    aes(group = interaction(tail_contrast, plot_segment)),
    linewidth = 0.75
  ) +
  geom_text(
    data = week_tail_balance_annotations,
    aes(x = survey_n, y = annotation_y, label = "*"),
    color = "black",
    vjust = -0.35,
    size = 4,
    show.legend = FALSE
  ) +
  facet_wrap(~ model_sedimentation_class) +
  labs(
    x = "Construction survey",
    y = "Positive-tail excess over matched negative tail (cm)",
    color = "Tail contrast",
    fill = "Tail contrast",
    title = "Week-specific sediment-gain tail balance",
    subtitle = paste(
      "Positive tail percentile minus absolute matched negative tail percentile;",
      "positive values indicate larger sediment-gain extremes"
    ),
    caption = paste(
      "* Transect bootstrap CI for upper-tail excess is entirely above zero.",
      "CIs are only shown for survey/class groups with at least 3 transects."
    )
  ) +
  theme_classic()

week_tail_balance_plot
```
