---
title: "Sedimentation Monitoring"
author: "Ross Cunning"
date: "2026-04-29"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
```

## Import

```{r paths}
sediment_dir <- "data/sedimentation_monitoring"
processed_dir <- "data/processed"
audit_dir <- "analysis/sedimentation_raw_data_audit"

dir.create(processed_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(audit_dir, showWarnings = FALSE, recursive = TRUE)

gps_file <- file.path(sediment_dir, "PE_SB_Station_Transect_GPS.xlsx")

sediment_files <- list.files(
  sediment_dir,
  pattern = "xlsx$",
  recursive = TRUE,
  full.names = TRUE
) %>%
  discard(~ basename(.x) == basename(gps_file))
```

```{r helpers}
excel_date <- function(x) {
  if (inherits(x, "Date")) return(x)
  if (inherits(x, "POSIXt")) return(as.Date(x))

  suppressWarnings(as.Date(as.numeric(x), origin = "1899-12-30"))
}

survey_number <- function(x) {
  if_else(
    x == "Baseline",
    0L,
    suppressWarnings(as.integer(str_extract(x, "\\d+")))
  )
}

normalize_transect_id <- function(x) {
  x %>%
    str_squish() %>%
    str_replace_all("\\s+", "_")
}

model_sedimentation_class <- function(transect) {
  case_when(
    transect == "ST22" ~ "solid_blue",
    transect == "ST31" ~ "dashed_red",
    transect %in% c("ST28", "ST5") ~ "solid_red",
    TRUE ~ "outside"
  ) %>%
    factor(
      levels = c("solid_blue", "dashed_red", "solid_red", "outside"),
      ordered = TRUE
    )
}

threshold_class <- function(x) {
  case_when(
    is.na(x) ~ NA_character_,
    x > 1 ~ ">1 cm",
    x > 0.5 ~ ">5 mm to 1 cm",
    x > 0.15 ~ ">1.5 to 5 mm",
    x >= 0 ~ "0 to 1.5 mm",
    TRUE ~ "below baseline"
  ) %>%
    factor(
      levels = c(
        "below baseline",
        "0 to 1.5 mm",
        ">1.5 to 5 mm",
        ">5 mm to 1 cm",
        ">1 cm"
      )
    )
}

read_sediment_raw <- function(path) {
  read_excel(path, sheet = "SedimentDepth_RawData", guess_max = 20000) %>%
    clean_names() %>%
    mutate(
      source_file = path,
      source_file_name = basename(path),
      source_modified = file.info(path)$mtime,
      monitoring_phase = case_when(
        str_detect(path, "1_Pre-Construction") ~ "pre_construction",
        str_detect(path, "2_During-Construction") ~ "during_construction",
        TRUE ~ "unknown"
      ),
      monitoring_type = case_when(
        str_detect(path, "Weekly Monitoring") ~ "weekly",
        str_detect(path, "Exceedance Triggered Monitoring") ~ "exceedance_triggered",
        TRUE ~ "compiled"
      ),
      date = excel_date(date),
      survey = str_squish(survey),
      survey_n = survey_number(survey),
      transect_type = str_squish(transect_type),
      transect = normalize_transect_id(transect),
      model_sedimentation_class = model_sedimentation_class(transect),
      meter = as.integer(meter),
      expected_permanent_length_m = case_when(
        transect_type == "Permanent" & transect %in% c("ST1", "ST38") ~ 150L,
        transect_type == "Permanent" ~ 50L,
        TRUE ~ NA_integer_
      ),
      across(starts_with("sed_depth_cm_"), as.numeric)
    )
}
```

## GPS Lookup

```{r gps}
hardbottom_gps <- read_excel(gps_file, sheet = "Hardbottom") %>%
  clean_names() %>%
  transmute(
    habitat = "hardbottom",
    transect = normalize_transect_id(str_remove(station_transect_id, "\\s+\\d+m$")),
    meter = as.integer(str_extract(station_transect_id, "\\d+(?=m$)")),
    gps_note = notes,
    lat_dd,
    lon_dd
  ) %>%
  filter(!is.na(transect), !is.na(meter))

seagrass_gps <- read_excel(gps_file, sheet = "Seagrass") %>%
  clean_names() %>%
  transmute(
    habitat = "seagrass",
    transect = normalize_transect_id(name),
    meter = NA_integer_,
    gps_note = location,
    lat_dd,
    lon_dd
  )

transect_meter_gps <- hardbottom_gps %>%
  filter(n() > 1, .by = transect) %>%
  arrange(transect, meter) %>%
  group_by(transect) %>%
  group_modify(~ {
    meter_seq <- seq(min(.x$meter), max(.x$meter))

    tibble(
      meter = meter_seq,
      lat_dd = approx(.x$meter, .x$lat_dd, xout = meter_seq, rule = 2)$y,
      lon_dd = approx(.x$meter, .x$lon_dd, xout = meter_seq, rule = 2)$y
    )
  }) %>%
  ungroup()
```

## Raw Sediment Data

The monitoring workbooks are cumulative, so later files repeat earlier surveys.
After reading all workbooks, this keeps the most recent source for each
survey/transect/meter record while retaining source metadata for traceability.

```{r sediment-raw}
sediment_raw_sources <- map_dfr(sediment_files, read_sediment_raw) %>%
  mutate(
    source_max_survey_n = max(survey_n, na.rm = TRUE),
    n_replicates_present = rowSums(!is.na(pick(starts_with("sed_depth_cm_")))),
    .by = source_file
  )

sediment_raw <- sediment_raw_sources %>%
  arrange(desc(source_max_survey_n), desc(source_modified), source_file_name) %>%
  distinct(
    date,
    survey,
    survey_n,
    transect_type,
    transect,
    meter,
    sed_depth_cm_1,
    sed_depth_cm_2,
    sed_depth_cm_3,
    .keep_all = TRUE
  ) %>%
  arrange(survey_n, transect, meter)

sediment_replicates <- sediment_raw %>%
  pivot_longer(
    starts_with("sed_depth_cm_"),
    names_to = "replicate",
    values_to = "sediment_depth_cm",
    values_drop_na = TRUE
  ) %>%
  mutate(
    replicate = as.integer(str_extract(replicate, "\\d+$")),
    sediment_depth_class = threshold_class(sediment_depth_cm)
  ) %>%
  left_join(transect_meter_gps, by = c("transect", "meter"))
```

## Raw Data Audit

These summaries scope only the `SedimentDepth_RawData` sheets. The processed
mean/difference sheets in the workbooks are not used.

```{r raw-data-audit}
raw_file_audit <- sediment_raw_sources %>%
  summarize(
    raw_rows = n(),
    unique_surveys = n_distinct(survey),
    surveys = paste(unique(survey[order(survey_n)]), collapse = ", "),
    n_transects = n_distinct(transect),
    n_transect_meter_positions = n_distinct(paste(transect, meter)),
    min_date = min(date, na.rm = TRUE),
    max_date = max(date, na.rm = TRUE),
    replicate_count_values = paste(sort(unique(n_replicates_present)), collapse = ", "),
    .by = c(monitoring_phase, monitoring_type, source_file_name)
  ) %>%
  arrange(monitoring_phase, monitoring_type, source_file_name)

raw_overall_audit <- sediment_raw %>%
  summarize(
    unique_raw_rows = n(),
    survey_events = n_distinct(survey),
    survey_date_combinations = n_distinct(paste(survey, date)),
    unique_dates = n_distinct(date),
    transects = n_distinct(transect),
    permanent_transects = n_distinct(transect[transect_type == "Permanent"]),
    adjacent_transects = n_distinct(transect[transect_type == "Adjacent"]),
    transect_meter_positions = n_distinct(paste(transect, meter)),
    replicate_columns = paste(
      names(select(sediment_raw, starts_with("sed_depth_cm_"))),
      collapse = ", "
    ),
    replicate_count_values = paste(sort(unique(n_replicates_present)), collapse = ", "),
    total_replicate_measurements = sum(n_replicates_present),
    min_date = min(date, na.rm = TRUE),
    max_date = max(date, na.rm = TRUE)
  )

raw_transect_audit <- sediment_raw %>%
  summarize(
    transect_type = first(transect_type),
    model_sedimentation_class = first(model_sedimentation_class),
    min_meter = min(meter),
    max_meter = max(meter),
    expected_permanent_length_m = if_else(
      all(is.na(expected_permanent_length_m)),
      NA_integer_,
      first(na.omit(expected_permanent_length_m))
    ),
    observed_length_m = max_meter - min_meter,
    length_matches_expected = if_else(
      transect_type == "Permanent",
      observed_length_m == expected_permanent_length_m,
      NA
    ),
    n_meters = n_distinct(meter),
    n_survey_events = n_distinct(survey),
    surveys = paste(unique(survey[order(survey_n)]), collapse = ", "),
    n_raw_meter_rows = n(),
    n_rep_measurements = sum(n_replicates_present),
    .by = transect
  ) %>%
  arrange(transect_type, parse_number(transect), transect)

raw_survey_audit <- sediment_raw %>%
  summarize(
    dates = paste(sort(unique(date)), collapse = ", "),
    n_dates = n_distinct(date),
    n_transects = n_distinct(transect),
    transects = paste(sort(unique(transect)), collapse = ", "),
    n_meter_rows = n(),
    n_permanent_transects = n_distinct(transect[transect_type == "Permanent"]),
    n_adjacent_transects = n_distinct(transect[transect_type == "Adjacent"]),
    total_rep_measurements = sum(n_replicates_present),
    .by = c(survey_n, survey)
  ) %>%
  arrange(survey_n)

raw_replicate_audit <- sediment_raw %>%
  summarize(
    n_meter_rows = n(),
    n_rep_1_present = sum(!is.na(sed_depth_cm_1)),
    n_rep_2_present = sum(!is.na(sed_depth_cm_2)),
    n_rep_3_present = sum(!is.na(sed_depth_cm_3)),
    min_reps_per_meter = min(n_replicates_present),
    max_reps_per_meter = max(n_replicates_present),
    rows_with_0_reps = sum(n_replicates_present == 0),
    rows_with_1_rep = sum(n_replicates_present == 1),
    rows_with_2_reps = sum(n_replicates_present == 2),
    rows_with_3_reps = sum(n_replicates_present == 3),
    .by = c(survey_n, survey)
  ) %>%
  arrange(survey_n)

raw_workbook_duplication_audit <- sediment_raw_sources %>%
  summarize(
    source_file_count = n_distinct(source_file),
    .by = c(
      date,
      survey,
      survey_n,
      transect_type,
      transect,
      meter,
      sed_depth_cm_1,
      sed_depth_cm_2,
      sed_depth_cm_3
    )
  ) %>%
  count(source_file_count, name = "unique_raw_rows") %>%
  arrange(source_file_count)

raw_value_conflict_audit <- sediment_raw_sources %>%
  summarize(
    n_value_versions = n_distinct(
      paste(sed_depth_cm_1, sed_depth_cm_2, sed_depth_cm_3)
    ),
    values = paste(
      unique(paste(sed_depth_cm_1, sed_depth_cm_2, sed_depth_cm_3, sep = "/")),
      collapse = " | "
    ),
    source_files = paste(sort(unique(source_file_name)), collapse = " | "),
    .by = c(date, survey, survey_n, transect_type, transect, meter)
  ) %>%
  filter(n_value_versions > 1) %>%
  arrange(survey_n, transect, meter)

raw_gps_audit <- sediment_raw %>%
  distinct(transect_type, transect) %>%
  left_join(
    hardbottom_gps %>%
      summarize(
        gps_points = n(),
        gps_meters = paste(sort(meter), collapse = ", "),
        .by = transect
      ),
    by = "transect"
  ) %>%
  arrange(transect_type, parse_number(transect), transect)

raw_file_audit
raw_overall_audit
raw_transect_audit
raw_survey_audit
raw_replicate_audit
raw_workbook_duplication_audit
raw_value_conflict_audit
raw_gps_audit
```

```{r export-raw-audit}
write_csv(raw_file_audit, file.path(audit_dir, "raw_file_audit.csv"))
write_csv(raw_overall_audit, file.path(audit_dir, "raw_overall_audit.csv"))
write_csv(raw_transect_audit, file.path(audit_dir, "raw_transect_audit.csv"))
write_csv(raw_survey_audit, file.path(audit_dir, "raw_survey_audit.csv"))
write_csv(raw_replicate_audit, file.path(audit_dir, "raw_replicate_audit.csv"))
write_csv(
  raw_workbook_duplication_audit,
  file.path(audit_dir, "raw_workbook_duplication_audit.csv")
)
write_csv(
  raw_value_conflict_audit,
  file.path(audit_dir, "raw_value_conflict_audit.csv")
)
write_csv(raw_gps_audit, file.path(audit_dir, "raw_gps_audit.csv"))
```

## Baseline Differences

Sedimentation impact is defined in the Biological Monitoring Plan as an increase
in mean sediment depth of 1 cm or more above mean baseline sediment depth. This
chunk creates both raw replicate-level deltas and meter-level mean deltas so
individual high measurements remain visible before any transect averaging.

```{r baseline-differences}
baseline_by_meter <- sediment_raw %>%
  filter(survey == "Baseline") %>%
  transmute(
    transect_type,
    transect,
    meter,
    baseline_mean_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    ),
    baseline_max_cm = do.call(
      pmax,
      c(pick(starts_with("sed_depth_cm_")), list(na.rm = TRUE))
    )
  )

sediment_replicates_processed <- sediment_replicates %>%
  left_join(
    baseline_by_meter,
    by = c("transect_type", "transect", "meter")
  ) %>%
  mutate(
    sediment_delta_from_baseline_cm = sediment_depth_cm - baseline_mean_cm,
    sediment_delta_class = threshold_class(sediment_delta_from_baseline_cm),
    raw_gt_1cm_above_baseline = sediment_delta_from_baseline_cm >= 1
  )

sediment_meter_processed <- sediment_raw %>%
  mutate(
    mean_sediment_depth_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    ),
    max_sediment_depth_cm = do.call(
      pmax,
      c(pick(starts_with("sed_depth_cm_")), list(na.rm = TRUE))
    ),
    n_replicates = rowSums(!is.na(pick(starts_with("sed_depth_cm_"))))
  ) %>%
  select(
    date,
    survey,
    survey_n,
    transect_type,
    transect,
    model_sedimentation_class,
    meter,
    expected_permanent_length_m,
    mean_sediment_depth_cm,
    max_sediment_depth_cm,
    n_replicates,
    monitoring_phase,
    monitoring_type,
    source_file,
    source_file_name
  ) %>%
  left_join(baseline_by_meter, by = c("transect_type", "transect", "meter")) %>%
  left_join(transect_meter_gps, by = c("transect", "meter")) %>%
  mutate(
    mean_delta_from_baseline_cm = mean_sediment_depth_cm - baseline_mean_cm,
    max_delta_from_baseline_cm = max_sediment_depth_cm - baseline_mean_cm,
    mean_delta_class = threshold_class(mean_delta_from_baseline_cm),
    max_delta_class = threshold_class(max_delta_from_baseline_cm),
    sedimentation_impact = mean_delta_from_baseline_cm >= 1,
    main_50m_segment = transect %in% c("ST1", "ST38") & between(meter, 0, 50)
  )
```

## Transect Summaries

```{r transect-summaries}
sediment_transect_processed <- sediment_meter_processed %>%
  summarize(
    n_meters = n(),
    transect_mean_depth_cm = mean(mean_sediment_depth_cm, na.rm = TRUE),
    transect_mean_baseline_cm = mean(baseline_mean_cm, na.rm = TRUE),
    transect_mean_delta_cm = transect_mean_depth_cm - transect_mean_baseline_cm,
    max_meter_mean_delta_cm = max(mean_delta_from_baseline_cm, na.rm = TRUE),
    max_raw_delta_cm = max(max_delta_from_baseline_cm, na.rm = TRUE),
    n_meters_mean_gt_1cm = sum(mean_delta_from_baseline_cm >= 1, na.rm = TRUE),
    pct_meters_mean_gt_1cm = n_meters_mean_gt_1cm / n_meters,
    sedimentation_impact = transect_mean_delta_cm >= 1,
    main_50m_sedimentation_impact = any(
      main_50m_segment & mean_delta_from_baseline_cm >= 1,
      na.rm = TRUE
    ),
    .by = c(
      date,
      survey,
      survey_n,
      transect_type,
      transect,
      model_sedimentation_class
    )
  ) %>%
  arrange(survey_n, model_sedimentation_class, transect)

sediment_transect_processed %>%
  filter(survey != "Baseline") %>%
  arrange(desc(max_raw_delta_cm))
```

## Export Processed Tables

```{r export}
write_csv(
  sediment_replicates_processed,
  file.path(processed_dir, "sedimentation_replicates_processed.csv")
)

write_csv(
  sediment_meter_processed,
  file.path(processed_dir, "sedimentation_meter_processed.csv")
)

write_csv(
  sediment_transect_processed,
  file.path(processed_dir, "sedimentation_transect_processed.csv")
)
```

## Quick Visual Checks

The raw sediment sheets contain 19 transects total. Six adjacent transects are
Baseline-only, so construction-only plots show the 13 permanent transects.

```{r raw-coverage-plot, fig.width=11, fig.height=6}
sediment_raw %>%
  distinct(transect_type, transect, survey_n, survey) %>%
  mutate(
    transect = fct_reorder(transect, survey_n, .fun = max),
    survey = fct_reorder(survey, survey_n)
  ) %>%
  ggplot(aes(x = survey, y = transect, fill = transect_type)) +
  geom_tile(color = "white", linewidth = 0.2) +
  scale_fill_manual(
    values = c("Adjacent" = "#7f8c8d", "Permanent" = "#2a9d8f")
  ) +
  labs(
    x = "Survey event",
    y = "Transect",
    fill = "Transect type"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r raw-baseline-depth-plot, fig.width=11, fig.height=7}
sediment_replicates %>%
  filter(survey == "Baseline") %>%
  ggplot(aes(x = meter, y = sediment_depth_cm)) +
  geom_point(aes(color = sediment_depth_class), alpha = 0.55, size = 0.7) +
  facet_wrap(~ transect, scales = "free_x") +
  scale_color_manual(
    values = c(
      "below baseline" = "grey90",
      "0 to 1.5 mm" = "#8ecae6",
      ">1.5 to 5 mm" = "#ffb703",
      ">5 mm to 1 cm" = "#fb8500",
      ">1 cm" = "#d00000"
    ),
    drop = FALSE
  ) +
  labs(
    x = "Transect meter",
    y = "Raw sediment depth (cm)",
    color = "Raw depth class"
  ) +
  theme_classic()
```

```{r construction-delta-plot, fig.width=11, fig.height=7}
sediment_meter_processed %>%
  filter(survey != "Baseline") %>%
  mutate(
    transect = fct_reorder(
      transect,
      as.integer(model_sedimentation_class) * 1000 + parse_number(transect),
      .fun = min
    )
  ) %>%
  ggplot(aes(x = meter, y = survey_n, fill = mean_delta_class)) +        # MEAN or MAX
  geom_tile() +
  facet_wrap(~ model_sedimentation_class + transect, scales = "free_x") +
  scale_fill_manual(
    values = c(
      "below baseline" = "grey90",
      "0 to 1.5 mm" = "#8ecae6",
      ">1.5 to 5 mm" = "#ffb703",
      ">5 mm to 1 cm" = "#fb8500",
      ">1 cm" = "#d00000"
    ),
    drop = FALSE
  ) +
  labs(
    x = "Transect meter",
    y = "Construction survey",
    fill = "Max delta from baseline"
  ) +
  theme_classic()
```
