---
title: "Localized Sedimentation Analysis"
author: "Ross Cunning"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(ggrepel)
library(ggspatial)
library(patchwork)

project_dir <- if (basename(getwd()) == "code") dirname(getwd()) else getwd()
knitr::opts_knit$set(root.dir = project_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
```

## Data sources

```{r paths-and-helpers}
sediment_dir <- "data/sedimentation_monitoring"
gps_file <- file.path(sediment_dir, "PE_SB_Station_Transect_GPS.xlsx")

normalize_transect_id <- function(x) {
  x %>%
    str_squish() %>%
    str_replace_all("\\s+", "_")
}

model_sedimentation_class <- function(transect) {
  case_when(
    transect %in% c("ST22", "ST31") ~ "1.0 cm predicted",
    transect %in% c("ST28", "ST5") ~ "1.5 mm predicted",
    TRUE ~ "0 cm predicted"
  ) %>%
    factor(
      levels = c(
        "1.0 cm predicted",
        "1.5 mm predicted",
        "0 cm predicted"
      ),
      ordered = TRUE
    )
}

construction_files <- list.files(
  file.path(sediment_dir, "2_During-Construction", "Data"),
  pattern = "\\.xlsx$",
  recursive = TRUE,
  full.names = TRUE
)

construction_file_index <- tibble(source_file = construction_files) %>%
  mutate(
    source_file_name = basename(source_file),
    construction_survey_n = map_int(
      str_extract_all(source_file_name, "(?i)(?<=con)\\d+"),
      ~ if (length(.x) == 0) NA_integer_ else max(as.integer(.x))
    )
  ) %>%
  filter(!is.na(construction_survey_n)) %>%
  arrange(desc(construction_survey_n), desc(file.info(source_file)$mtime))

latest_construction_file <- construction_file_index %>%
  slice(1) %>%
  pull(source_file)
```

The transect inventory is taken from the newest cumulative construction
workbook: `r basename(latest_construction_file)`.

```{r transect-inventory}
monitoring_transects <- read_excel(
  latest_construction_file,
  sheet = "SedimentDepth_RawData"
) %>%
  clean_names() %>%
  transmute(
    transect = normalize_transect_id(transect),
    transect_type = str_squish(transect_type)
  ) %>%
  distinct()

transect_gps <- read_excel(gps_file, sheet = "Hardbottom") %>%
  clean_names() %>%
  transmute(
    transect = normalize_transect_id(
      str_remove(station_transect_id, "\\s+\\d+m$")
    ),
    meter = as.integer(str_extract(station_transect_id, "\\d+(?=m$)")),
    gps_note = notes,
    lat_dd = as.numeric(lat_dd),
    lon_dd = as.numeric(lon_dd)
  ) %>%
  filter(!is.na(transect), !is.na(meter)) %>%
  arrange(transect, meter) %>%
  left_join(monitoring_transects, by = "transect") %>%
  mutate(model_sedimentation_class = model_sedimentation_class(transect))

transect_coverage <- full_join(
  monitoring_transects %>% mutate(in_monitoring_data = TRUE),
  transect_gps %>%
    summarize(
      n_gps_points = n(),
      min_meter = min(meter),
      max_meter = max(meter),
      coordinates_complete = all(!is.na(lat_dd) & !is.na(lon_dd)),
      .by = transect
    ) %>%
    mutate(in_gps_file = TRUE),
  by = "transect"
) %>%
  mutate(
    across(c(in_monitoring_data, in_gps_file), ~ replace_na(.x, FALSE)),
    map_ready = in_monitoring_data & in_gps_file & coordinates_complete
  ) %>%
  arrange(transect_type, transect)

stopifnot(
  nrow(monitoring_transects) == 19,
  all(transect_coverage$map_ready)
)

transect_coverage
```

## Transect map

Lines connect the surveyed GPS anchors in meter order. Most 50 m transects have
coordinates at both ends; longer and adjacent transects include intermediate
anchors where they were supplied in the GPS workbook.

```{r transect-map-data, fig.width=11, fig.height=8.5}
transect_labels <- transect_gps %>%
  summarize(
    lon_dd = mean(range(lon_dd)),
    lat_dd = mean(range(lat_dd)),
    transect_type = first(transect_type),
    model_sedimentation_class = first(model_sedimentation_class),
    .by = transect
  )

transect_map <- ggplot(
  transect_gps,
  aes(
    x = lon_dd,
    y = lat_dd,
    group = transect,
    color = model_sedimentation_class
  )
) +
  geom_path(linewidth = 1.05, lineend = "round") +
  geom_point(size = 1.7, shape = 21, fill = "white", stroke = 0.7) +
  geom_text_repel(
    data = transect_labels,
    aes(label = transect),
    size = 3.2,
    fontface = "bold",
    color = "grey15",
    seed = 1406,
    box.padding = 0.3,
    point.padding = 0.15,
    force = 1.5,
    max.iter = 20000,
    min.segment.length = 0,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  annotation_scale(
    location = "bl",
    width_hint = 0.22,
    text_cex = 0.75
  ) +
  annotation_north_arrow(
    location = "tl",
    which_north = "true",
    height = unit(0.75, "cm"),
    width = unit(0.75, "cm"),
    style = north_arrow_minimal(text_size = 8)
  ) +
  scale_color_manual(
    values = c(
      "1.0 cm predicted" = "#c1121f",
      "1.5 mm predicted" = "#f28e2b",
      "0 cm predicted" = "#287271"
    ),
    drop = FALSE
  ) +
  coord_sf(
    crs = sf::st_crs(4326),
    default_crs = sf::st_crs(4326),
    expand = TRUE,
    datum = sf::st_crs(4326)
  ) +
  labs(
    title = "Sand Bypass sediment monitoring transects",
    subtitle = "GPS anchors and transect alignments",
    x = "Longitude",
    y = "Latitude",
    color = "Predicted sedimentation",
    caption = "Source: PE_SB_Station_Transect_GPS.xlsx"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.box = "vertical",
    plot.title.position = "plot"
  )

transect_map
```

# Individual transect case studies

## Case study 1: ST22

### Sediment deltas

For each survey, the three sediment-depth replicates are averaged at each meter
position. The baseline mean at that same meter position is then subtracted from
the construction-survey mean:

\[
\Delta_{m,t} = \overline{x}_{m,t} - \overline{x}_{m,baseline}
\]

Positive values indicate sediment accumulation relative to baseline. This
meter-matched calculation retains spatial variation that would be lost by first
averaging across the full transect.

```{r st22-meter-deltas}
st22_raw <- read_excel(
  latest_construction_file,
  sheet = "SedimentDepth_RawData"
) %>%
  clean_names() %>%
  mutate(
    date = as.Date(date),
    survey = str_squish(survey),
    survey_n = if_else(
      survey == "Baseline",
      0L,
      as.integer(str_extract(survey, "\\d+"))
    ),
    transect = normalize_transect_id(transect),
    meter = as.integer(meter),
    meter_mean_sediment_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    )
  ) %>%
  filter(transect == "ST22")

st22_baseline <- st22_raw %>%
  filter(survey == "Baseline") %>%
  transmute(
    meter,
    baseline_mean_sediment_cm = meter_mean_sediment_cm
  )

st22_meter_deltas <- st22_raw %>%
  filter(survey != "Baseline") %>%
  left_join(st22_baseline, by = "meter") %>%
  mutate(
    sediment_delta_cm =
      meter_mean_sediment_cm - baseline_mean_sediment_cm
  ) %>%
  select(
    date,
    survey,
    survey_n,
    transect,
    meter,
    baseline_mean_sediment_cm,
    meter_mean_sediment_cm,
    sediment_delta_cm
  ) %>%
  arrange(survey_n, meter)

st22_delta_coverage <- st22_meter_deltas %>%
  summarize(
    n_meter_positions = n(),
    n_missing_deltas = sum(is.na(sediment_delta_cm)),
    .by = c(date, survey, survey_n)
  ) %>%
  arrange(survey_n)

stopifnot(
  all(st22_delta_coverage$n_meter_positions == 51),
  all(st22_delta_coverage$n_missing_deltas == 0)
)

st22_delta_coverage
```

### Alternative Baseline-Con1-Con2 reference

This sensitivity analysis defines a meter-specific alternative reference as
the arithmetic mean of the Baseline, Con1, and Con2 meter means. Each survey
receives equal weight. Because Con1 and Con2 contribute to the reference,
alternative-reference results begin with Con3. This does not replace the BMP
baseline analysis above.

```{r st22-alternative-reference}
st22_alternative_reference <- st22_raw %>%
  filter(survey_n %in% 0:2) %>%
  summarize(
    alternative_reference_cm = mean(meter_mean_sediment_cm),
    n_reference_surveys = n(),
    .by = meter
  ) %>%
  left_join(st22_baseline, by = "meter") %>%
  mutate(
    reference_difference_cm =
      alternative_reference_cm - baseline_mean_sediment_cm
  )

stopifnot(
  nrow(st22_alternative_reference) == 51,
  all(st22_alternative_reference$n_reference_surveys == 3)
)

st22_reference_summary <- st22_alternative_reference %>%
  summarize(
    n_meter_positions = n(),
    official_baseline_mean_cm = mean(baseline_mean_sediment_cm),
    alternative_reference_mean_cm = mean(alternative_reference_cm),
    mean_reference_difference_cm = mean(reference_difference_cm),
    median_reference_difference_cm = median(reference_difference_cm),
    pct_alternative_lower = mean(reference_difference_cm < 0),
    pct_alternative_higher = mean(reference_difference_cm > 0)
  )

st22_reference_summary
```

```{r st22-alternative-deltas}
st22_alternative_deltas <- st22_raw %>%
  filter(survey_n >= 3) %>%
  left_join(
    st22_alternative_reference %>%
      select(meter, alternative_reference_cm),
    by = "meter"
  ) %>%
  left_join(st22_baseline, by = "meter") %>%
  mutate(
    alternative_delta_cm =
      meter_mean_sediment_cm - alternative_reference_cm,
    official_delta_cm =
      meter_mean_sediment_cm - baseline_mean_sediment_cm
  ) %>%
  select(
    date,
    survey,
    survey_n,
    transect,
    meter,
    meter_mean_sediment_cm,
    baseline_mean_sediment_cm,
    alternative_reference_cm,
    official_delta_cm,
    alternative_delta_cm
  ) %>%
  arrange(survey_n, meter)
```

### Meter-position diagnostic

This heatmap provides a relatively raw view of sediment change before the
transect is summarized with exceedance curves. It shows whether accumulation
or loss is scattered among meter positions or forms spatially coherent patterns
that persist through multiple surveys. Results use the Baseline-Con1-Con2
alternative reference; outlined cells meet or exceed 1 cm accumulation.

```{r st22-meter-position-heatmap, fig.width=11, fig.height=7.5}
st22_spatial_deltas <- st22_alternative_deltas %>%
  mutate(
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n, .desc = TRUE)
  )

st22_spatial_color_limit <- max(
  abs(
    quantile(
      st22_spatial_deltas$alternative_delta_cm,
      probs = c(0.02, 0.98),
      na.rm = TRUE
    )
  )
)

st22_meter_position_heatmap <- ggplot(
  st22_spatial_deltas,
  aes(x = meter, y = survey_label, fill = alternative_delta_cm)
) +
  geom_tile() +
  geom_point(
    data = st22_spatial_deltas %>%
      filter(alternative_delta_cm >= 1),
    shape = 21,
    fill = NA,
    color = "#67001f",
    size = 0.9,
    stroke = 0.35
  ) +
  scale_fill_gradient2(
    low = "#2166ac",
    mid = "white",
    high = "#b2182b",
    midpoint = 0,
    limits = c(-st22_spatial_color_limit, st22_spatial_color_limit),
    oob = scales::squish
  ) +
  scale_x_continuous(breaks = seq(0, 50, by = 5), expand = c(0, 0)) +
  labs(
    title = "ST22 sediment change through time and along the transect",
    subtitle = "Baseline-Con1-Con2 reference; outlined cells meet or exceed 1 cm accumulation",
    x = "Meter position",
    y = "Construction survey",
    fill = "Sediment delta (cm)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid = element_blank(),
    plot.title.position = "plot"
  )

st22_meter_position_heatmap
```

### Empirical tail-balance analysis

The following views use empirical exceedance fractions rather than smoothed
density estimates. At magnitude \(t\), accumulation exceedance is the fraction
of meter positions with \(\Delta \ge t\), loss exceedance is the fraction with
\(\Delta \le -t\), and tail balance is their difference.

```{r st22-empirical-tail-profiles}
st22_reference_deltas_long <- st22_alternative_deltas %>%
  select(
    date,
    survey,
    survey_n,
    meter,
    official_delta_cm,
    alternative_delta_cm
  ) %>%
  pivot_longer(
    c(official_delta_cm, alternative_delta_cm),
    names_to = "reference",
    values_to = "sediment_delta_cm"
  ) %>%
  mutate(
    reference = recode(
      reference,
      official_delta_cm = "Official baseline",
      alternative_delta_cm = "Baseline-Con1-Con2 average"
    ),
    reference = factor(
      reference,
      levels = c("Official baseline", "Baseline-Con1-Con2 average")
    ),
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n)
  )

tail_profile_max_cm <- ceiling(
  max(abs(st22_reference_deltas_long$sediment_delta_cm))
)

tail_profile_thresholds <- tibble(
  threshold_cm = seq(0, tail_profile_max_cm, by = 0.05)
)

st22_empirical_tail_profiles <- crossing(
  st22_reference_deltas_long,
  tail_profile_thresholds
) %>%
  summarize(
    accumulation_fraction = if_else(
      first(threshold_cm) == 0,
      mean(sediment_delta_cm > 0),
      mean(sediment_delta_cm >= threshold_cm)
    ),
    loss_fraction = if_else(
      first(threshold_cm) == 0,
      mean(sediment_delta_cm < 0),
      mean(sediment_delta_cm <= -threshold_cm)
    ),
    tail_balance = accumulation_fraction - loss_fraction,
    .by = c(
      date,
      survey,
      survey_n,
      survey_label,
      reference,
      threshold_cm
    )
  ) %>%
  arrange(reference, survey_n, threshold_cm)

st22_alternative_exceedance_profiles <- st22_empirical_tail_profiles %>%
  filter(reference == "Baseline-Con1-Con2 average")

tail_plot_max_cm <- ceiling(
  2 * max(abs(st22_alternative_deltas$alternative_delta_cm))
) / 2

tail_plot_breaks_cm <- sort(unique(c(
  seq(0, floor(tail_plot_max_cm), by = 1),
  tail_plot_max_cm
)))

st22_net_exceedance_profiles <- st22_alternative_exceedance_profiles %>%
  filter(threshold_cm <= tail_plot_max_cm)

st22_weekly_mean_delta <- st22_alternative_deltas %>%
  summarize(
    mean_delta_cm = mean(alternative_delta_cm),
    n_meter_positions = n(),
    .by = c(date, survey, survey_n)
  ) %>%
  mutate(
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n),
    mean_absolute_delta_cm = abs(mean_delta_cm),
    mean_direction = if_else(
      mean_delta_cm >= 0,
      "Positive transect mean",
      "Negative transect mean"
    )
  )
```

#### Mirrored empirical curves with net balance

The red and blue curves remain the actual empirical accumulation and loss
prevalences. The black curve overlays their difference: accumulation prevalence
minus equivalent-loss prevalence. The shared x-axis extends through the full
observed alternative-reference change range, rounded up to the next 0.5 cm.

```{r st22-mirrored-empirical-with-net, fig.width=11, fig.height=15}
st22_empirical_mirrored_profiles <-
  st22_alternative_exceedance_profiles %>%
  filter(threshold_cm <= tail_plot_max_cm) %>%
  select(
    date,
    survey,
    survey_n,
    survey_label,
    threshold_cm,
    accumulation_fraction,
    loss_fraction
  ) %>%
  pivot_longer(
    c(accumulation_fraction, loss_fraction),
    names_to = "tail",
    values_to = "fraction"
  ) %>%
  mutate(
    tail = recode(
      tail,
      accumulation_fraction = "Accumulation",
      loss_fraction = "Equivalent loss"
    ),
    tail = factor(tail, levels = c("Accumulation", "Equivalent loss")),
    signed_fraction = if_else(tail == "Accumulation", fraction, -fraction),
    threshold_end = lead(threshold_cm, default = tail_plot_max_cm),
    .by = c(survey_n, tail)
  )

st22_mirrored_empirical_net_plot <- ggplot() +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_rect(
    data = st22_empirical_mirrored_profiles,
    aes(
      xmin = threshold_cm,
      xmax = threshold_end,
      ymin = pmin(signed_fraction, 0),
      ymax = pmax(signed_fraction, 0),
      fill = tail
    ),
    alpha = 0.4,
    color = NA
  ) +
  geom_step(
    data = st22_empirical_mirrored_profiles,
    aes(
      x = threshold_cm,
      y = signed_fraction,
      color = tail,
      group = tail
    ),
    linewidth = 0.65,
    direction = "hv"
  ) +
  geom_step(
    data = st22_net_exceedance_profiles,
    aes(x = threshold_cm, y = tail_balance),
    color = "white",
    linewidth = 1.4,
    direction = "hv"
  ) +
  geom_step(
    data = st22_net_exceedance_profiles,
    aes(x = threshold_cm, y = tail_balance, color = "Net balance"),
    linewidth = 0.7,
    direction = "hv"
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  geom_vline(
    data = st22_weekly_mean_delta,
    aes(xintercept = mean_absolute_delta_cm, color = mean_direction),
    linewidth = 0.75,
    linetype = "longdash"
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_x_continuous(
    limits = c(0, tail_plot_max_cm),
    breaks = tail_plot_breaks_cm,
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = ~ scales::percent(abs(.x), accuracy = 1),
    breaks = seq(-0.75, 0.75, by = 0.25),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  coord_cartesian(ylim = c(-0.75, 0.75)) +
  scale_color_manual(
    values = c(
      "Accumulation" = "#b2182b",
      "Equivalent loss" = "#2166ac",
      "Net balance" = "#1a1a1a",
      "Positive transect mean" = "#7f0000",
      "Negative transect mean" = "#084594"
    )
  ) +
  scale_fill_manual(
    values = c("Accumulation" = "#d6604d", "Equivalent loss" = "#4393c3")
  ) +
  labs(
    title = "ST22 empirical cumulative tails with net balance",
    subtitle = "Black = accumulation prevalence minus equivalent-loss prevalence",
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Meter positions at threshold or greater",
    color = NULL,
    fill = NULL
  ) +
  guides(fill = "none") +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_mirrored_empirical_net_plot
```

#### Animated empirical curves with maximum-net memory

The red, blue, and black curves show the current survey. At each absolute
sediment-depth threshold, the dashed green curve records the largest signed net
balance observed from Con3 through the current survey. It therefore acts as a
threshold-specific memory of the strongest accumulation-dominant result seen
so far.

```{r st22-animated-empirical-data}
st22_animation_frames <- st22_net_exceedance_profiles %>%
  distinct(survey_n, survey_label) %>%
  left_join(
    st22_weekly_mean_delta %>%
      select(
        survey_n,
        mean_delta_cm,
        mean_absolute_delta_cm,
        mean_direction
      ),
    by = "survey_n"
  ) %>%
  arrange(survey_n) %>%
  mutate(
    frame_label = paste0(
      str_replace(survey_label, "\\n", "  |  ")
    ),
    frame_label = factor(frame_label, levels = frame_label)
  )

st22_animation_tails <- st22_empirical_mirrored_profiles %>%
  left_join(st22_animation_frames, by = c("survey_n", "survey_label"))

st22_animation_net <- st22_net_exceedance_profiles %>%
  left_join(st22_animation_frames, by = c("survey_n", "survey_label"))

st22_animation_max_net <- st22_animation_net %>%
  arrange(threshold_cm, survey_n) %>%
  mutate(
    maximum_net_to_date = cummax(tail_balance),
    .by = threshold_cm
  ) %>%
  arrange(survey_n, threshold_cm)
```

```{r st22-static-animation-frames, fig.width=10, fig.height=7}
make_st22_empirical_frame <- function(frame_survey_n) {
  frame_info <- st22_animation_frames %>%
    filter(survey_n == frame_survey_n)
  frame_tails <- st22_animation_tails %>%
    filter(survey_n == frame_survey_n)
  frame_net <- st22_animation_net %>%
    filter(survey_n == frame_survey_n)
  frame_max_net <- st22_animation_max_net %>%
    filter(survey_n == frame_survey_n)

  ggplot() +
    geom_hline(yintercept = 0, color = "grey35", linewidth = 0.45) +
    geom_area(
      data = frame_tails,
      aes(
        x = threshold_cm,
        y = signed_fraction,
        fill = tail,
        group = tail
      ),
      alpha = 0.4,
      position = "identity"
    ) +
    geom_step(
      data = frame_tails,
      aes(
        x = threshold_cm,
        y = signed_fraction,
        color = tail,
        group = tail
      ),
      linewidth = 0.9,
      direction = "hv"
    ) +
    geom_step(
      data = frame_net,
      aes(x = threshold_cm, y = tail_balance),
      color = "white",
      linewidth = 2,
      direction = "hv"
    ) +
    geom_step(
      data = frame_net,
      aes(x = threshold_cm, y = tail_balance, color = "Current net balance"),
      linewidth = 1,
      direction = "hv"
    ) +
    geom_step(
      data = frame_max_net,
      aes(x = threshold_cm, y = maximum_net_to_date),
      color = "white",
      linewidth = 2.2,
      direction = "hv"
    ) +
    geom_step(
      data = frame_max_net,
      aes(
        x = threshold_cm,
        y = maximum_net_to_date,
        color = "Maximum net to date"
      ),
      linewidth = 1.1,
      linetype = "longdash",
      direction = "hv"
    ) +
    geom_vline(
      xintercept = c(0.15, 0.5, 1),
      color = "grey45",
      linewidth = 0.4,
      linetype = "dotted"
    ) +
    geom_vline(
      data = frame_info,
      aes(xintercept = mean_absolute_delta_cm, color = mean_direction),
      linewidth = 0.9,
      linetype = "longdash"
    ) +
    scale_x_continuous(
      limits = c(0, tail_plot_max_cm),
      breaks = tail_plot_breaks_cm,
      expand = expansion(mult = c(0, 0.01))
    ) +
    scale_y_continuous(
      labels = ~ scales::percent(abs(.x), accuracy = 1),
      breaks = seq(-0.75, 0.75, by = 0.25),
      expand = expansion(mult = c(0.01, 0.01))
    ) +
    coord_cartesian(ylim = c(-0.75, 0.75)) +
    scale_color_manual(
      values = c(
        "Accumulation" = "#b2182b",
        "Equivalent loss" = "#2166ac",
        "Current net balance" = "#1a1a1a",
        "Maximum net to date" = "#1b7837",
        "Positive transect mean" = "#7f0000",
        "Negative transect mean" = "#084594"
      )
    ) +
    scale_fill_manual(
      values = c("Accumulation" = "#d6604d", "Equivalent loss" = "#4393c3")
    ) +
    labs(
      title = paste0("ST22 empirical cumulative tails: ", frame_info$frame_label),
      subtitle = "Dashed green line remembers the maximum net balance reached at each threshold",
      x = "Absolute sediment-depth change threshold (cm)",
      y = "Meter positions at threshold or greater",
      color = NULL,
      fill = NULL
    ) +
    guides(
      fill = "none",
      color = guide_legend(
        override.aes = list(
          linetype = c(
            "solid",
            "solid",
            "solid",
            "longdash",
            "longdash"
          ),
          linewidth = c(1, 1, 1, 1.1, 0.9)
        )
      )
    ) +
    theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      panel.border = element_rect(
        color = "grey75",
        fill = NA,
        linewidth = 0.4
      ),
      legend.position = "bottom",
      plot.title = element_text(face = "bold"),
      plot.title.position = "plot"
    )
}

st22_frame_dir <- file.path("code", "figures", "st22_animation_frames")
dir.create(st22_frame_dir, recursive = TRUE, showWarnings = FALSE)

st22_frame_files <- st22_animation_frames %>%
  transmute(
    survey_n,
    frame_file = file.path(
      st22_frame_dir,
      sprintf("st22_frame_%02d.png", survey_n)
    )
  )

pwalk(
  st22_frame_files,
  function(survey_n, frame_file) {
    ggsave(
      frame_file,
      plot = make_st22_empirical_frame(survey_n),
      width = 10,
      height = 7,
      units = "in",
      dpi = 96,
      bg = "white"
    )
  }
)

stopifnot(all(file.exists(st22_frame_files$frame_file)))

st22_final_survey_n <- max(st22_animation_frames$survey_n)
st22_final_net <- st22_animation_net %>%
  filter(survey_n == st22_final_survey_n)
st22_final_max_net <- st22_animation_max_net %>%
  filter(survey_n == st22_final_survey_n)
st22_final_frame_label <- st22_animation_frames %>%
  filter(survey_n == st22_final_survey_n) %>%
  pull(frame_label) %>%
  as.character()
st22_final_mean_delta <- st22_animation_frames %>%
  filter(survey_n == st22_final_survey_n) %>%
  select(mean_delta_cm, mean_absolute_delta_cm, mean_direction)

st22_final_summary_plot <- ggplot() +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.45) +
  geom_step(
    data = st22_final_net,
    aes(x = threshold_cm, y = tail_balance, color = "Most recent net balance"),
    linewidth = 1.15,
    direction = "hv"
  ) +
  geom_step(
    data = st22_final_max_net,
    aes(
      x = threshold_cm,
      y = maximum_net_to_date,
      color = "Maximum net over monitoring period"
    ),
    linewidth = 1.2,
    linetype = "longdash",
    direction = "hv"
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.4,
    linetype = "dotted"
  ) +
  geom_vline(
    data = st22_final_mean_delta,
    aes(xintercept = mean_absolute_delta_cm, color = mean_direction),
    linewidth = 0.9,
    linetype = "longdash"
  ) +
  scale_x_continuous(
    limits = c(0, tail_plot_max_cm),
    breaks = tail_plot_breaks_cm,
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    breaks = seq(-0.75, 0.75, by = 0.25),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  coord_cartesian(ylim = c(-0.75, 0.75)) +
  scale_color_manual(
    values = c(
      "Most recent net balance" = "#1a1a1a",
      "Maximum net over monitoring period" = "#1b7837",
      "Positive transect mean" = "#7f0000",
      "Negative transect mean" = "#084594"
    )
  ) +
  labs(
    title = paste0("ST22 net-balance summary: ", st22_final_frame_label),
    subtitle = "Black is the most recent net balance; dashed green is the maximum observed at each threshold",
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Accumulation minus equivalent-loss prevalence",
    color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey75", fill = NA, linewidth = 0.4),
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot"
  )

st22_final_summary_file <- file.path(
  st22_frame_dir,
  "st22_frame_final_summary.png"
)
ggsave(
  st22_final_summary_file,
  plot = st22_final_summary_plot,
  width = 10,
  height = 7,
  units = "in",
  dpi = 96,
  bg = "white"
)
```

```{r st22-interactive-empirical-player, echo=FALSE, results='asis'}
st22_player_id <- "st22-empirical-memory-player"
st22_player_frames <- c(
  map_chr(st22_frame_files$frame_file, knitr::image_uri),
  knitr::image_uri(st22_final_summary_file)
)
st22_player_frames_json <- jsonlite::toJSON(
  st22_player_frames,
  auto_unbox = TRUE
)

st22_player <- htmltools::tags$div(
  id = st22_player_id,
  class = "st22-player",
  htmltools::tags$style(htmltools::HTML(paste(
    ".st22-player { max-width: 960px; margin: 1rem auto; }",
    ".st22-player img { width: 100%; height: auto; display: block; cursor: pointer; border: 1px solid #d4d4d4; }",
    ".st22-player-controls { display: flex; align-items: center; gap: 0.65rem; margin-top: 0.55rem; }",
    ".st22-player button { border: 1px solid #777; background: #fff; color: #222; padding: 0.35rem 0.75rem; cursor: pointer; }",
    ".st22-player button:hover { background: #f0f0f0; }",
    ".st22-player input[type=range] { flex: 1; }",
    ".st22-player-status { min-width: 5.5rem; text-align: right; font-variant-numeric: tabular-nums; }",
    sep = "\n"
  ))),
  htmltools::tags$img(
    src = st22_player_frames[[1]],
    alt = "Animated ST22 empirical accumulation, loss, and net-balance curves"
  ),
  htmltools::tags$div(
    class = "st22-player-controls",
    htmltools::tags$button(type = "button", `data-action` = "play", "Play"),
    htmltools::tags$button(type = "button", `data-action` = "restart", "Restart"),
    htmltools::tags$input(
      type = "range",
      min = 0,
      max = length(st22_player_frames) - 1,
      value = 0,
      step = 1,
      `aria-label` = "Animation frame"
    ),
    htmltools::tags$span(
      class = "st22-player-status",
      paste0("1 / ", length(st22_player_frames))
    )
  ),
  htmltools::tags$script(htmltools::HTML(sprintf(paste0(
    "(function() {",
    "const root=document.getElementById('%s');",
    "const frames=%s;",
    "const image=root.querySelector('img');",
    "const playButton=root.querySelector('[data-action=play]');",
    "const restartButton=root.querySelector('[data-action=restart]');",
    "const slider=root.querySelector('input[type=range]');",
    "const status=root.querySelector('.st22-player-status');",
    "let index=0; let timer=null;",
    "function showFrame(nextIndex){",
    "index=Math.max(0,Math.min(nextIndex,frames.length-1));",
    "image.src=frames[index]; slider.value=index;",
    "status.textContent=(index+1)+' / '+frames.length;}",
    "function stop(){if(timer!==null)window.clearInterval(timer);",
    "timer=null; playButton.textContent=index===frames.length-1?'Play again':'Play';}",
    "function play(){if(timer!==null){stop();return;}",
    "if(index===frames.length-1)showFrame(0); playButton.textContent='Pause';",
    "timer=window.setInterval(function(){",
    "if(index>=frames.length-1){stop();return;} showFrame(index+1);",
    "if(index===frames.length-1)stop();},500);}",
    "playButton.addEventListener('click',play);",
    "image.addEventListener('click',play);",
    "restartButton.addEventListener('click',function(){stop();showFrame(0);});",
    "slider.addEventListener('input',function(){stop();showFrame(Number(this.value));});",
    "})();"
  ), st22_player_id, st22_player_frames_json)))
)

st22_player_html <- as.character(st22_player)
st22_player_html <- gsub(">[[:space:]]+<", "><", st22_player_html)
cat(st22_player_html)
```

#### Stacked net contributions by sediment-change band

This view decomposes cumulative net exceedance into three mutually exclusive
magnitude bands. The `1 cm+` segment is the net balance at 1 cm or greater; the
two lighter segments are the additional net contributions between successive
thresholds. The signed segment values sum to the cumulative net balance at
1.5 mm or greater.

```{r st22-stacked-net-threshold-contributions, fig.width=12, fig.height=8}
st22_cumulative_net_thresholds <-
  st22_alternative_exceedance_profiles %>%
  mutate(threshold_key_cm = round(threshold_cm, 2)) %>%
  filter(threshold_key_cm %in% c(0.15, 0.5, 1)) %>%
  mutate(
    survey_timepoint = factor(
      survey,
      levels = survey[order(survey_n)] %>% unique()
    )
  )

st22_stacked_net_wide <- st22_cumulative_net_thresholds %>%
  mutate(
    threshold_code = recode(
      as.character(threshold_key_cm),
      `0.15` = "net_1_5mm",
      `0.5` = "net_5mm",
      `1` = "net_1cm"
    )
  ) %>%
  select(
    date,
    survey,
    survey_n,
    survey_timepoint,
    threshold_code,
    tail_balance
  ) %>%
  pivot_wider(names_from = threshold_code, values_from = tail_balance) %>%
  mutate(
    `1 cm+` = net_1cm,
    `5 mm to <1 cm` = net_5mm - net_1cm,
    `1.5 mm to <5 mm` = net_1_5mm - net_5mm,
    total_net_1_5mm = net_1_5mm
  )

st22_stacked_net_contributions <- st22_stacked_net_wide %>%
  pivot_longer(
    c(`1 cm+`, `5 mm to <1 cm`, `1.5 mm to <5 mm`),
    names_to = "magnitude_band",
    values_to = "net_contribution"
  ) %>%
  mutate(
    magnitude_band = factor(
      magnitude_band,
      levels = c("1 cm+", "5 mm to <1 cm", "1.5 mm to <5 mm")
    ),
    contribution_direction = factor(
      if_else(net_contribution >= 0, "Accumulation", "Loss"),
      levels = c("Accumulation", "Loss")
    )
  )

st22_stacked_identity_check <- st22_stacked_net_contributions %>%
  summarize(
    reconstructed_total = sum(net_contribution),
    expected_total = first(total_net_1_5mm),
    .by = survey_n
  )

stopifnot(all(
  near(
    st22_stacked_identity_check$reconstructed_total,
    st22_stacked_identity_check$expected_total
  )
))

st22_stacked_net_threshold_plot <- ggplot(
  st22_stacked_net_contributions,
  aes(
    x = survey_timepoint,
    y = net_contribution,
    fill = contribution_direction,
    alpha = magnitude_band,
    group = magnitude_band
  )
) +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_col(
    width = 0.75,
    position = position_stack(reverse = TRUE)
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(-0.75, 0.75),
    breaks = seq(-0.75, 0.75, by = 0.25),
    expand = expansion(mult = c(0.03, 0.03))
  ) +
  scale_fill_manual(
    values = c("Accumulation" = "#b2182b", "Loss" = "#2166ac")
  ) +
  scale_alpha_manual(
    values = c(
      "1 cm+" = 1,
      "5 mm to <1 cm" = 0.65,
      "1.5 mm to <5 mm" = 0.35
    )
  ) +
  labs(
    title = "ST22 net exceedance decomposed by sediment-change band",
    subtitle = "Stacked segments are signed, mutually exclusive contributions; points below show the transect mean",
    x = NULL,
    y = "Net prevalence contribution",
    fill = "Direction",
    alpha = "Magnitude band"
  ) +
  guides(
    fill = guide_legend(order = 1),
    alpha = guide_legend(order = 2)
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_stacked_mean_plot <- st22_weekly_mean_delta %>%
  mutate(
    survey_timepoint = factor(
      survey,
      levels = levels(st22_stacked_net_contributions$survey_timepoint)
    )
  ) %>%
  ggplot(aes(x = survey_timepoint, y = mean_delta_cm)) +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_point(aes(color = mean_direction), size = 2.3) +
  scale_color_manual(
    values = c(
      "Positive transect mean" = "#7f0000",
      "Negative transect mean" = "#084594"
    ),
    guide = "none"
  ) +
  labs(
    x = "Construction survey",
    y = "Transect mean\ndelta (cm)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1)
  )

st22_stacked_net_threshold_plot / st22_stacked_mean_plot +
  plot_layout(heights = c(4, 1), guides = "collect") &
  theme(legend.position = "bottom")
```


## Case study 2: ST28

### Sediment deltas

For each survey, the three sediment-depth replicates are averaged at each meter
position. The baseline mean at that same meter position is then subtracted from
the construction-survey mean:

\[
\Delta_{m,t} = \overline{x}_{m,t} - \overline{x}_{m,baseline}
\]

Positive values indicate sediment accumulation relative to baseline. This
meter-matched calculation retains spatial variation that would be lost by first
averaging across the full transect.

```{r st28-meter-deltas}
st28_raw <- read_excel(
  latest_construction_file,
  sheet = "SedimentDepth_RawData"
) %>%
  clean_names() %>%
  mutate(
    date = as.Date(date),
    survey = str_squish(survey),
    survey_n = if_else(
      survey == "Baseline",
      0L,
      as.integer(str_extract(survey, "\\d+"))
    ),
    transect = normalize_transect_id(transect),
    meter = as.integer(meter),
    meter_mean_sediment_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    )
  ) %>%
  filter(transect == "ST28")

st28_baseline <- st28_raw %>%
  filter(survey == "Baseline") %>%
  transmute(
    meter,
    baseline_mean_sediment_cm = meter_mean_sediment_cm
  )

st28_meter_deltas <- st28_raw %>%
  filter(survey != "Baseline") %>%
  left_join(st28_baseline, by = "meter") %>%
  mutate(
    sediment_delta_cm =
      meter_mean_sediment_cm - baseline_mean_sediment_cm
  ) %>%
  select(
    date,
    survey,
    survey_n,
    transect,
    meter,
    baseline_mean_sediment_cm,
    meter_mean_sediment_cm,
    sediment_delta_cm
  ) %>%
  arrange(survey_n, meter)

st28_delta_coverage <- st28_meter_deltas %>%
  summarize(
    n_meter_positions = n(),
    n_missing_deltas = sum(is.na(sediment_delta_cm)),
    .by = c(date, survey, survey_n)
  ) %>%
  arrange(survey_n)

stopifnot(
  all(st28_delta_coverage$n_meter_positions == 51),
  all(st28_delta_coverage$n_missing_deltas == 0)
)

st28_delta_coverage
```

### Alternative Baseline-Con1-Con2 reference

This sensitivity analysis defines a meter-specific alternative reference as
the arithmetic mean of the Baseline, Con1, and Con2 meter means. Each survey
receives equal weight. Because Con1 and Con2 contribute to the reference,
alternative-reference results begin with Con3. This does not replace the BMP
baseline analysis above.

```{r st28-alternative-reference}
st28_alternative_reference <- st28_raw %>%
  filter(survey_n %in% 0:2) %>%
  summarize(
    alternative_reference_cm = mean(meter_mean_sediment_cm),
    n_reference_surveys = n(),
    .by = meter
  ) %>%
  left_join(st28_baseline, by = "meter") %>%
  mutate(
    reference_difference_cm =
      alternative_reference_cm - baseline_mean_sediment_cm
  )

stopifnot(
  nrow(st28_alternative_reference) == 51,
  all(st28_alternative_reference$n_reference_surveys == 3)
)

st28_reference_summary <- st28_alternative_reference %>%
  summarize(
    n_meter_positions = n(),
    official_baseline_mean_cm = mean(baseline_mean_sediment_cm),
    alternative_reference_mean_cm = mean(alternative_reference_cm),
    mean_reference_difference_cm = mean(reference_difference_cm),
    median_reference_difference_cm = median(reference_difference_cm),
    pct_alternative_lower = mean(reference_difference_cm < 0),
    pct_alternative_higher = mean(reference_difference_cm > 0)
  )

st28_reference_summary
```

```{r st28-alternative-deltas}
st28_alternative_deltas <- st28_raw %>%
  filter(survey_n >= 3) %>%
  left_join(
    st28_alternative_reference %>%
      select(meter, alternative_reference_cm),
    by = "meter"
  ) %>%
  left_join(st28_baseline, by = "meter") %>%
  mutate(
    alternative_delta_cm =
      meter_mean_sediment_cm - alternative_reference_cm,
    official_delta_cm =
      meter_mean_sediment_cm - baseline_mean_sediment_cm
  ) %>%
  select(
    date,
    survey,
    survey_n,
    transect,
    meter,
    meter_mean_sediment_cm,
    baseline_mean_sediment_cm,
    alternative_reference_cm,
    official_delta_cm,
    alternative_delta_cm
  ) %>%
  arrange(survey_n, meter)
```

### Meter-position diagnostic

This heatmap provides a relatively raw view of sediment change before the
transect is summarized with exceedance curves. It shows whether accumulation
or loss is scattered among meter positions or forms spatially coherent patterns
that persist through multiple surveys. Results use the Baseline-Con1-Con2
alternative reference; outlined cells meet or exceed 1 cm accumulation.

```{r st28-meter-position-heatmap, fig.width=11, fig.height=7.5}
st28_spatial_deltas <- st28_alternative_deltas %>%
  mutate(
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n, .desc = TRUE)
  )

st28_spatial_color_limit <- max(
  abs(
    quantile(
      st28_spatial_deltas$alternative_delta_cm,
      probs = c(0.02, 0.98),
      na.rm = TRUE
    )
  )
)

st28_meter_position_heatmap <- ggplot(
  st28_spatial_deltas,
  aes(x = meter, y = survey_label, fill = alternative_delta_cm)
) +
  geom_tile() +
  geom_point(
    data = st28_spatial_deltas %>%
      filter(alternative_delta_cm >= 1),
    shape = 21,
    fill = NA,
    color = "#67001f",
    size = 0.9,
    stroke = 0.35
  ) +
  scale_fill_gradient2(
    low = "#2166ac",
    mid = "white",
    high = "#b2182b",
    midpoint = 0,
    limits = c(-st28_spatial_color_limit, st28_spatial_color_limit),
    oob = scales::squish
  ) +
  scale_x_continuous(breaks = seq(0, 50, by = 5), expand = c(0, 0)) +
  labs(
    title = "ST28 sediment change through time and along the transect",
    subtitle = "Baseline-Con1-Con2 reference; outlined cells meet or exceed 1 cm accumulation",
    x = "Meter position",
    y = "Construction survey",
    fill = "Sediment delta (cm)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid = element_blank(),
    plot.title.position = "plot"
  )

st28_meter_position_heatmap
```

### Empirical tail-balance analysis

The following views use empirical exceedance fractions rather than smoothed
density estimates. At magnitude \(t\), accumulation exceedance is the fraction
of meter positions with \(\Delta \ge t\), loss exceedance is the fraction with
\(\Delta \le -t\), and tail balance is their difference.

```{r st28-empirical-tail-profiles}
st28_reference_deltas_long <- st28_alternative_deltas %>%
  select(
    date,
    survey,
    survey_n,
    meter,
    official_delta_cm,
    alternative_delta_cm
  ) %>%
  pivot_longer(
    c(official_delta_cm, alternative_delta_cm),
    names_to = "reference",
    values_to = "sediment_delta_cm"
  ) %>%
  mutate(
    reference = recode(
      reference,
      official_delta_cm = "Official baseline",
      alternative_delta_cm = "Baseline-Con1-Con2 average"
    ),
    reference = factor(
      reference,
      levels = c("Official baseline", "Baseline-Con1-Con2 average")
    ),
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n)
  )

tail_profile_max_cm <- ceiling(
  max(abs(st28_reference_deltas_long$sediment_delta_cm))
)

tail_profile_thresholds <- tibble(
  threshold_cm = seq(0, tail_profile_max_cm, by = 0.05)
)

st28_empirical_tail_profiles <- crossing(
  st28_reference_deltas_long,
  tail_profile_thresholds
) %>%
  summarize(
    accumulation_fraction = if_else(
      first(threshold_cm) == 0,
      mean(sediment_delta_cm > 0),
      mean(sediment_delta_cm >= threshold_cm)
    ),
    loss_fraction = if_else(
      first(threshold_cm) == 0,
      mean(sediment_delta_cm < 0),
      mean(sediment_delta_cm <= -threshold_cm)
    ),
    tail_balance = accumulation_fraction - loss_fraction,
    .by = c(
      date,
      survey,
      survey_n,
      survey_label,
      reference,
      threshold_cm
    )
  ) %>%
  arrange(reference, survey_n, threshold_cm)

st28_alternative_exceedance_profiles <- st28_empirical_tail_profiles %>%
  filter(reference == "Baseline-Con1-Con2 average")

st28_observed_tail_max_cm <- max(
  abs(st28_alternative_deltas$alternative_delta_cm)
)

st28_tail_plot_max_cm <- ceiling(
  2 * st28_observed_tail_max_cm
) / 2

st28_threshold_axis_trans <- scales::pseudo_log_trans(
  sigma = 0.15,
  base = 10
)

st28_tail_plot_breaks_cm <- sort(unique(c(
  c(0, 0.15, 0.5, 1, 5, 10, 15)[
    c(0, 0.15, 0.5, 1, 5, 10, 15) <= st28_tail_plot_max_cm
  ],
  st28_tail_plot_max_cm
)))

st28_net_exceedance_profiles <- st28_alternative_exceedance_profiles %>%
  filter(threshold_cm <= st28_tail_plot_max_cm)

st28_weekly_mean_delta <- st28_alternative_deltas %>%
  summarize(
    mean_delta_cm = mean(alternative_delta_cm),
    n_meter_positions = n(),
    .by = c(date, survey, survey_n)
  ) %>%
  mutate(
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n),
    mean_absolute_delta_cm = abs(mean_delta_cm),
    mean_direction = if_else(
      mean_delta_cm >= 0,
      "Positive transect mean",
      "Negative transect mean"
    )
  )
```

#### Mirrored empirical curves with net balance

The red and blue curves remain the actual empirical accumulation and loss
prevalences. The black curve overlays their difference: accumulation prevalence
minus equivalent-loss prevalence. The shared x-axis extends through the full
observed alternative-reference change range, rounded up to the next 0.5 cm.

```{r st28-mirrored-empirical-with-net, fig.width=11, fig.height=15}
st28_empirical_mirrored_profiles <-
  st28_alternative_exceedance_profiles %>%
  filter(threshold_cm <= st28_tail_plot_max_cm) %>%
  select(
    date,
    survey,
    survey_n,
    survey_label,
    threshold_cm,
    accumulation_fraction,
    loss_fraction
  ) %>%
  pivot_longer(
    c(accumulation_fraction, loss_fraction),
    names_to = "tail",
    values_to = "fraction"
  ) %>%
  mutate(
    tail = recode(
      tail,
      accumulation_fraction = "Accumulation",
      loss_fraction = "Equivalent loss"
    ),
    tail = factor(tail, levels = c("Accumulation", "Equivalent loss")),
    signed_fraction = if_else(tail == "Accumulation", fraction, -fraction),
    threshold_end = lead(threshold_cm, default = st28_tail_plot_max_cm),
    .by = c(survey_n, tail)
  )

st28_mirrored_empirical_net_plot <- ggplot() +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_rect(
    data = st28_empirical_mirrored_profiles,
    aes(
      xmin = threshold_cm,
      xmax = threshold_end,
      ymin = pmin(signed_fraction, 0),
      ymax = pmax(signed_fraction, 0),
      fill = tail
    ),
    alpha = 0.4,
    color = NA
  ) +
  geom_step(
    data = st28_empirical_mirrored_profiles,
    aes(
      x = threshold_cm,
      y = signed_fraction,
      color = tail,
      group = tail
    ),
    linewidth = 0.65,
    direction = "hv"
  ) +
  geom_step(
    data = st28_net_exceedance_profiles,
    aes(x = threshold_cm, y = tail_balance),
    color = "white",
    linewidth = 1.4,
    direction = "hv"
  ) +
  geom_step(
    data = st28_net_exceedance_profiles,
    aes(x = threshold_cm, y = tail_balance, color = "Net balance"),
    linewidth = 0.7,
    direction = "hv"
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  geom_vline(
    data = st28_weekly_mean_delta,
    aes(xintercept = mean_absolute_delta_cm, color = mean_direction),
    linewidth = 0.75,
    linetype = "longdash"
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_x_continuous(
    trans = st28_threshold_axis_trans,
    limits = c(0, st28_tail_plot_max_cm),
    breaks = st28_tail_plot_breaks_cm,
    labels = ~ str_remove(
      format(.x, trim = TRUE, scientific = FALSE),
      "\\.?0+$"
    ),
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = ~ scales::percent(abs(.x), accuracy = 1),
    breaks = seq(-0.75, 0.75, by = 0.25),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  coord_cartesian(ylim = c(-0.75, 0.75)) +
  scale_color_manual(
    values = c(
      "Accumulation" = "#b2182b",
      "Equivalent loss" = "#2166ac",
      "Net balance" = "#1a1a1a",
      "Positive transect mean" = "#7f0000",
      "Negative transect mean" = "#084594"
    )
  ) +
  scale_fill_manual(
    values = c("Accumulation" = "#d6604d", "Equivalent loss" = "#4393c3")
  ) +
  labs(
    title = "ST28 empirical cumulative tails with net balance",
    subtitle = "Black = accumulation prevalence minus equivalent-loss prevalence",
    x = "Absolute sediment-depth change threshold (cm; pseudo-log scale)",
    y = "Meter positions at threshold or greater",
    color = NULL,
    fill = NULL
  ) +
  guides(fill = "none") +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    panel.spacing.x = unit(1, "lines"),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st28_mirrored_empirical_net_plot
```

#### Animated empirical curves with maximum-net memory

The red, blue, and black curves show the current survey. At each absolute
sediment-depth threshold, the dashed green curve records the largest signed net
balance observed from Con3 through the current survey. It therefore acts as a
threshold-specific memory of the strongest accumulation-dominant result seen
so far.

```{r st28-animated-empirical-data}
st28_animation_frames <- st28_net_exceedance_profiles %>%
  distinct(survey_n, survey_label) %>%
  left_join(
    st28_weekly_mean_delta %>%
      select(
        survey_n,
        mean_delta_cm,
        mean_absolute_delta_cm,
        mean_direction
      ),
    by = "survey_n"
  ) %>%
  arrange(survey_n) %>%
  mutate(
    frame_label = paste0(
      str_replace(survey_label, "\\n", "  |  ")
    ),
    frame_label = factor(frame_label, levels = frame_label)
  )

st28_animation_tails <- st28_empirical_mirrored_profiles %>%
  left_join(st28_animation_frames, by = c("survey_n", "survey_label"))

st28_animation_net <- st28_net_exceedance_profiles %>%
  left_join(st28_animation_frames, by = c("survey_n", "survey_label"))

st28_animation_max_net <- st28_animation_net %>%
  arrange(threshold_cm, survey_n) %>%
  mutate(
    maximum_net_to_date = cummax(tail_balance),
    .by = threshold_cm
  ) %>%
  arrange(survey_n, threshold_cm)
```

```{r st28-static-animation-frames, fig.width=10, fig.height=7}
make_st28_empirical_frame <- function(frame_survey_n) {
  frame_info <- st28_animation_frames %>%
    filter(survey_n == frame_survey_n)
  frame_tails <- st28_animation_tails %>%
    filter(survey_n == frame_survey_n)
  frame_net <- st28_animation_net %>%
    filter(survey_n == frame_survey_n)
  frame_max_net <- st28_animation_max_net %>%
    filter(survey_n == frame_survey_n)

  ggplot() +
    geom_hline(yintercept = 0, color = "grey35", linewidth = 0.45) +
    geom_area(
      data = frame_tails,
      aes(
        x = threshold_cm,
        y = signed_fraction,
        fill = tail,
        group = tail
      ),
      alpha = 0.4,
      position = "identity"
    ) +
    geom_step(
      data = frame_tails,
      aes(
        x = threshold_cm,
        y = signed_fraction,
        color = tail,
        group = tail
      ),
      linewidth = 0.9,
      direction = "hv"
    ) +
    geom_step(
      data = frame_net,
      aes(x = threshold_cm, y = tail_balance),
      color = "white",
      linewidth = 2,
      direction = "hv"
    ) +
    geom_step(
      data = frame_net,
      aes(x = threshold_cm, y = tail_balance, color = "Current net balance"),
      linewidth = 1,
      direction = "hv"
    ) +
    geom_step(
      data = frame_max_net,
      aes(x = threshold_cm, y = maximum_net_to_date),
      color = "white",
      linewidth = 2.2,
      direction = "hv"
    ) +
    geom_step(
      data = frame_max_net,
      aes(
        x = threshold_cm,
        y = maximum_net_to_date,
        color = "Maximum net to date"
      ),
      linewidth = 1.1,
      linetype = "longdash",
      direction = "hv"
    ) +
    geom_vline(
      xintercept = c(0.15, 0.5, 1),
      color = "grey45",
      linewidth = 0.4,
      linetype = "dotted"
    ) +
    geom_vline(
      data = frame_info,
      aes(xintercept = mean_absolute_delta_cm, color = mean_direction),
      linewidth = 0.9,
      linetype = "longdash"
    ) +
    scale_x_continuous(
      trans = st28_threshold_axis_trans,
      limits = c(0, st28_tail_plot_max_cm),
      breaks = st28_tail_plot_breaks_cm,
      labels = ~ str_remove(
        format(.x, trim = TRUE, scientific = FALSE),
        "\\.?0+$"
      ),
      expand = expansion(mult = c(0, 0.01))
    ) +
    scale_y_continuous(
      labels = ~ scales::percent(abs(.x), accuracy = 1),
      breaks = seq(-0.75, 0.75, by = 0.25),
      expand = expansion(mult = c(0.01, 0.01))
    ) +
    coord_cartesian(ylim = c(-0.75, 0.75)) +
    scale_color_manual(
      values = c(
        "Accumulation" = "#b2182b",
        "Equivalent loss" = "#2166ac",
        "Current net balance" = "#1a1a1a",
        "Maximum net to date" = "#1b7837",
        "Positive transect mean" = "#7f0000",
        "Negative transect mean" = "#084594"
      )
    ) +
    scale_fill_manual(
      values = c("Accumulation" = "#d6604d", "Equivalent loss" = "#4393c3")
    ) +
    labs(
      title = paste0("ST28 empirical cumulative tails: ", frame_info$frame_label),
      subtitle = "Dashed green line remembers the maximum net balance reached at each threshold",
      x = "Absolute sediment-depth change threshold (cm; pseudo-log scale)",
      y = "Meter positions at threshold or greater",
      color = NULL,
      fill = NULL
    ) +
    guides(
      fill = "none",
      color = guide_legend(
        override.aes = list(
          linetype = c(
            "solid",
            "solid",
            "solid",
            "longdash",
            "longdash"
          ),
          linewidth = c(1, 1, 1, 1.1, 0.9)
        )
      )
    ) +
    theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      panel.border = element_rect(
        color = "grey75",
        fill = NA,
        linewidth = 0.4
      ),
      legend.position = "bottom",
      plot.title = element_text(face = "bold"),
      plot.title.position = "plot"
    )
}

st28_frame_dir <- file.path("code", "figures", "st28_animation_frames")
dir.create(st28_frame_dir, recursive = TRUE, showWarnings = FALSE)

st28_frame_files <- st28_animation_frames %>%
  transmute(
    survey_n,
    frame_file = file.path(
      st28_frame_dir,
      sprintf("st28_frame_%02d.png", survey_n)
    )
  )

pwalk(
  st28_frame_files,
  function(survey_n, frame_file) {
    ggsave(
      frame_file,
      plot = make_st28_empirical_frame(survey_n),
      width = 10,
      height = 7,
      units = "in",
      dpi = 96,
      bg = "white"
    )
  }
)

stopifnot(all(file.exists(st28_frame_files$frame_file)))

st28_final_survey_n <- max(st28_animation_frames$survey_n)
st28_final_net <- st28_animation_net %>%
  filter(survey_n == st28_final_survey_n)
st28_final_max_net <- st28_animation_max_net %>%
  filter(survey_n == st28_final_survey_n)
st28_final_frame_label <- st28_animation_frames %>%
  filter(survey_n == st28_final_survey_n) %>%
  pull(frame_label) %>%
  as.character()
st28_final_mean_delta <- st28_animation_frames %>%
  filter(survey_n == st28_final_survey_n) %>%
  select(mean_delta_cm, mean_absolute_delta_cm, mean_direction)

st28_final_summary_plot <- ggplot() +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.45) +
  geom_step(
    data = st28_final_net,
    aes(x = threshold_cm, y = tail_balance, color = "Most recent net balance"),
    linewidth = 1.15,
    direction = "hv"
  ) +
  geom_step(
    data = st28_final_max_net,
    aes(
      x = threshold_cm,
      y = maximum_net_to_date,
      color = "Maximum net over monitoring period"
    ),
    linewidth = 1.2,
    linetype = "longdash",
    direction = "hv"
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.4,
    linetype = "dotted"
  ) +
  geom_vline(
    data = st28_final_mean_delta,
    aes(xintercept = mean_absolute_delta_cm, color = mean_direction),
    linewidth = 0.9,
    linetype = "longdash"
  ) +
  scale_x_continuous(
    trans = st28_threshold_axis_trans,
    limits = c(0, st28_tail_plot_max_cm),
    breaks = st28_tail_plot_breaks_cm,
    labels = ~ str_remove(
      format(.x, trim = TRUE, scientific = FALSE),
      "\\.?0+$"
    ),
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    breaks = seq(-0.75, 0.75, by = 0.25),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  coord_cartesian(ylim = c(-0.75, 0.75)) +
  scale_color_manual(
    values = c(
      "Most recent net balance" = "#1a1a1a",
      "Maximum net over monitoring period" = "#1b7837",
      "Positive transect mean" = "#7f0000",
      "Negative transect mean" = "#084594"
    )
  ) +
  labs(
    title = paste0("ST28 net-balance summary: ", st28_final_frame_label),
    subtitle = "Black is the most recent net balance; dashed green is the maximum observed at each threshold",
    x = "Absolute sediment-depth change threshold (cm; pseudo-log scale)",
    y = "Accumulation minus equivalent-loss prevalence",
    color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey75", fill = NA, linewidth = 0.4),
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot"
  )

st28_final_summary_file <- file.path(
  st28_frame_dir,
  "st28_frame_final_summary.png"
)
ggsave(
  st28_final_summary_file,
  plot = st28_final_summary_plot,
  width = 10,
  height = 7,
  units = "in",
  dpi = 96,
  bg = "white"
)
```

```{r st28-interactive-empirical-player, echo=FALSE, results='asis'}
st28_player_id <- "st28-empirical-memory-player"
st28_player_frames <- c(
  map_chr(st28_frame_files$frame_file, knitr::image_uri),
  knitr::image_uri(st28_final_summary_file)
)
st28_player_frames_json <- jsonlite::toJSON(
  st28_player_frames,
  auto_unbox = TRUE
)

st28_player <- htmltools::tags$div(
  id = st28_player_id,
  class = "st28-player",
  htmltools::tags$style(htmltools::HTML(paste(
    ".st28-player { max-width: 960px; margin: 1rem auto; }",
    ".st28-player img { width: 100%; height: auto; display: block; cursor: pointer; border: 1px solid #d4d4d4; }",
    ".st28-player-controls { display: flex; align-items: center; gap: 0.65rem; margin-top: 0.55rem; }",
    ".st28-player button { border: 1px solid #777; background: #fff; color: #222; padding: 0.35rem 0.75rem; cursor: pointer; }",
    ".st28-player button:hover { background: #f0f0f0; }",
    ".st28-player input[type=range] { flex: 1; }",
    ".st28-player-status { min-width: 5.5rem; text-align: right; font-variant-numeric: tabular-nums; }",
    sep = "\n"
  ))),
  htmltools::tags$img(
    src = st28_player_frames[[1]],
    alt = "Animated ST28 empirical accumulation, loss, and net-balance curves"
  ),
  htmltools::tags$div(
    class = "st28-player-controls",
    htmltools::tags$button(type = "button", `data-action` = "play", "Play"),
    htmltools::tags$button(type = "button", `data-action` = "restart", "Restart"),
    htmltools::tags$input(
      type = "range",
      min = 0,
      max = length(st28_player_frames) - 1,
      value = 0,
      step = 1,
      `aria-label` = "Animation frame"
    ),
    htmltools::tags$span(
      class = "st28-player-status",
      paste0("1 / ", length(st28_player_frames))
    )
  ),
  htmltools::tags$script(htmltools::HTML(sprintf(paste0(
    "(function() {",
    "const root=document.getElementById('%s');",
    "const frames=%s;",
    "const image=root.querySelector('img');",
    "const playButton=root.querySelector('[data-action=play]');",
    "const restartButton=root.querySelector('[data-action=restart]');",
    "const slider=root.querySelector('input[type=range]');",
    "const status=root.querySelector('.st28-player-status');",
    "let index=0; let timer=null;",
    "function showFrame(nextIndex){",
    "index=Math.max(0,Math.min(nextIndex,frames.length-1));",
    "image.src=frames[index]; slider.value=index;",
    "status.textContent=(index+1)+' / '+frames.length;}",
    "function stop(){if(timer!==null)window.clearInterval(timer);",
    "timer=null; playButton.textContent=index===frames.length-1?'Play again':'Play';}",
    "function play(){if(timer!==null){stop();return;}",
    "if(index===frames.length-1)showFrame(0); playButton.textContent='Pause';",
    "timer=window.setInterval(function(){",
    "if(index>=frames.length-1){stop();return;} showFrame(index+1);",
    "if(index===frames.length-1)stop();},500);}",
    "playButton.addEventListener('click',play);",
    "image.addEventListener('click',play);",
    "restartButton.addEventListener('click',function(){stop();showFrame(0);});",
    "slider.addEventListener('input',function(){stop();showFrame(Number(this.value));});",
    "})();"
  ), st28_player_id, st28_player_frames_json)))
)

st28_player_html <- as.character(st28_player)
st28_player_html <- gsub(">[[:space:]]+<", "><", st28_player_html)
cat(st28_player_html)
```

#### Stacked net contributions by sediment-change band

This view decomposes cumulative net exceedance into three mutually exclusive
magnitude bands. The `1 cm+` segment is the net balance at 1 cm or greater; the
two lighter segments are the additional net contributions between successive
thresholds. The signed segment values sum to the cumulative net balance at
1.5 mm or greater.

```{r st28-stacked-net-threshold-contributions, fig.width=12, fig.height=8}
st28_cumulative_net_thresholds <-
  st28_alternative_exceedance_profiles %>%
  mutate(threshold_key_cm = round(threshold_cm, 2)) %>%
  filter(threshold_key_cm %in% c(0.15, 0.5, 1)) %>%
  mutate(
    survey_timepoint = factor(
      survey,
      levels = survey[order(survey_n)] %>% unique()
    )
  )

st28_stacked_net_wide <- st28_cumulative_net_thresholds %>%
  mutate(
    threshold_code = recode(
      as.character(threshold_key_cm),
      `0.15` = "net_1_5mm",
      `0.5` = "net_5mm",
      `1` = "net_1cm"
    )
  ) %>%
  select(
    date,
    survey,
    survey_n,
    survey_timepoint,
    threshold_code,
    tail_balance
  ) %>%
  pivot_wider(names_from = threshold_code, values_from = tail_balance) %>%
  mutate(
    `1 cm+` = net_1cm,
    `5 mm to <1 cm` = net_5mm - net_1cm,
    `1.5 mm to <5 mm` = net_1_5mm - net_5mm,
    total_net_1_5mm = net_1_5mm
  )

st28_stacked_net_contributions <- st28_stacked_net_wide %>%
  pivot_longer(
    c(`1 cm+`, `5 mm to <1 cm`, `1.5 mm to <5 mm`),
    names_to = "magnitude_band",
    values_to = "net_contribution"
  ) %>%
  mutate(
    magnitude_band = factor(
      magnitude_band,
      levels = c("1 cm+", "5 mm to <1 cm", "1.5 mm to <5 mm")
    ),
    contribution_direction = factor(
      if_else(net_contribution >= 0, "Accumulation", "Loss"),
      levels = c("Accumulation", "Loss")
    )
  )

st28_stacked_identity_check <- st28_stacked_net_contributions %>%
  summarize(
    reconstructed_total = sum(net_contribution),
    expected_total = first(total_net_1_5mm),
    .by = survey_n
  )

stopifnot(all(
  near(
    st28_stacked_identity_check$reconstructed_total,
    st28_stacked_identity_check$expected_total
  )
))

st28_stacked_net_threshold_plot <- ggplot(
  st28_stacked_net_contributions,
  aes(
    x = survey_timepoint,
    y = net_contribution,
    fill = contribution_direction,
    alpha = magnitude_band,
    group = magnitude_band
  )
) +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_col(
    width = 0.75,
    position = position_stack(reverse = TRUE)
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(-0.75, 0.75),
    breaks = seq(-0.75, 0.75, by = 0.25),
    expand = expansion(mult = c(0.03, 0.03))
  ) +
  scale_fill_manual(
    values = c("Accumulation" = "#b2182b", "Loss" = "#2166ac")
  ) +
  scale_alpha_manual(
    values = c(
      "1 cm+" = 1,
      "5 mm to <1 cm" = 0.65,
      "1.5 mm to <5 mm" = 0.35
    )
  ) +
  labs(
    title = "ST28 net exceedance decomposed by sediment-change band",
    subtitle = "Stacked segments are signed, mutually exclusive contributions; points below show the transect mean",
    x = NULL,
    y = "Net prevalence contribution",
    fill = "Direction",
    alpha = "Magnitude band"
  ) +
  guides(
    fill = guide_legend(order = 1),
    alpha = guide_legend(order = 2)
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st28_stacked_mean_plot <- st28_weekly_mean_delta %>%
  mutate(
    survey_timepoint = factor(
      survey,
      levels = levels(st28_stacked_net_contributions$survey_timepoint)
    )
  ) %>%
  ggplot(aes(x = survey_timepoint, y = mean_delta_cm)) +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_point(aes(color = mean_direction), size = 2.3) +
  scale_color_manual(
    values = c(
      "Positive transect mean" = "#7f0000",
      "Negative transect mean" = "#084594"
    ),
    guide = "none"
  ) +
  labs(
    x = "Construction survey",
    y = "Transect mean\ndelta (cm)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1)
  )

st28_stacked_net_threshold_plot / st28_stacked_mean_plot +
  plot_layout(heights = c(4, 1), guides = "collect") &
  theme(legend.position = "bottom")
```
