---
title: "Localized Sedimentation Analysis"
author: "Ross Cunning"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(ggrepel)
library(ggspatial)

project_dir <- if (basename(getwd()) == "code") dirname(getwd()) else getwd()
knitr::opts_knit$set(root.dir = project_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
```

## Data sources

```{r paths-and-helpers}
sediment_dir <- "data/sedimentation_monitoring"
gps_file <- file.path(sediment_dir, "PE_SB_Station_Transect_GPS.xlsx")

normalize_transect_id <- function(x) {
  x %>%
    str_squish() %>%
    str_replace_all("\\s+", "_")
}

model_sedimentation_class <- function(transect) {
  case_when(
    transect %in% c("ST22", "ST31") ~ "1.0 cm predicted",
    transect %in% c("ST28", "ST5") ~ "1.5 mm predicted",
    TRUE ~ "0 cm predicted"
  ) %>%
    factor(
      levels = c(
        "1.0 cm predicted",
        "1.5 mm predicted",
        "0 cm predicted"
      ),
      ordered = TRUE
    )
}

construction_files <- list.files(
  file.path(sediment_dir, "2_During-Construction", "Data"),
  pattern = "\\.xlsx$",
  recursive = TRUE,
  full.names = TRUE
)

construction_file_index <- tibble(source_file = construction_files) %>%
  mutate(
    source_file_name = basename(source_file),
    construction_survey_n = map_int(
      str_extract_all(source_file_name, "(?i)(?<=con)\\d+"),
      ~ if (length(.x) == 0) NA_integer_ else max(as.integer(.x))
    )
  ) %>%
  filter(!is.na(construction_survey_n)) %>%
  arrange(desc(construction_survey_n), desc(file.info(source_file)$mtime))

latest_construction_file <- construction_file_index %>%
  slice(1) %>%
  pull(source_file)
```

The transect inventory is taken from the newest cumulative construction
workbook: `r basename(latest_construction_file)`.

```{r transect-inventory}
monitoring_transects <- read_excel(
  latest_construction_file,
  sheet = "SedimentDepth_RawData"
) %>%
  clean_names() %>%
  transmute(
    transect = normalize_transect_id(transect),
    transect_type = str_squish(transect_type)
  ) %>%
  distinct()

transect_gps <- read_excel(gps_file, sheet = "Hardbottom") %>%
  clean_names() %>%
  transmute(
    transect = normalize_transect_id(
      str_remove(station_transect_id, "\\s+\\d+m$")
    ),
    meter = as.integer(str_extract(station_transect_id, "\\d+(?=m$)")),
    gps_note = notes,
    lat_dd = as.numeric(lat_dd),
    lon_dd = as.numeric(lon_dd)
  ) %>%
  filter(!is.na(transect), !is.na(meter)) %>%
  arrange(transect, meter) %>%
  left_join(monitoring_transects, by = "transect") %>%
  mutate(model_sedimentation_class = model_sedimentation_class(transect))

transect_coverage <- full_join(
  monitoring_transects %>% mutate(in_monitoring_data = TRUE),
  transect_gps %>%
    summarize(
      n_gps_points = n(),
      min_meter = min(meter),
      max_meter = max(meter),
      coordinates_complete = all(!is.na(lat_dd) & !is.na(lon_dd)),
      .by = transect
    ) %>%
    mutate(in_gps_file = TRUE),
  by = "transect"
) %>%
  mutate(
    across(c(in_monitoring_data, in_gps_file), ~ replace_na(.x, FALSE)),
    map_ready = in_monitoring_data & in_gps_file & coordinates_complete
  ) %>%
  arrange(transect_type, transect)

stopifnot(
  nrow(monitoring_transects) == 19,
  all(transect_coverage$map_ready)
)

transect_coverage
```

## Transect map

Lines connect the surveyed GPS anchors in meter order. Most 50 m transects have
coordinates at both ends; longer and adjacent transects include intermediate
anchors where they were supplied in the GPS workbook.

```{r transect-map-data, fig.width=11, fig.height=8.5}
transect_labels <- transect_gps %>%
  summarize(
    lon_dd = mean(range(lon_dd)),
    lat_dd = mean(range(lat_dd)),
    transect_type = first(transect_type),
    model_sedimentation_class = first(model_sedimentation_class),
    .by = transect
  )

transect_map <- ggplot(
  transect_gps,
  aes(
    x = lon_dd,
    y = lat_dd,
    group = transect,
    color = model_sedimentation_class
  )
) +
  geom_path(linewidth = 1.05, lineend = "round") +
  geom_point(size = 1.7, shape = 21, fill = "white", stroke = 0.7) +
  geom_text_repel(
    data = transect_labels,
    aes(label = transect),
    size = 3.2,
    fontface = "bold",
    color = "grey15",
    seed = 1406,
    box.padding = 0.3,
    point.padding = 0.15,
    force = 1.5,
    max.iter = 20000,
    min.segment.length = 0,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  annotation_scale(
    location = "bl",
    width_hint = 0.22,
    text_cex = 0.75
  ) +
  annotation_north_arrow(
    location = "tl",
    which_north = "true",
    height = unit(0.75, "cm"),
    width = unit(0.75, "cm"),
    style = north_arrow_minimal(text_size = 8)
  ) +
  scale_color_manual(
    values = c(
      "1.0 cm predicted" = "#c1121f",
      "1.5 mm predicted" = "#f28e2b",
      "0 cm predicted" = "#287271"
    ),
    drop = FALSE
  ) +
  coord_sf(
    crs = sf::st_crs(4326),
    default_crs = sf::st_crs(4326),
    expand = TRUE,
    datum = sf::st_crs(4326)
  ) +
  labs(
    title = "Sand Bypass sediment monitoring transects",
    subtitle = "GPS anchors and transect alignments",
    x = "Longitude",
    y = "Latitude",
    color = "Predicted sedimentation",
    caption = "Source: PE_SB_Station_Transect_GPS.xlsx"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.box = "vertical",
    plot.title.position = "plot"
  )

transect_map
```

## ST22 sediment deltas

For each survey, the three sediment-depth replicates are averaged at each meter
position. The baseline mean at that same meter position is then subtracted from
the construction-survey mean:

\[
\Delta_{m,t} = \overline{x}_{m,t} - \overline{x}_{m,baseline}
\]

Positive values indicate sediment accumulation relative to baseline. This
meter-matched calculation retains spatial variation that would be lost by first
averaging across the full transect.

```{r st22-meter-deltas}
st22_raw <- read_excel(
  latest_construction_file,
  sheet = "SedimentDepth_RawData"
) %>%
  clean_names() %>%
  mutate(
    date = as.Date(date),
    survey = str_squish(survey),
    survey_n = if_else(
      survey == "Baseline",
      0L,
      as.integer(str_extract(survey, "\\d+"))
    ),
    transect = normalize_transect_id(transect),
    meter = as.integer(meter),
    meter_mean_sediment_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    )
  ) %>%
  filter(transect == "ST22")

st22_baseline <- st22_raw %>%
  filter(survey == "Baseline") %>%
  transmute(
    meter,
    baseline_mean_sediment_cm = meter_mean_sediment_cm
  )

st22_meter_deltas <- st22_raw %>%
  filter(survey != "Baseline") %>%
  left_join(st22_baseline, by = "meter") %>%
  mutate(
    sediment_delta_cm =
      meter_mean_sediment_cm - baseline_mean_sediment_cm
  ) %>%
  select(
    date,
    survey,
    survey_n,
    transect,
    meter,
    baseline_mean_sediment_cm,
    meter_mean_sediment_cm,
    sediment_delta_cm
  ) %>%
  arrange(survey_n, meter)

st22_delta_coverage <- st22_meter_deltas %>%
  summarize(
    n_meter_positions = n(),
    n_missing_deltas = sum(is.na(sediment_delta_cm)),
    .by = c(date, survey, survey_n)
  ) %>%
  arrange(survey_n)

stopifnot(
  all(st22_delta_coverage$n_meter_positions == 51),
  all(st22_delta_coverage$n_missing_deltas == 0)
)

st22_delta_coverage
```

### Weekly density distributions

The density curves use a common horizontal scale. Fill colors classify the
delta represented by each portion of the curve, rather than classifying an
entire survey from a summary statistic.

```{r st22-density-data}
st22_delta_limits <- range(st22_meter_deltas$sediment_delta_cm) + c(-0.5, 0.5)

st22_density <- st22_meter_deltas %>%
  group_by(date, survey, survey_n) %>%
  group_modify(~ {
    density_estimate <- density(
      .x$sediment_delta_cm,
      from = st22_delta_limits[1],
      to = st22_delta_limits[2],
      n = 1024,
      na.rm = TRUE
    )

    tibble(
      sediment_delta_cm = density_estimate$x,
      density = density_estimate$y
    )
  }) %>%
  ungroup() %>%
  mutate(
    delta_band = case_when(
      sediment_delta_cm >= 1 ~ ">= 1 cm",
      sediment_delta_cm >= 0.5 ~ ">= 5 mm to < 1 cm",
      sediment_delta_cm >= 0.15 ~ ">= 1.5 mm to < 5 mm",
      TRUE ~ "< 1.5 mm"
    ),
    delta_band = factor(
      delta_band,
      levels = c(
        ">= 1 cm",
        ">= 5 mm to < 1 cm",
        ">= 1.5 mm to < 5 mm",
        "< 1.5 mm"
      )
    ),
    survey_label = paste0(
      survey,
      "\n",
      format(date, "%Y-%m-%d")
    ),
    survey_label = fct_reorder(survey_label, survey_n)
  )
```

```{r st22-density-plot, fig.width=11, fig.height=15}
st22_density_plot <- ggplot(
  st22_density,
  aes(x = sediment_delta_cm, y = density)
) +
  geom_area(
    aes(fill = delta_band, group = delta_band),
    color = NA,
    alpha = 0.9
  ) +
  geom_line(linewidth = 0.45, color = "grey15") +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_fill_manual(
    values = c(
      ">= 1 cm" = "#d73027",
      ">= 5 mm to < 1 cm" = "#f28e2b",
      ">= 1.5 mm to < 5 mm" = "#ffd54f",
      "< 1.5 mm" = "white"
    ),
    drop = FALSE
  ) +
  scale_x_continuous(
    breaks = seq(
      floor(st22_delta_limits[1] / 2) * 2,
      ceiling(st22_delta_limits[2] / 2) * 2,
      by = 2
    )
  ) +
  labs(
    title = "ST22 sediment accumulation relative to baseline",
    subtitle = "Density of meter-level mean sediment-depth deltas by construction survey",
    x = "Sediment-depth delta from meter-matched baseline (cm)",
    y = "Density",
    fill = "Delta range"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_density_plot
```

### Accumulation versus mirrored loss

This view compares accumulation and loss at the same absolute magnitude. The
solid curve is the density at positive deltas. The dashed curve reflects the
negative tail onto the positive axis. Colored ribbons show only magnitudes at
which accumulation density is greater than the density of equivalent sediment
loss. Where the dashed curve is higher, loss is more prevalent and no colored
ribbon is drawn.

```{r st22-tail-balance-data}
st22_max_absolute_delta <-
  max(abs(st22_meter_deltas$sediment_delta_cm)) + 0.5

st22_mirrored_density <- st22_meter_deltas %>%
  group_by(date, survey, survey_n) %>%
  group_modify(~ {
    density_estimate <- density(
      .x$sediment_delta_cm,
      from = -st22_max_absolute_delta,
      to = st22_max_absolute_delta,
      n = 2049,
      na.rm = TRUE
    )

    positive_magnitude <- density_estimate$x[
      density_estimate$x >= 0
    ]

    tibble(
      magnitude_cm = positive_magnitude,
      accumulation_density = approx(
        density_estimate$x,
        density_estimate$y,
        xout = positive_magnitude
      )$y,
      mirrored_loss_density = approx(
        density_estimate$x,
        density_estimate$y,
        xout = -positive_magnitude
      )$y
    )
  }) %>%
  ungroup() %>%
  mutate(
    accumulation_exceeds_loss =
      accumulation_density > mirrored_loss_density,
    delta_band = case_when(
      magnitude_cm >= 1 ~ ">= 1 cm",
      magnitude_cm >= 0.5 ~ ">= 5 mm to < 1 cm",
      magnitude_cm >= 0.15 ~ ">= 1.5 mm to < 5 mm",
      TRUE ~ "< 1.5 mm"
    ),
    delta_band = factor(
      delta_band,
      levels = c(
        ">= 1 cm",
        ">= 5 mm to < 1 cm",
        ">= 1.5 mm to < 5 mm",
        "< 1.5 mm"
      )
    ),
    excess_run = cumsum(
      accumulation_exceeds_loss != lag(
        accumulation_exceeds_loss,
        default = first(accumulation_exceeds_loss)
      ) |
        delta_band != lag(delta_band, default = first(delta_band))
    ),
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n),
    .by = c(date, survey, survey_n)
  )

st22_mirrored_curves <- st22_mirrored_density %>%
  select(
    date,
    survey,
    survey_n,
    survey_label,
    magnitude_cm,
    accumulation_density,
    mirrored_loss_density
  ) %>%
  pivot_longer(
    c(accumulation_density, mirrored_loss_density),
    names_to = "curve",
    values_to = "density"
  ) %>%
  mutate(
    curve = recode(
      curve,
      accumulation_density = "Accumulation (+delta)",
      mirrored_loss_density = "Mirrored loss (-delta)"
    ),
    curve = factor(
      curve,
      levels = c("Accumulation (+delta)", "Mirrored loss (-delta)")
    )
  )

tail_balance_thresholds <- tribble(
  ~threshold_cm, ~threshold_label,
  0.15, "1.5 mm",
  0.5, "5 mm",
  1, "1 cm"
)

st22_weekly_tail_balance <- crossing(
  st22_meter_deltas,
  tail_balance_thresholds
) %>%
  summarize(
    n_meter_positions = n(),
    accumulation_fraction = mean(sediment_delta_cm >= threshold_cm),
    loss_fraction = mean(sediment_delta_cm <= -threshold_cm),
    tail_balance = accumulation_fraction - loss_fraction,
    .by = c(
      date,
      survey,
      survey_n,
      threshold_cm,
      threshold_label
    )
  ) %>%
  arrange(survey_n, threshold_cm)
```

```{r st22-tail-balance-plot, fig.width=11, fig.height=15}
st22_tail_balance_plot <- ggplot(
  st22_mirrored_curves,
  aes(x = magnitude_cm, y = density)
) +
  geom_ribbon(
    data = st22_mirrored_density %>%
      filter(accumulation_exceeds_loss),
    aes(
      x = magnitude_cm,
      ymin = mirrored_loss_density,
      ymax = accumulation_density,
      fill = delta_band,
      group = interaction(delta_band, excess_run)
    ),
    inherit.aes = FALSE,
    alpha = 0.9
  ) +
  geom_line(
    aes(color = curve, linetype = curve),
    linewidth = 0.55
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey55",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_fill_manual(
    values = c(
      ">= 1 cm" = "#d73027",
      ">= 5 mm to < 1 cm" = "#f28e2b",
      ">= 1.5 mm to < 5 mm" = "#ffd54f",
      "< 1.5 mm" = "white"
    ),
    drop = FALSE
  ) +
  scale_color_manual(
    values = c(
      "Accumulation (+delta)" = "grey10",
      "Mirrored loss (-delta)" = "#2c7fb8"
    )
  ) +
  scale_linetype_manual(
    values = c(
      "Accumulation (+delta)" = "solid",
      "Mirrored loss (-delta)" = "22"
    )
  ) +
  scale_x_continuous(
    breaks = seq(0, floor(st22_max_absolute_delta), by = 2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    title = "ST22 accumulation compared with equivalent sediment loss",
    subtitle = paste(
      "Colored area marks density where accumulation is more prevalent",
      "than loss of the same magnitude"
    ),
    x = "Absolute sediment-depth change from baseline (cm)",
    y = "Density",
    fill = "Excess accumulation range",
    color = NULL,
    linetype = NULL
  ) +
  guides(
    color = guide_legend(order = 1),
    linetype = guide_legend(order = 1),
    fill = guide_legend(order = 2)
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    legend.box = "vertical",
    plot.title.position = "plot"
  )

st22_tail_balance_plot
```

The table below reports the same comparison directly from the 51 observed
meter positions. Tail balance is the fraction at or above the positive
threshold minus the fraction at or below the equivalent negative threshold.
Positive values indicate that accumulation is more prevalent than equivalent
loss.

```{r st22-tail-balance-table}
st22_weekly_tail_balance %>%
  select(
    date,
    survey,
    survey_n,
    threshold_label,
    accumulation_fraction,
    loss_fraction,
    tail_balance
  ) %>%
  pivot_wider(
    names_from = threshold_label,
    values_from = c(
      accumulation_fraction,
      loss_fraction,
      tail_balance
    ),
    names_glue = "{.value}_{threshold_label}"
  ) %>%
  arrange(survey_n)
```

## ST22 alternative Baseline-Con1-Con2 reference

This sensitivity analysis defines a meter-specific alternative reference as
the arithmetic mean of the Baseline, Con1, and Con2 meter means. Each survey
receives equal weight. Because Con1 and Con2 contribute to the reference,
alternative-reference results begin with Con3. This does not replace the BMP
baseline analysis above.

```{r st22-alternative-reference}
st22_alternative_reference <- st22_raw %>%
  filter(survey_n %in% 0:2) %>%
  summarize(
    alternative_reference_cm = mean(meter_mean_sediment_cm),
    n_reference_surveys = n(),
    .by = meter
  ) %>%
  left_join(st22_baseline, by = "meter") %>%
  mutate(
    reference_difference_cm =
      alternative_reference_cm - baseline_mean_sediment_cm
  )

stopifnot(
  nrow(st22_alternative_reference) == 51,
  all(st22_alternative_reference$n_reference_surveys == 3)
)

st22_reference_summary <- st22_alternative_reference %>%
  summarize(
    n_meter_positions = n(),
    official_baseline_mean_cm = mean(baseline_mean_sediment_cm),
    alternative_reference_mean_cm = mean(alternative_reference_cm),
    mean_reference_difference_cm = mean(reference_difference_cm),
    median_reference_difference_cm = median(reference_difference_cm),
    pct_alternative_lower = mean(reference_difference_cm < 0),
    pct_alternative_higher = mean(reference_difference_cm > 0)
  )

st22_reference_summary
```

```{r st22-alternative-deltas}
st22_alternative_deltas <- st22_raw %>%
  filter(survey_n >= 3) %>%
  left_join(
    st22_alternative_reference %>%
      select(meter, alternative_reference_cm),
    by = "meter"
  ) %>%
  left_join(st22_baseline, by = "meter") %>%
  mutate(
    alternative_delta_cm =
      meter_mean_sediment_cm - alternative_reference_cm,
    official_delta_cm =
      meter_mean_sediment_cm - baseline_mean_sediment_cm
  ) %>%
  select(
    date,
    survey,
    survey_n,
    transect,
    meter,
    meter_mean_sediment_cm,
    baseline_mean_sediment_cm,
    alternative_reference_cm,
    official_delta_cm,
    alternative_delta_cm
  ) %>%
  arrange(survey_n, meter)
```

### Alternative-reference density distributions

```{r st22-alternative-density-data}
st22_alternative_delta_limits <-
  range(st22_alternative_deltas$alternative_delta_cm) + c(-0.5, 0.5)

st22_alternative_density <- st22_alternative_deltas %>%
  group_by(date, survey, survey_n) %>%
  group_modify(~ {
    density_estimate <- density(
      .x$alternative_delta_cm,
      from = st22_alternative_delta_limits[1],
      to = st22_alternative_delta_limits[2],
      n = 1024,
      na.rm = TRUE
    )

    tibble(
      sediment_delta_cm = density_estimate$x,
      density = density_estimate$y
    )
  }) %>%
  ungroup() %>%
  mutate(
    delta_band = case_when(
      sediment_delta_cm >= 1 ~ ">= 1 cm",
      sediment_delta_cm >= 0.5 ~ ">= 5 mm to < 1 cm",
      sediment_delta_cm >= 0.15 ~ ">= 1.5 mm to < 5 mm",
      TRUE ~ "< 1.5 mm"
    ),
    delta_band = factor(
      delta_band,
      levels = c(
        ">= 1 cm",
        ">= 5 mm to < 1 cm",
        ">= 1.5 mm to < 5 mm",
        "< 1.5 mm"
      )
    ),
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n)
  )
```

```{r st22-alternative-density-plot, fig.width=11, fig.height=14}
st22_alternative_density_plot <- ggplot(
  st22_alternative_density,
  aes(x = sediment_delta_cm, y = density)
) +
  geom_area(
    aes(fill = delta_band, group = delta_band),
    color = NA,
    alpha = 0.9
  ) +
  geom_line(linewidth = 0.45, color = "grey15") +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_fill_manual(
    values = c(
      ">= 1 cm" = "#d73027",
      ">= 5 mm to < 1 cm" = "#f28e2b",
      ">= 1.5 mm to < 5 mm" = "#ffd54f",
      "< 1.5 mm" = "white"
    ),
    drop = FALSE
  ) +
  scale_x_continuous(
    breaks = seq(
      floor(st22_alternative_delta_limits[1] / 2) * 2,
      ceiling(st22_alternative_delta_limits[2] / 2) * 2,
      by = 2
    )
  ) +
  labs(
    title = "ST22 sediment accumulation using the alternative reference",
    subtitle = "Reference is the meter-specific mean of Baseline, Con1, and Con2",
    x = "Sediment-depth delta from alternative reference (cm)",
    y = "Density",
    fill = "Delta range"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_alternative_density_plot
```

### Alternative-reference accumulation versus mirrored loss

```{r st22-alternative-tail-balance-data}
st22_alternative_max_absolute_delta <-
  max(abs(st22_alternative_deltas$alternative_delta_cm)) + 0.5

st22_alternative_mirrored_density <- st22_alternative_deltas %>%
  group_by(date, survey, survey_n) %>%
  group_modify(~ {
    density_estimate <- density(
      .x$alternative_delta_cm,
      from = -st22_alternative_max_absolute_delta,
      to = st22_alternative_max_absolute_delta,
      n = 2049,
      na.rm = TRUE
    )

    positive_magnitude <- density_estimate$x[
      density_estimate$x >= 0
    ]

    tibble(
      magnitude_cm = positive_magnitude,
      accumulation_density = approx(
        density_estimate$x,
        density_estimate$y,
        xout = positive_magnitude
      )$y,
      mirrored_loss_density = approx(
        density_estimate$x,
        density_estimate$y,
        xout = -positive_magnitude
      )$y
    )
  }) %>%
  ungroup() %>%
  mutate(
    accumulation_exceeds_loss =
      accumulation_density > mirrored_loss_density,
    delta_band = case_when(
      magnitude_cm >= 1 ~ ">= 1 cm",
      magnitude_cm >= 0.5 ~ ">= 5 mm to < 1 cm",
      magnitude_cm >= 0.15 ~ ">= 1.5 mm to < 5 mm",
      TRUE ~ "< 1.5 mm"
    ),
    delta_band = factor(
      delta_band,
      levels = c(
        ">= 1 cm",
        ">= 5 mm to < 1 cm",
        ">= 1.5 mm to < 5 mm",
        "< 1.5 mm"
      )
    ),
    excess_run = cumsum(
      accumulation_exceeds_loss != lag(
        accumulation_exceeds_loss,
        default = first(accumulation_exceeds_loss)
      ) |
        delta_band != lag(delta_band, default = first(delta_band))
    ),
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n),
    .by = c(date, survey, survey_n)
  )

st22_alternative_mirrored_curves <- st22_alternative_mirrored_density %>%
  select(
    date,
    survey,
    survey_n,
    survey_label,
    magnitude_cm,
    accumulation_density,
    mirrored_loss_density
  ) %>%
  pivot_longer(
    c(accumulation_density, mirrored_loss_density),
    names_to = "curve",
    values_to = "density"
  ) %>%
  mutate(
    curve = recode(
      curve,
      accumulation_density = "Accumulation (+delta)",
      mirrored_loss_density = "Mirrored loss (-delta)"
    ),
    curve = factor(
      curve,
      levels = c("Accumulation (+delta)", "Mirrored loss (-delta)")
    )
  )
```

```{r st22-alternative-tail-balance-plot, fig.width=11, fig.height=15}
st22_alternative_tail_balance_plot <- ggplot(
  st22_alternative_mirrored_curves,
  aes(x = magnitude_cm, y = density)
) +
  geom_ribbon(
    data = st22_alternative_mirrored_density %>%
      filter(accumulation_exceeds_loss),
    aes(
      x = magnitude_cm,
      ymin = mirrored_loss_density,
      ymax = accumulation_density,
      fill = delta_band,
      group = interaction(delta_band, excess_run)
    ),
    inherit.aes = FALSE,
    alpha = 0.9
  ) +
  geom_line(
    aes(color = curve, linetype = curve),
    linewidth = 0.55
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey55",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_fill_manual(
    values = c(
      ">= 1 cm" = "#d73027",
      ">= 5 mm to < 1 cm" = "#f28e2b",
      ">= 1.5 mm to < 5 mm" = "#ffd54f",
      "< 1.5 mm" = "white"
    ),
    drop = FALSE
  ) +
  scale_color_manual(
    values = c(
      "Accumulation (+delta)" = "grey10",
      "Mirrored loss (-delta)" = "#2c7fb8"
    )
  ) +
  scale_linetype_manual(
    values = c(
      "Accumulation (+delta)" = "solid",
      "Mirrored loss (-delta)" = "22"
    )
  ) +
  scale_x_continuous(
    breaks = seq(0, floor(st22_alternative_max_absolute_delta), by = 2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    title = "ST22 tail balance using the alternative reference",
    subtitle = "Reference is the meter-specific mean of Baseline, Con1, and Con2",
    x = "Absolute sediment-depth change from alternative reference (cm)",
    y = "Density",
    fill = "Excess accumulation range",
    color = NULL,
    linetype = NULL
  ) +
  guides(
    color = guide_legend(order = 1),
    linetype = guide_legend(order = 1),
    fill = guide_legend(order = 2)
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    legend.box = "vertical",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(b = 4)),
    plot.subtitle = element_text(margin = margin(b = 8))
  )

st22_alternative_tail_balance_plot
```

### Reference comparison

```{r st22-reference-comparison}
st22_reference_comparison <- st22_alternative_deltas %>%
  select(
    date,
    survey,
    survey_n,
    meter,
    official_delta_cm,
    alternative_delta_cm
  ) %>%
  pivot_longer(
    c(official_delta_cm, alternative_delta_cm),
    names_to = "reference",
    values_to = "sediment_delta_cm"
  ) %>%
  mutate(
    reference = recode(
      reference,
      official_delta_cm = "Official baseline",
      alternative_delta_cm = "Baseline-Con1-Con2 average"
    )
  ) %>%
  crossing(tail_balance_thresholds) %>%
  summarize(
    n_meter_positions = n(),
    n_accumulation_exceedances =
      sum(sediment_delta_cm >= threshold_cm),
    pct_accumulation_exceedances =
      mean(sediment_delta_cm >= threshold_cm),
    n_equivalent_losses =
      sum(sediment_delta_cm <= -threshold_cm),
    pct_equivalent_losses =
      mean(sediment_delta_cm <= -threshold_cm),
    tail_balance =
      pct_accumulation_exceedances - pct_equivalent_losses,
    .by = c(
      date,
      survey,
      survey_n,
      reference,
      threshold_cm,
      threshold_label
    )
  ) %>%
  arrange(survey_n, threshold_cm, reference)

st22_reference_comparison
```

## ST22 tail-balance visualization comparison

The following views use empirical exceedance fractions rather than smoothed
density estimates. At magnitude \(t\), accumulation exceedance is the fraction
of meter positions with \(\Delta \ge t\), loss exceedance is the fraction with
\(\Delta \le -t\), and tail balance is their difference.

```{r st22-empirical-tail-profiles}
st22_reference_deltas_long <- st22_alternative_deltas %>%
  select(
    date,
    survey,
    survey_n,
    meter,
    official_delta_cm,
    alternative_delta_cm
  ) %>%
  pivot_longer(
    c(official_delta_cm, alternative_delta_cm),
    names_to = "reference",
    values_to = "sediment_delta_cm"
  ) %>%
  mutate(
    reference = recode(
      reference,
      official_delta_cm = "Official baseline",
      alternative_delta_cm = "Baseline-Con1-Con2 average"
    ),
    reference = factor(
      reference,
      levels = c("Official baseline", "Baseline-Con1-Con2 average")
    ),
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n)
  )

tail_profile_max_cm <- ceiling(
  max(abs(st22_reference_deltas_long$sediment_delta_cm))
)

tail_profile_thresholds <- tibble(
  threshold_cm = seq(0, tail_profile_max_cm, by = 0.05)
)

st22_empirical_tail_profiles <- crossing(
  st22_reference_deltas_long,
  tail_profile_thresholds
) %>%
  summarize(
    accumulation_fraction = if_else(
      first(threshold_cm) == 0,
      mean(sediment_delta_cm > 0),
      mean(sediment_delta_cm >= threshold_cm)
    ),
    loss_fraction = if_else(
      first(threshold_cm) == 0,
      mean(sediment_delta_cm < 0),
      mean(sediment_delta_cm <= -threshold_cm)
    ),
    tail_balance = accumulation_fraction - loss_fraction,
    .by = c(
      date,
      survey,
      survey_n,
      survey_label,
      reference,
      threshold_cm
    )
  ) %>%
  arrange(reference, survey_n, threshold_cm) %>%
  mutate(
    balance_direction = case_when(
      tail_balance > 0 ~ "Accumulation exceeds loss",
      tail_balance < 0 ~ "Loss exceeds accumulation",
      TRUE ~ "Balanced"
    ),
    balance_direction = factor(
      balance_direction,
      levels = c(
        "Accumulation exceeds loss",
        "Loss exceeds accumulation",
        "Balanced"
      )
    ),
    direction_run = cumsum(
      balance_direction != lag(
        balance_direction,
        default = first(balance_direction)
      )
    ),
    .by = c(reference, survey_n)
  )
```

### Empirical exceedance curves

```{r st22-exceedance-curves, fig.width=11, fig.height=15}
st22_alternative_exceedance_profiles <- st22_empirical_tail_profiles %>%
  filter(reference == "Baseline-Con1-Con2 average")

st22_alternative_exceedance_curves <-
  st22_alternative_exceedance_profiles %>%
  select(
    date,
    survey,
    survey_n,
    survey_label,
    threshold_cm,
    accumulation_fraction,
    loss_fraction
  ) %>%
  pivot_longer(
    c(accumulation_fraction, loss_fraction),
    names_to = "tail",
    values_to = "fraction"
  ) %>%
  mutate(
    tail = recode(
      tail,
      accumulation_fraction = "Accumulation",
      loss_fraction = "Equivalent loss"
    ),
    tail = factor(tail, levels = c("Accumulation", "Equivalent loss"))
  )

st22_exceedance_curve_plot <- ggplot(
  st22_alternative_exceedance_curves,
  aes(x = threshold_cm, y = fraction)
) +
  geom_ribbon(
    data = st22_alternative_exceedance_profiles %>%
      filter(balance_direction != "Balanced"),
    aes(
      x = threshold_cm,
      ymin = pmin(accumulation_fraction, loss_fraction),
      ymax = pmax(accumulation_fraction, loss_fraction),
      fill = balance_direction,
      group = interaction(survey_n, direction_run)
    ),
    inherit.aes = FALSE,
    alpha = 0.55
  ) +
  geom_step(
    aes(color = tail, linetype = tail),
    linewidth = 0.55,
    direction = "hv"
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_x_continuous(
    limits = c(0, tail_profile_max_cm),
    breaks = seq(0, tail_profile_max_cm, by = 2),
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(0, 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_color_manual(
    values = c("Accumulation" = "#b2182b", "Equivalent loss" = "#2166ac")
  ) +
  scale_linetype_manual(
    values = c("Accumulation" = "solid", "Equivalent loss" = "22")
  ) +
  scale_fill_manual(
    values = c(
      "Accumulation exceeds loss" = "#ef8a62",
      "Loss exceeds accumulation" = "#67a9cf"
    )
  ) +
  labs(
    title = "ST22 empirical accumulation and loss exceedance curves",
    subtitle = "Baseline-Con1-Con2 alternative reference",
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Meter positions exceeding threshold",
    color = NULL,
    linetype = NULL,
    fill = "Tail balance"
  ) +
  guides(
    color = guide_legend(order = 1),
    linetype = guide_legend(order = 1),
    fill = guide_legend(order = 2)
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    legend.box = "vertical",
    plot.title.position = "plot"
  )

st22_exceedance_curve_plot
```

### Net accumulation-exceedance curves

This is the direct difference between the two curves above. At each sediment-
change threshold, the curve is the percentage of meter positions exceeding the
threshold through accumulation minus the percentage exceeding the same
threshold through loss. Positive values indicate excess accumulation, negative
values indicate excess loss, and zero indicates equal prevalence. The view is
limited to 3 cm so that the regulatory thresholds and the main body of the
empirical distributions remain legible.

```{r st22-net-exceedance-curves, fig.width=11, fig.height=15}
st22_net_exceedance_profiles <- st22_alternative_exceedance_profiles %>%
  filter(threshold_cm <= 3)

st22_net_exceedance_curve_plot <- ggplot(
  st22_net_exceedance_profiles,
  aes(x = threshold_cm, y = tail_balance)
) +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_ribbon(
    aes(
      ymin = pmin(tail_balance, 0),
      ymax = pmax(tail_balance, 0),
      fill = balance_direction,
      group = interaction(survey_n, direction_run)
    ),
    alpha = 0.7
  ) +
  geom_step(
    color = "grey15",
    linewidth = 0.55,
    direction = "hv"
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_x_continuous(
    limits = c(0, 3),
    breaks = 0:3,
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.25),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_fill_manual(
    values = c(
      "Accumulation exceeds loss" = "#d6604d",
      "Loss exceeds accumulation" = "#4393c3",
      "Balanced" = "grey80"
    ),
    drop = TRUE
  ) +
  labs(
    title = "ST22 net accumulation exceedance",
    subtitle = paste(
      "Baseline-Con1-Con2 alternative reference; positive = accumulation",
      "exceeds equivalent loss"
    ),
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Accumulation minus loss prevalence",
    fill = NULL
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_net_exceedance_curve_plot
```

### Smoothed cumulative net exceedance

This experimental view estimates a kernel density for each survey and then
integrates its two tails. At threshold (t), the smooth curve is the estimated
(P(\Delta \ge t) - P(\Delta \le -t)); it is not a comparison of the density
heights at (+t) and (-t). The larger points retain the unsmoothed empirical
results at the three regulatory thresholds.

```{r st22-smoothed-net-exceedance-data}
smooth_tail_profile <- function(delta, thresholds = seq(0, 3, by = 0.01)) {
  kernel <- density(delta, bw = "nrd0", n = 4096, cut = 4)
  interval_area <- diff(kernel$x) *
    (head(kernel$y, -1) + tail(kernel$y, -1)) / 2
  kernel_cdf <- c(0, cumsum(interval_area))
  kernel_cdf <- kernel_cdf / max(kernel_cdf)
  cdf_at <- approxfun(
    kernel$x,
    kernel_cdf,
    yleft = 0,
    yright = 1
  )

  tibble(
    threshold_cm = thresholds,
    smoothed_accumulation_fraction = 1 - cdf_at(thresholds),
    smoothed_loss_fraction = cdf_at(-thresholds),
    smoothed_tail_balance =
      smoothed_accumulation_fraction - smoothed_loss_fraction
  )
}

st22_smoothed_net_profiles <- st22_alternative_deltas %>%
  group_by(date, survey, survey_n) %>%
  group_modify(~ smooth_tail_profile(.x$alternative_delta_cm)) %>%
  ungroup() %>%
  mutate(
    survey_label = paste0(survey, "\n", format(date, "%Y-%m-%d")),
    survey_label = fct_reorder(survey_label, survey_n),
    balance_direction = if_else(
      smoothed_tail_balance >= 0,
      "Accumulation exceeds loss",
      "Loss exceeds accumulation"
    ),
    direction_run = cumsum(
      balance_direction != lag(
        balance_direction,
        default = first(balance_direction)
      )
    ),
    .by = survey_n
  )

st22_empirical_regulatory_points <-
  st22_alternative_exceedance_profiles %>%
  filter(threshold_cm %in% c(0.15, 0.5, 1))
```

```{r st22-smoothed-net-exceedance-plot, fig.width=11, fig.height=15}
st22_smoothed_net_exceedance_plot <- ggplot(
  st22_smoothed_net_profiles,
  aes(x = threshold_cm, y = smoothed_tail_balance)
) +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_ribbon(
    aes(
      ymin = pmin(smoothed_tail_balance, 0),
      ymax = pmax(smoothed_tail_balance, 0),
      fill = balance_direction,
      group = interaction(survey_n, direction_run)
    ),
    alpha = 0.65
  ) +
  geom_line(color = "grey15", linewidth = 0.65) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  geom_point(
    data = st22_empirical_regulatory_points,
    aes(
      x = threshold_cm,
      y = tail_balance,
      fill = balance_direction
    ),
    inherit.aes = FALSE,
    shape = 21,
    color = "grey10",
    stroke = 0.45,
    size = 2.2
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_x_continuous(
    limits = c(0, 3),
    breaks = 0:3,
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.25),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_fill_manual(
    values = c(
      "Accumulation exceeds loss" = "#d6604d",
      "Loss exceeds accumulation" = "#4393c3"
    )
  ) +
  labs(
    title = "ST22 smoothed cumulative net exceedance",
    subtitle = paste(
      "Kernel-density tail balance; outlined points are unsmoothed values at",
      "1.5 mm, 5 mm, and 1 cm"
    ),
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Estimated accumulation minus loss prevalence",
    fill = NULL
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_smoothed_net_exceedance_plot
```

### Mirrored cumulative accumulation and loss

This view shows the two cumulative tail prevalences directly rather than their
difference. The red curve above zero is the estimated fraction accumulating at
least the threshold amount. The blue curve below zero is the estimated fraction
losing at least the same amount. Distance from zero is therefore an actual
prevalence, and both sides contract toward zero as the threshold increases.
Outlined points show the corresponding unsmoothed empirical fractions at the
three regulatory thresholds.

```{r st22-mirrored-cumulative-data}
st22_smoothed_mirrored_profiles <- st22_smoothed_net_profiles %>%
  select(
    date,
    survey,
    survey_n,
    survey_label,
    threshold_cm,
    smoothed_accumulation_fraction,
    smoothed_loss_fraction
  ) %>%
  pivot_longer(
    c(smoothed_accumulation_fraction, smoothed_loss_fraction),
    names_to = "tail",
    values_to = "fraction"
  ) %>%
  mutate(
    tail = recode(
      tail,
      smoothed_accumulation_fraction = "Accumulation",
      smoothed_loss_fraction = "Equivalent loss"
    ),
    tail = factor(tail, levels = c("Accumulation", "Equivalent loss")),
    signed_fraction = if_else(tail == "Accumulation", fraction, -fraction)
  )

st22_empirical_mirrored_points <- st22_empirical_regulatory_points %>%
  select(
    date,
    survey,
    survey_n,
    survey_label,
    threshold_cm,
    accumulation_fraction,
    loss_fraction
  ) %>%
  pivot_longer(
    c(accumulation_fraction, loss_fraction),
    names_to = "tail",
    values_to = "fraction"
  ) %>%
  mutate(
    tail = recode(
      tail,
      accumulation_fraction = "Accumulation",
      loss_fraction = "Equivalent loss"
    ),
    tail = factor(tail, levels = c("Accumulation", "Equivalent loss")),
    signed_fraction = if_else(tail == "Accumulation", fraction, -fraction)
  )
```

```{r st22-mirrored-cumulative-plot, fig.width=11, fig.height=15}
st22_mirrored_cumulative_plot <- ggplot(
  st22_smoothed_mirrored_profiles,
  aes(
    x = threshold_cm,
    y = signed_fraction,
    color = tail,
    fill = tail,
    group = tail
  )
) +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_area(alpha = 0.4, position = "identity") +
  geom_line(linewidth = 0.7) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  geom_point(
    data = st22_empirical_mirrored_points,
    aes(
      x = threshold_cm,
      y = signed_fraction,
      color = tail,
      fill = tail
    ),
    inherit.aes = FALSE,
    shape = 21,
    stroke = 0.5,
    size = 2.2
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_x_continuous(
    limits = c(0, 3),
    breaks = 0:3,
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = ~ scales::percent(abs(.x), accuracy = 1),
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.25),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_color_manual(
    values = c("Accumulation" = "#b2182b", "Equivalent loss" = "#2166ac")
  ) +
  scale_fill_manual(
    values = c("Accumulation" = "#d6604d", "Equivalent loss" = "#4393c3")
  ) +
  labs(
    title = "ST22 mirrored cumulative accumulation and loss",
    subtitle = paste(
      "Distance from zero is the prevalence at that threshold or greater;",
      "outlined points are empirical values"
    ),
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Meter positions at threshold or greater",
    color = NULL,
    fill = NULL
  ) +
  guides(fill = "none") +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_mirrored_cumulative_plot
```

### Mirrored empirical curves with net balance

The red and blue curves remain the actual empirical accumulation and loss
prevalences. The black curve overlays their difference: accumulation prevalence
minus equivalent-loss prevalence.

```{r st22-mirrored-empirical-with-net, fig.width=11, fig.height=15}
st22_empirical_mirrored_profiles <-
  st22_alternative_exceedance_profiles %>%
  filter(threshold_cm <= 3) %>%
  select(
    date,
    survey,
    survey_n,
    survey_label,
    threshold_cm,
    accumulation_fraction,
    loss_fraction
  ) %>%
  pivot_longer(
    c(accumulation_fraction, loss_fraction),
    names_to = "tail",
    values_to = "fraction"
  ) %>%
  mutate(
    tail = recode(
      tail,
      accumulation_fraction = "Accumulation",
      loss_fraction = "Equivalent loss"
    ),
    tail = factor(tail, levels = c("Accumulation", "Equivalent loss")),
    signed_fraction = if_else(tail == "Accumulation", fraction, -fraction),
    threshold_end = lead(threshold_cm, default = 3),
    .by = c(survey_n, tail)
  )

st22_mirrored_empirical_net_plot <- ggplot() +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_rect(
    data = st22_empirical_mirrored_profiles,
    aes(
      xmin = threshold_cm,
      xmax = threshold_end,
      ymin = pmin(signed_fraction, 0),
      ymax = pmax(signed_fraction, 0),
      fill = tail
    ),
    alpha = 0.4,
    color = NA
  ) +
  geom_step(
    data = st22_empirical_mirrored_profiles,
    aes(
      x = threshold_cm,
      y = signed_fraction,
      color = tail,
      group = tail
    ),
    linewidth = 0.65,
    direction = "hv"
  ) +
  geom_step(
    data = st22_net_exceedance_profiles,
    aes(x = threshold_cm, y = tail_balance),
    color = "white",
    linewidth = 1.4,
    direction = "hv"
  ) +
  geom_step(
    data = st22_net_exceedance_profiles,
    aes(x = threshold_cm, y = tail_balance, color = "Net balance"),
    linewidth = 0.7,
    direction = "hv"
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_x_continuous(
    limits = c(0, 3),
    breaks = 0:3,
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = ~ scales::percent(abs(.x), accuracy = 1),
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.25),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_color_manual(
    values = c(
      "Accumulation" = "#b2182b",
      "Equivalent loss" = "#2166ac",
      "Net balance" = "#1a1a1a"
    )
  ) +
  scale_fill_manual(
    values = c("Accumulation" = "#d6604d", "Equivalent loss" = "#4393c3")
  ) +
  labs(
    title = "ST22 empirical cumulative tails with net balance",
    subtitle = "Black = accumulation prevalence minus equivalent-loss prevalence",
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Meter positions at threshold or greater",
    color = NULL,
    fill = NULL
  ) +
  guides(fill = "none") +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_mirrored_empirical_net_plot
```

### Animated empirical curves with maximum-net memory

The red, blue, and black curves show the current survey. At each absolute
sediment-depth threshold, the dashed green curve records the largest signed net
balance observed from Con3 through the current survey. It therefore acts as a
threshold-specific memory of the strongest accumulation-dominant result seen
so far.

```{r st22-animated-empirical-data}
st22_animation_frames <- st22_net_exceedance_profiles %>%
  distinct(survey_n, survey_label) %>%
  arrange(survey_n) %>%
  mutate(
    frame_label = paste0(
      str_replace(survey_label, "\\n", "  |  ")
    ),
    frame_label = factor(frame_label, levels = frame_label)
  )

st22_animation_tails <- st22_empirical_mirrored_profiles %>%
  left_join(st22_animation_frames, by = c("survey_n", "survey_label"))

st22_animation_net <- st22_net_exceedance_profiles %>%
  left_join(st22_animation_frames, by = c("survey_n", "survey_label"))

st22_animation_max_net <- st22_animation_net %>%
  arrange(threshold_cm, survey_n) %>%
  mutate(
    maximum_net_to_date = cummax(tail_balance),
    .by = threshold_cm
  ) %>%
  arrange(survey_n, threshold_cm)
```

```{r st22-static-animation-frames, fig.width=10, fig.height=7}
make_st22_empirical_frame <- function(frame_survey_n) {
  frame_info <- st22_animation_frames %>%
    filter(survey_n == frame_survey_n)
  frame_tails <- st22_animation_tails %>%
    filter(survey_n == frame_survey_n)
  frame_net <- st22_animation_net %>%
    filter(survey_n == frame_survey_n)
  frame_max_net <- st22_animation_max_net %>%
    filter(survey_n == frame_survey_n)

  ggplot() +
    geom_hline(yintercept = 0, color = "grey35", linewidth = 0.45) +
    geom_area(
      data = frame_tails,
      aes(
        x = threshold_cm,
        y = signed_fraction,
        fill = tail,
        group = tail
      ),
      alpha = 0.4,
      position = "identity"
    ) +
    geom_step(
      data = frame_tails,
      aes(
        x = threshold_cm,
        y = signed_fraction,
        color = tail,
        group = tail
      ),
      linewidth = 0.9,
      direction = "hv"
    ) +
    geom_step(
      data = frame_net,
      aes(x = threshold_cm, y = tail_balance),
      color = "white",
      linewidth = 2,
      direction = "hv"
    ) +
    geom_step(
      data = frame_net,
      aes(x = threshold_cm, y = tail_balance, color = "Current net balance"),
      linewidth = 1,
      direction = "hv"
    ) +
    geom_step(
      data = frame_max_net,
      aes(x = threshold_cm, y = maximum_net_to_date),
      color = "white",
      linewidth = 2.2,
      direction = "hv"
    ) +
    geom_step(
      data = frame_max_net,
      aes(
        x = threshold_cm,
        y = maximum_net_to_date,
        color = "Maximum net to date"
      ),
      linewidth = 1.1,
      linetype = "longdash",
      direction = "hv"
    ) +
    geom_vline(
      xintercept = c(0.15, 0.5, 1),
      color = "grey45",
      linewidth = 0.4,
      linetype = "dotted"
    ) +
    scale_x_continuous(
      limits = c(0, 3),
      breaks = seq(0, 3, by = 0.5),
      expand = expansion(mult = c(0, 0.01))
    ) +
    scale_y_continuous(
      labels = ~ scales::percent(abs(.x), accuracy = 1),
      limits = c(-1, 1),
      breaks = seq(-1, 1, by = 0.2),
      expand = expansion(mult = c(0.01, 0.01))
    ) +
    scale_color_manual(
      values = c(
        "Accumulation" = "#b2182b",
        "Equivalent loss" = "#2166ac",
        "Current net balance" = "#1a1a1a",
        "Maximum net to date" = "#1b7837"
      )
    ) +
    scale_fill_manual(
      values = c("Accumulation" = "#d6604d", "Equivalent loss" = "#4393c3")
    ) +
    labs(
      title = paste0("ST22 empirical cumulative tails: ", frame_info$frame_label),
      subtitle = paste(
        "Dashed green line remembers the maximum net balance reached at each",
        "threshold"
      ),
      x = "Absolute sediment-depth change threshold (cm)",
      y = "Meter positions at threshold or greater",
      color = NULL,
      fill = NULL
    ) +
    guides(
      fill = "none",
      color = guide_legend(
        override.aes = list(
          linetype = c("solid", "solid", "solid", "longdash"),
          linewidth = c(1, 1, 1, 1.1)
        )
      )
    ) +
    theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      panel.border = element_rect(
        color = "grey75",
        fill = NA,
        linewidth = 0.4
      ),
      legend.position = "bottom",
      plot.title = element_text(face = "bold"),
      plot.title.position = "plot"
    )
}

st22_frame_dir <- file.path("code", "figures", "st22_animation_frames")
dir.create(st22_frame_dir, recursive = TRUE, showWarnings = FALSE)

st22_frame_files <- st22_animation_frames %>%
  transmute(
    survey_n,
    frame_file = file.path(
      st22_frame_dir,
      sprintf("st22_frame_%02d.png", survey_n)
    )
  )

pwalk(
  st22_frame_files,
  function(survey_n, frame_file) {
    ggsave(
      frame_file,
      plot = make_st22_empirical_frame(survey_n),
      width = 10,
      height = 7,
      units = "in",
      dpi = 96,
      bg = "white"
    )
  }
)

stopifnot(all(file.exists(st22_frame_files$frame_file)))

st22_final_survey_n <- max(st22_animation_frames$survey_n)
st22_final_net <- st22_animation_net %>%
  filter(survey_n == st22_final_survey_n)
st22_final_max_net <- st22_animation_max_net %>%
  filter(survey_n == st22_final_survey_n)
st22_final_frame_label <- st22_animation_frames %>%
  filter(survey_n == st22_final_survey_n) %>%
  pull(frame_label) %>%
  as.character()

st22_final_summary_plot <- ggplot() +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.45) +
  geom_step(
    data = st22_final_net,
    aes(x = threshold_cm, y = tail_balance, color = "Most recent net balance"),
    linewidth = 1.15,
    direction = "hv"
  ) +
  geom_step(
    data = st22_final_max_net,
    aes(
      x = threshold_cm,
      y = maximum_net_to_date,
      color = "Maximum net over monitoring period"
    ),
    linewidth = 1.2,
    linetype = "longdash",
    direction = "hv"
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.4,
    linetype = "dotted"
  ) +
  scale_x_continuous(
    limits = c(0, 3),
    breaks = seq(0, 3, by = 0.5),
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.2),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_color_manual(
    values = c(
      "Most recent net balance" = "#1a1a1a",
      "Maximum net over monitoring period" = "#1b7837"
    )
  ) +
  labs(
    title = paste0("ST22 net-balance summary: ", st22_final_frame_label),
    subtitle = paste(
      "Black is the most recent net balance; dashed green is the maximum",
      "observed at each threshold"
    ),
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Accumulation minus equivalent-loss prevalence",
    color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey75", fill = NA, linewidth = 0.4),
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot"
  )

st22_final_summary_file <- file.path(
  st22_frame_dir,
  "st22_frame_final_summary.png"
)
ggsave(
  st22_final_summary_file,
  plot = st22_final_summary_plot,
  width = 10,
  height = 7,
  units = "in",
  dpi = 96,
  bg = "white"
)

knitr::include_graphics(
  normalizePath(
    st22_frame_files$frame_file[
      c(1, ceiling(nrow(st22_frame_files) / 2), nrow(st22_frame_files))
    ]
  ),
  rel_path = FALSE
)
```

```{r st22-interactive-empirical-player, echo=FALSE, results='asis'}
st22_player_id <- "st22-empirical-memory-player"
st22_player_frames <- c(
  map_chr(st22_frame_files$frame_file, knitr::image_uri),
  knitr::image_uri(st22_final_summary_file)
)
st22_player_frames_json <- jsonlite::toJSON(
  st22_player_frames,
  auto_unbox = TRUE
)

st22_player <- htmltools::tags$div(
  id = st22_player_id,
  class = "st22-player",
  htmltools::tags$style(htmltools::HTML(paste(
    ".st22-player { max-width: 960px; margin: 1rem auto; }",
    ".st22-player img { width: 100%; height: auto; display: block; cursor: pointer; border: 1px solid #d4d4d4; }",
    ".st22-player-controls { display: flex; align-items: center; gap: 0.65rem; margin-top: 0.55rem; }",
    ".st22-player button { border: 1px solid #777; background: #fff; color: #222; padding: 0.35rem 0.75rem; cursor: pointer; }",
    ".st22-player button:hover { background: #f0f0f0; }",
    ".st22-player input[type=range] { flex: 1; }",
    ".st22-player-status { min-width: 5.5rem; text-align: right; font-variant-numeric: tabular-nums; }",
    sep = "\n"
  ))),
  htmltools::tags$img(
    src = st22_player_frames[[1]],
    alt = "Animated ST22 empirical accumulation, loss, and net-balance curves"
  ),
  htmltools::tags$div(
    class = "st22-player-controls",
    htmltools::tags$button(type = "button", `data-action` = "play", "Play"),
    htmltools::tags$button(type = "button", `data-action` = "restart", "Restart"),
    htmltools::tags$input(
      type = "range",
      min = 0,
      max = length(st22_player_frames) - 1,
      value = 0,
      step = 1,
      `aria-label` = "Animation frame"
    ),
    htmltools::tags$span(
      class = "st22-player-status",
      paste0("1 / ", length(st22_player_frames))
    )
  ),
  htmltools::tags$script(htmltools::HTML(sprintf(paste0(
    "(function() {",
    "const root=document.getElementById('%s');",
    "const frames=%s;",
    "const image=root.querySelector('img');",
    "const playButton=root.querySelector('[data-action=play]');",
    "const restartButton=root.querySelector('[data-action=restart]');",
    "const slider=root.querySelector('input[type=range]');",
    "const status=root.querySelector('.st22-player-status');",
    "let index=0; let timer=null;",
    "function showFrame(nextIndex){",
    "index=Math.max(0,Math.min(nextIndex,frames.length-1));",
    "image.src=frames[index]; slider.value=index;",
    "status.textContent=(index+1)+' / '+frames.length;}",
    "function stop(){if(timer!==null)window.clearInterval(timer);",
    "timer=null; playButton.textContent=index===frames.length-1?'Play again':'Play';}",
    "function play(){if(timer!==null){stop();return;}",
    "if(index===frames.length-1)showFrame(0); playButton.textContent='Pause';",
    "timer=window.setInterval(function(){",
    "if(index>=frames.length-1){stop();return;} showFrame(index+1);",
    "if(index===frames.length-1)stop();},500);}",
    "playButton.addEventListener('click',play);",
    "image.addEventListener('click',play);",
    "restartButton.addEventListener('click',function(){stop();showFrame(0);});",
    "slider.addEventListener('input',function(){stop();showFrame(Number(this.value));});",
    "})();"
  ), st22_player_id, st22_player_frames_json)))
)

st22_player_html <- as.character(st22_player)
st22_player_html <- gsub(">[[:space:]]+<", "><", st22_player_html)
cat(st22_player_html)
```

```{r st22-animated-empirical-plot, fig.width=10, fig.height=7, eval=FALSE}
st22_animated_empirical_plot <- ggplot() +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.45) +
  geom_area(
    data = st22_animation_tails,
    aes(
      x = threshold_cm,
      y = signed_fraction,
      fill = tail,
      group = tail
    ),
    alpha = 0.4,
    position = "identity"
  ) +
  geom_step(
    data = st22_animation_tails,
    aes(
      x = threshold_cm,
      y = signed_fraction,
      color = tail,
      group = tail
    ),
    linewidth = 0.9,
    direction = "hv"
  ) +
  geom_step(
    data = st22_animation_net,
    aes(x = threshold_cm, y = tail_balance),
    color = "white",
    linewidth = 2,
    direction = "hv"
  ) +
  geom_step(
    data = st22_animation_net,
    aes(x = threshold_cm, y = tail_balance, color = "Current net balance"),
    linewidth = 1,
    direction = "hv"
  ) +
  geom_step(
    data = st22_animation_max_net,
    aes(x = threshold_cm, y = maximum_net_to_date),
    color = "white",
    linewidth = 2.2,
    direction = "hv"
  ) +
  geom_step(
    data = st22_animation_max_net,
    aes(
      x = threshold_cm,
      y = maximum_net_to_date,
      color = "Maximum net to date"
    ),
    linewidth = 1.1,
    linetype = "longdash",
    direction = "hv"
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.4,
    linetype = "dotted"
  ) +
  scale_x_continuous(
    limits = c(0, 3),
    breaks = seq(0, 3, by = 0.5),
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = ~ scales::percent(abs(.x), accuracy = 1),
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.2),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_color_manual(
    values = c(
      "Accumulation" = "#b2182b",
      "Equivalent loss" = "#2166ac",
      "Current net balance" = "#1a1a1a",
      "Maximum net to date" = "#1b7837"
    )
  ) +
  scale_fill_manual(
    values = c("Accumulation" = "#d6604d", "Equivalent loss" = "#4393c3")
  ) +
  labs(
    title = "ST22 empirical cumulative tails: {current_frame}",
    subtitle = paste(
      "Dashed green line remembers the maximum net balance reached at each",
      "threshold"
    ),
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Meter positions at threshold or greater",
    color = NULL,
    fill = NULL
  ) +
  guides(
    fill = "none",
    color = guide_legend(
      override.aes = list(
        linetype = c("solid", "solid", "solid", "longdash"),
        linewidth = c(1, 1, 1, 1.1)
      )
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey75", fill = NA, linewidth = 0.4),
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot"
  ) +
  transition_manual(frame_label)

st22_animation_file <- file.path(
  "code",
  "figures",
  "st22_mirrored_empirical_max_net.gif"
)
dir.create(dirname(st22_animation_file), recursive = TRUE, showWarnings = FALSE)

st22_animated_empirical_gif <- animate(
  st22_animated_empirical_plot,
  nframes = nrow(st22_animation_frames),
  fps = 2,
  end_pause = 5,
  width = 720,
  height = 504,
  res = 72,
  renderer = gifski_renderer(loop = TRUE)
)

anim_save(
  st22_animation_file,
  animation = st22_animated_empirical_gif
)

knitr::include_graphics(st22_animation_file)
```

### Mirrored smoothed curves with net balance

This version applies the same display to the kernel-density tail estimates.
Outlined red and blue points are the empirical tail prevalences at the three
regulatory thresholds; outlined black points are their empirical differences.

```{r st22-mirrored-smoothed-with-net, fig.width=11, fig.height=15}
st22_empirical_net_points <- st22_empirical_regulatory_points %>%
  transmute(
    date,
    survey,
    survey_n,
    survey_label,
    threshold_cm,
    tail_balance
  )

st22_mirrored_smoothed_net_plot <- ggplot() +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_area(
    data = st22_smoothed_mirrored_profiles,
    aes(
      x = threshold_cm,
      y = signed_fraction,
      fill = tail,
      group = tail
    ),
    alpha = 0.4,
    position = "identity"
  ) +
  geom_line(
    data = st22_smoothed_mirrored_profiles,
    aes(
      x = threshold_cm,
      y = signed_fraction,
      color = tail,
      group = tail
    ),
    linewidth = 0.65
  ) +
  geom_line(
    data = st22_smoothed_net_profiles,
    aes(x = threshold_cm, y = smoothed_tail_balance),
    color = "white",
    linewidth = 1.4
  ) +
  geom_line(
    data = st22_smoothed_net_profiles,
    aes(
      x = threshold_cm,
      y = smoothed_tail_balance,
      color = "Net balance"
    ),
    linewidth = 0.7
  ) +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey45",
    linewidth = 0.3,
    linetype = "dotted"
  ) +
  geom_point(
    data = st22_empirical_mirrored_points,
    aes(
      x = threshold_cm,
      y = signed_fraction,
      color = tail,
      fill = tail
    ),
    shape = 21,
    stroke = 0.5,
    size = 2.1
  ) +
  geom_point(
    data = st22_empirical_net_points,
    aes(x = threshold_cm, y = tail_balance, color = "Net balance"),
    fill = "white",
    shape = 21,
    stroke = 0.55,
    size = 2.1
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_x_continuous(
    limits = c(0, 3),
    breaks = 0:3,
    expand = expansion(mult = c(0, 0.01))
  ) +
  scale_y_continuous(
    labels = ~ scales::percent(abs(.x), accuracy = 1),
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.25),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_color_manual(
    values = c(
      "Accumulation" = "#b2182b",
      "Equivalent loss" = "#2166ac",
      "Net balance" = "#1a1a1a"
    )
  ) +
  scale_fill_manual(
    values = c("Accumulation" = "#d6604d", "Equivalent loss" = "#4393c3")
  ) +
  labs(
    title = "ST22 smoothed cumulative tails with net balance",
    subtitle = paste(
      "Black = accumulation prevalence minus equivalent-loss prevalence;",
      "outlined points are empirical values"
    ),
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Estimated meter positions at threshold or greater",
    color = NULL,
    fill = NULL
  ) +
  guides(fill = "none") +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_mirrored_smoothed_net_plot
```

### Cumulative net exceedance at regulatory thresholds

The curve above is already cumulative: its value at threshold (t) is
(P(\Delta \ge t) - P(\Delta \le -t)). This companion view extracts that
quantity at the three thresholds of interest and labels it directly. Thus, the
1 cm bar represents the prevalence of accumulation of 1 cm **or greater** minus
the prevalence of loss of 1 cm **or greater**.

```{r st22-cumulative-net-thresholds, fig.width=11, fig.height=15}
st22_cumulative_net_thresholds <-
  st22_alternative_exceedance_profiles %>%
  filter(threshold_cm %in% c(0.15, 0.5, 1)) %>%
  mutate(
    threshold_label = factor(
      threshold_cm,
      levels = c(0.15, 0.5, 1),
      labels = c("1.5 mm+", "5 mm+", "1 cm+")
    ),
    net_direction = if_else(
      tail_balance >= 0,
      "Net accumulation",
      "Net loss"
    ),
    label_y = tail_balance + if_else(tail_balance >= 0, 0.035, -0.035),
    label_vjust = if_else(tail_balance >= 0, 0, 1)
  )

st22_cumulative_net_threshold_plot <- ggplot(
  st22_cumulative_net_thresholds,
  aes(x = threshold_label, y = tail_balance, fill = net_direction)
) +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_col(width = 0.65) +
  geom_text(
    aes(
      y = label_y,
      label = scales::percent(tail_balance, accuracy = 1),
      vjust = label_vjust
    ),
    size = 2.5,
    show.legend = FALSE
  ) +
  facet_wrap(~ survey_label, ncol = 4) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(-0.75, 0.75),
    breaks = seq(-0.75, 0.75, by = 0.25),
    expand = expansion(mult = c(0.03, 0.03))
  ) +
  scale_fill_manual(
    values = c("Net accumulation" = "#d6604d", "Net loss" = "#4393c3")
  ) +
  labs(
    title = "ST22 cumulative net exceedance at regulatory thresholds",
    subtitle = paste(
      "Each value is accumulation prevalence minus equivalent loss prevalence",
      "at that threshold or greater"
    ),
    x = NULL,
    y = "Net prevalence at threshold or greater",
    fill = NULL
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold", size = 8.5),
    axis.text.x = element_text(size = 8),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_cumulative_net_threshold_plot
```

### Tail-balance heatmap

```{r st22-tail-balance-heatmap, fig.width=11, fig.height=11}
tail_balance_limit <- max(
  abs(st22_empirical_tail_profiles$tail_balance),
  na.rm = TRUE
)

st22_tail_balance_heatmap <- ggplot(
  st22_empirical_tail_profiles,
  aes(x = threshold_cm, y = survey_label, fill = tail_balance)
) +
  geom_raster() +
  geom_vline(
    xintercept = c(0.15, 0.5, 1),
    color = "grey15",
    linewidth = 0.35,
    linetype = "dotted"
  ) +
  facet_wrap(~ reference, ncol = 1) +
  scale_fill_gradient2(
    low = "#2166ac",
    mid = "white",
    high = "#b2182b",
    midpoint = 0,
    limits = c(-tail_balance_limit, tail_balance_limit),
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, tail_profile_max_cm, by = 1),
    expand = expansion(mult = c(0, 0))
  ) +
  labs(
    title = "ST22 tail balance across weeks and sediment-change thresholds",
    subtitle = "Red indicates more accumulation; blue indicates more equivalent loss",
    x = "Absolute sediment-depth change threshold (cm)",
    y = "Construction survey",
    fill = "Accumulation minus loss"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold"),
    plot.title.position = "plot"
  )

st22_tail_balance_heatmap
```

### Regulatory-threshold time series

```{r st22-threshold-time-series, fig.width=11, fig.height=5.5}
st22_threshold_time_series <- st22_reference_comparison %>%
  mutate(
    threshold_label = factor(
      threshold_label,
      levels = c("1.5 mm", "5 mm", "1 cm")
    ),
    reference = factor(
      reference,
      levels = c("Official baseline", "Baseline-Con1-Con2 average")
    )
  ) %>%
  ggplot(
    aes(
      x = date,
      y = tail_balance,
      color = reference,
      group = reference
    )
  ) +
  geom_hline(yintercept = 0, color = "grey35", linewidth = 0.4) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.8) +
  facet_wrap(~ threshold_label, nrow = 1) +
  scale_color_manual(
    values = c(
      "Official baseline" = "#555555",
      "Baseline-Con1-Con2 average" = "#b2182b"
    )
  ) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
  labs(
    title = "ST22 tail balance at the three sediment thresholds",
    subtitle = "Positive values indicate accumulation exceeds equivalent loss",
    x = NULL,
    y = "Accumulation fraction minus loss fraction",
    color = "Reference"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom",
    plot.title.position = "plot"
  )

st22_threshold_time_series
```

### Meter-position heatmap

```{r st22-meter-position-heatmap, fig.width=11, fig.height=12}
st22_spatial_color_limit <- max(
  abs(
    quantile(
      st22_reference_deltas_long$sediment_delta_cm,
      probs = c(0.02, 0.98),
      na.rm = TRUE
    )
  )
)

st22_meter_position_heatmap <- ggplot(
  st22_reference_deltas_long,
  aes(x = meter, y = survey_label, fill = sediment_delta_cm)
) +
  geom_tile() +
  geom_point(
    data = st22_reference_deltas_long %>%
      filter(sediment_delta_cm >= 1),
    shape = 21,
    fill = NA,
    color = "#67001f",
    size = 0.9,
    stroke = 0.35
  ) +
  facet_wrap(~ reference, ncol = 1) +
  scale_fill_gradient2(
    low = "#2166ac",
    mid = "white",
    high = "#b2182b",
    midpoint = 0,
    limits = c(-st22_spatial_color_limit, st22_spatial_color_limit),
    oob = scales::squish
  ) +
  scale_x_continuous(breaks = seq(0, 50, by = 5), expand = c(0, 0)) +
  labs(
    title = "ST22 sediment change through time and along the transect",
    subtitle = "Outlined cells meet or exceed 1 cm accumulation; color limits use the 2nd and 98th percentiles",
    x = "Meter position",
    y = "Construction survey",
    fill = "Sediment delta (cm)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold"),
    plot.title.position = "plot"
  )

st22_meter_position_heatmap
```
