---
title: "Sedimentation Baseline Robustness Analysis"
author: "Ross Cunning"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(patchwork)

project_dir <- if (basename(getwd()) == "code") dirname(getwd()) else getwd()
knitr::opts_knit$set(root.dir = project_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)

theme_set(
  theme_minimal(base_size = 10) +
    theme(
      panel.grid.minor = element_blank(),
      plot.title.position = "plot"
    )
)
```

# Purpose and interpretation

The official pre-construction survey appears higher than much of the subsequent
monitoring record at many transects. This report tests whether that pattern is
large, spatially consistent, and unusual relative to ordinary survey-to-survey
variation. It then compares alternative meter-specific reference formulations.

These analyses diagnose **representativeness**, not intent or data quality. A
high official baseline could reflect a real but transient sediment state,
seasonality, measurement differences, or natural redistribution. None of the
tests alone establishes the cause. Likewise, construction-period observations
are not automatically impact-free; alternatives using Con1 or Con2 must be
interpreted as sensitivity references rather than replacements for the
regulatory baseline unless their use is independently justified.

# Data preparation

```{r data-preparation}
sediment_dir <- "data/sedimentation_monitoring"

construction_files <- list.files(
  file.path(sediment_dir, "2_During-Construction", "Data"),
  pattern = "\\.xlsx$",
  recursive = TRUE,
  full.names = TRUE
)

construction_file_index <- tibble(source_file = construction_files) %>%
  mutate(
    source_file_name = basename(source_file),
    construction_survey_n = map_int(
      str_extract_all(source_file_name, "(?i)(?<=con)\\d+"),
      ~ if (length(.x) == 0) NA_integer_ else max(as.integer(.x))
    )
  ) %>%
  filter(!is.na(construction_survey_n)) %>%
  arrange(desc(construction_survey_n), desc(file.info(source_file)$mtime))

latest_construction_file <- construction_file_index %>%
  slice(1) %>%
  pull(source_file)

sediment_raw <- read_excel(
  latest_construction_file,
  sheet = "SedimentDepth_RawData"
) %>%
  clean_names() %>%
  mutate(
    date = as.Date(date),
    survey = str_squish(survey),
    survey_n = if_else(
      survey == "Baseline",
      0L,
      as.integer(str_extract(survey, "\\d+"))
    ),
    transect = str_replace_all(str_squish(transect), "\\s+", "_"),
    meter = as.integer(meter),
    meter_mean_sediment_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    )
  )

transect_survey_coverage <- sediment_raw %>%
  summarize(
    n_meter_positions = n_distinct(meter),
    n_surveys = n_distinct(survey),
    latest_survey_n = max(survey_n),
    has_baseline = any(survey_n == 0),
    has_con1 = any(survey_n == 1),
    has_con2 = any(survey_n == 2),
    .by = transect
  ) %>%
  arrange(transect)

analysis_transects <- transect_survey_coverage %>%
  filter(has_baseline, has_con1, has_con2, latest_survey_n >= 3) %>%
  pull(transect)

baseline_only_transects <- transect_survey_coverage %>%
  filter(!transect %in% analysis_transects) %>%
  pull(transect)

sediment_analysis <- sediment_raw %>%
  filter(
    transect %in% analysis_transects,
    meter <= 50
  ) %>%
  mutate(
    predicted_impact_class = case_when(
      transect %in% c("ST22", "ST31") ~ "1 cm predicted",
      transect %in% c("ST28", "ST5") ~ "1.5 mm predicted",
      TRUE ~ "0 cm predicted"
    ),
    predicted_impact_class = factor(
      predicted_impact_class,
      levels = c("1 cm predicted", "1.5 mm predicted", "0 cm predicted")
    )
  )

analysis_coverage <- sediment_analysis %>%
  summarize(
    n_meter_positions = n_distinct(meter),
    first_date = min(date),
    last_date = max(date),
    latest_survey_n = max(survey_n),
    .by = transect
  ) %>%
  arrange(transect)

stopifnot(
  length(analysis_transects) == 13,
  all(analysis_coverage$n_meter_positions == 51),
  all(!is.na(sediment_analysis$meter_mean_sediment_cm))
)

analysis_coverage
```

The analysis includes `r length(analysis_transects)` transects with Baseline,
Con1, Con2, and later construction monitoring. For ST1 and ST38, the analysis
uses meters 0--50 because Con1 and Con2 do not cover their full 150 m extent.
The baseline-only transects (`r paste(baseline_only_transects, collapse = ", ")`)
cannot be used to compare reference formulations through time.

# Is the official baseline unusually high?

## Transect-wide survey means

```{r transect-survey-means, fig.width=12, fig.height=8}
transect_survey_means <- sediment_analysis %>%
  summarize(
    transect_mean_cm = mean(meter_mean_sediment_cm),
    transect_median_cm = median(meter_mean_sediment_cm),
    .by = c(transect, date, survey, survey_n)
  ) %>%
  arrange(transect, survey_n)

early_transect_means <- transect_survey_means %>%
  filter(survey_n <= 6) %>%
  mutate(
    survey_label = factor(
      survey,
      levels = c("Baseline", paste0("Con", 1:6))
    )
  )

ggplot(
  early_transect_means,
  aes(x = survey_label, y = transect_mean_cm, group = transect)
) +
  geom_line(color = "grey65", linewidth = 0.5) +
  geom_point(
    aes(color = survey_label == "Baseline"),
    size = 1.8
  ) +
  facet_wrap(~ transect, scales = "free_y", ncol = 4) +
  scale_color_manual(
    values = c(`TRUE` = "#b2182b", `FALSE` = "#333333"),
    guide = "none"
  ) +
  labs(
    title = "Early transect-wide mean sediment depth",
    subtitle = "Official Baseline is highlighted in red; free y-scales show within-transect changes",
    x = NULL,
    y = "Mean sediment depth (cm)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold")
  )
```

```{r all-transect-survey-means, fig.width=12, fig.height=9}
ggplot(
  transect_survey_means,
  aes(x = date, y = transect_mean_cm, group = transect)
) +
  geom_line(color = "grey55", linewidth = 0.55) +
  geom_point(
    aes(color = survey == "Baseline"),
    size = 1.5
  ) +
  facet_wrap(~ transect, scales = "free_y", ncol = 4) +
  scale_color_manual(
    values = c(`TRUE` = "#b2182b", `FALSE` = "#333333"),
    guide = "none"
  ) +
  scale_x_date(
    date_breaks = "2 months",
    date_labels = "%b\n%Y",
    expand = expansion(mult = c(0.02, 0.03))
  ) +
  labs(
    title = "Transect-wide mean sediment depth across the complete monitoring record",
    subtitle = "Official Baseline is highlighted in red; free y-scales show within-transect changes",
    x = NULL,
    y = "Mean sediment depth (cm)"
  ) +
  theme(
    axis.text.x = element_text(size = 7),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold")
  )
```

## Meter-matched elevation above Con1 and Con2

Comparisons are paired by transect and meter. Positive differences mean that
the official baseline depth was greater than the comparison survey at the same
location.

```{r baseline-paired-differences}
early_meter_wide <- sediment_analysis %>%
  filter(survey_n %in% 0:2) %>%
  select(transect, meter, survey_n, meter_mean_sediment_cm) %>%
  mutate(
    survey_key = recode(
      as.character(survey_n),
      `0` = "baseline_cm",
      `1` = "con1_cm",
      `2` = "con2_cm"
    )
  ) %>%
  select(-survey_n) %>%
  pivot_wider(names_from = survey_key, values_from = meter_mean_sediment_cm) %>%
  mutate(
    baseline_minus_con1_cm = baseline_cm - con1_cm,
    baseline_minus_con2_cm = baseline_cm - con2_cm,
    baseline_minus_early_mean_cm = baseline_cm - (con1_cm + con2_cm) / 2
  )

paired_long <- early_meter_wide %>%
  pivot_longer(
    c(baseline_minus_con1_cm, baseline_minus_con2_cm),
    names_to = "comparison",
    values_to = "baseline_excess_cm"
  ) %>%
  mutate(
    comparison = recode(
      comparison,
      baseline_minus_con1_cm = "Baseline - Con1",
      baseline_minus_con2_cm = "Baseline - Con2"
    )
  )

paired_summary <- paired_long %>%
  summarize(
    n_meter_positions = n(),
    mean_excess_cm = mean(baseline_excess_cm),
    median_excess_cm = median(baseline_excess_cm),
    pct_baseline_higher = mean(baseline_excess_cm > 0),
    .by = c(transect, comparison)
  )

paired_summary
```

```{r paired-difference-plot, fig.width=12, fig.height=8}
ggplot(
  paired_long,
  aes(x = comparison, y = baseline_excess_cm, fill = comparison)
) +
  geom_hline(yintercept = 0, color = "grey30", linewidth = 0.4) +
  geom_boxplot(width = 0.65, outlier.size = 0.7) +
  facet_wrap(~ transect, ncol = 4) +
  scale_fill_manual(values = c(
    "Baseline - Con1" = "#b2182b",
    "Baseline - Con2" = "#ef8a62"
  )) +
  labs(
    title = "Meter-matched official baseline elevation",
    subtitle = "Positive values indicate that official Baseline exceeds the early construction survey",
    x = NULL,
    y = "Official baseline excess (cm)",
    fill = NULL
  ) +
  guides(fill = "none") +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.3),
    strip.text = element_text(face = "bold")
  )
```

## Paired tests by transect

The Wilcoxon signed-rank test evaluates whether paired differences are centered
at zero. The exact binomial sign test asks only whether positive differences
outnumber negative differences, making it less sensitive to unusually large
values. False-discovery-rate-adjusted p-values account for testing multiple
transects.

```{r paired-tests}
safe_wilcox_p <- function(x) {
  tryCatch(
    wilcox.test(x, mu = 0, exact = FALSE)$p.value,
    error = function(e) NA_real_
  )
}

safe_sign_p <- function(x) {
  x <- x[x != 0 & !is.na(x)]
  if (length(x) == 0) return(NA_real_)
  binom.test(sum(x > 0), length(x), p = 0.5)$p.value
}

paired_tests <- paired_long %>%
  summarize(
    n = n(),
    median_excess_cm = median(baseline_excess_cm),
    pct_baseline_higher = mean(baseline_excess_cm > 0),
    wilcoxon_p = safe_wilcox_p(baseline_excess_cm),
    sign_test_p = safe_sign_p(baseline_excess_cm),
    .by = c(transect, comparison)
  ) %>%
  mutate(
    wilcoxon_fdr = p.adjust(wilcoxon_p, method = "BH"),
    sign_test_fdr = p.adjust(sign_test_p, method = "BH"),
    .by = comparison
  ) %>%
  arrange(comparison, transect)

paired_tests
```

## Transect-level robustness tests

Adjacent meter positions are not independent ecological replicates. The tests
below therefore reduce each transect to its mean paired difference and treat the
13 transects as the independent analysis units. These are the primary tests for
whether official-baseline elevation is consistent across the monitoring area;
the meter-level tests above describe within-transect spatial consistency.

```{r transect-level-tests}
transect_level_differences <- paired_long %>%
  summarize(
    mean_baseline_excess_cm = mean(baseline_excess_cm),
    median_baseline_excess_cm = median(baseline_excess_cm),
    .by = c(transect, comparison)
  )

transect_level_tests <- transect_level_differences %>%
  summarize(
    n_transects = n(),
    mean_of_transect_means_cm = mean(mean_baseline_excess_cm),
    median_of_transect_means_cm = median(mean_baseline_excess_cm),
    pct_transects_baseline_higher = mean(mean_baseline_excess_cm > 0),
    wilcoxon_p = safe_wilcox_p(mean_baseline_excess_cm),
    sign_test_p = safe_sign_p(mean_baseline_excess_cm),
    .by = comparison
  )

transect_level_tests
```

## Is Baseline to Con1 an unusual transition?

This comparison uses transect-wide means so each transect contributes equally.
The signed Baseline-to-Con1 shift is compared with each subsequent early shift
through Con6. A strongly negative first shift that is larger than ordinary
construction-to-construction changes supports the interpretation that Baseline
captured an unusual state.

```{r early-transition-analysis, fig.width=11, fig.height=6.5}
early_transitions <- transect_survey_means %>%
  filter(survey_n <= 6) %>%
  arrange(transect, survey_n) %>%
  mutate(
    previous_survey_n = lag(survey_n),
    transition_delta_cm = transect_mean_cm - lag(transect_mean_cm),
    transition = case_when(
      survey_n == 1 ~ "Baseline to Con1",
      TRUE ~ paste0("Con", previous_survey_n, " to Con", survey_n)
    ),
    .by = transect
  ) %>%
  filter(!is.na(transition_delta_cm)) %>%
  mutate(
    transition = factor(
      transition,
      levels = c(
        "Baseline to Con1",
        "Con1 to Con2",
        "Con2 to Con3",
        "Con3 to Con4",
        "Con4 to Con5",
        "Con5 to Con6"
      )
    )
  )

transition_summary <- early_transitions %>%
  summarize(
    n_transects = n(),
    mean_shift_cm = mean(transition_delta_cm),
    median_shift_cm = median(transition_delta_cm),
    mean_absolute_shift_cm = mean(abs(transition_delta_cm)),
    .by = transition
  )

transition_by_transect <- early_transitions %>%
  summarize(
    baseline_to_con1_cm = first(
      transition_delta_cm[transition == "Baseline to Con1"]
    ),
    typical_later_shift_cm = median(
      transition_delta_cm[transition != "Baseline to Con1"]
    ),
    baseline_absolute_shift_cm = abs(baseline_to_con1_cm),
    typical_later_absolute_shift_cm = median(
      abs(transition_delta_cm[transition != "Baseline to Con1"])
    ),
    .by = transect
  )

transition_test <- transition_by_transect %>%
  summarize(
    paired_absolute_shift_p = wilcox.test(
      baseline_absolute_shift_cm,
      typical_later_absolute_shift_cm,
      alternative = "greater",
      paired = TRUE,
      exact = FALSE
    )$p.value,
    paired_signed_shift_p = wilcox.test(
      baseline_to_con1_cm,
      typical_later_shift_cm,
      alternative = "less",
      paired = TRUE,
      exact = FALSE
    )$p.value
  )

ggplot(
  early_transitions,
  aes(x = transition, y = transition_delta_cm)
) +
  geom_hline(yintercept = 0, color = "grey30", linewidth = 0.4) +
  geom_boxplot(
    aes(fill = transition == "Baseline to Con1"),
    width = 0.65,
    outlier.shape = NA
  ) +
  geom_jitter(width = 0.1, height = 0, size = 1.6, alpha = 0.75) +
  scale_fill_manual(
    values = c(`TRUE` = "#b2182b", `FALSE` = "grey75"),
    guide = "none"
  ) +
  labs(
    title = "Early changes in transect-wide mean sediment depth",
    subtitle = "Each point is one transect; negative shifts indicate declining sediment depth",
    x = NULL,
    y = "Change from previous survey (cm)"
  ) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

transition_summary
transition_test
```

# When does a new sediment regime appear?

Later surveys may improve the estimate of ordinary background conditions, but
only until the system begins to depart from that regime. A defensible expanded
reference window therefore requires two findings: adding another survey should
improve prediction of the next survey, and it should not coincide with a new,
spatially selective or sustained shift that may reflect project-related flux.

## Complete survey-to-survey transition scan

For every transect and meter, each observation is compared with the immediately
preceding available survey. The transect mean is used as the independent unit
for regional tests; meter-level summaries retain information about whether a
shift is spatially widespread along a transect.

```{r full-transition-scan}
meter_transitions <- sediment_analysis %>%
  arrange(transect, meter, survey_n) %>%
  mutate(
    previous_survey = lag(survey),
    previous_survey_n = lag(survey_n),
    previous_depth_cm = lag(meter_mean_sediment_cm),
    transition_delta_cm = meter_mean_sediment_cm - previous_depth_cm,
    .by = c(transect, meter)
  ) %>%
  filter(!is.na(previous_depth_cm)) %>%
  mutate(
    transition_label = paste0(previous_survey, " to ", survey)
  )

transect_transitions <- meter_transitions %>%
  summarize(
    mean_transition_cm = mean(transition_delta_cm),
    median_transition_cm = median(transition_delta_cm),
    mean_absolute_meter_change_cm = mean(abs(transition_delta_cm)),
    pct_meter_positions_increasing = mean(transition_delta_cm > 0),
    meter_wilcoxon_p = safe_wilcox_p(transition_delta_cm),
    .by = c(
      transect,
      predicted_impact_class,
      date,
      survey,
      survey_n,
      previous_survey,
      previous_survey_n,
      transition_label
    )
  ) %>%
  mutate(
    meter_wilcoxon_fdr = p.adjust(meter_wilcoxon_p, method = "BH"),
    .by = transect
  )

regional_transition_tests <- transect_transitions %>%
  summarize(
    n_transects = n(),
    regional_mean_shift_cm = mean(mean_transition_cm),
    regional_median_shift_cm = median(mean_transition_cm),
    pct_transects_increasing = mean(mean_transition_cm > 0),
    transect_wilcoxon_p = safe_wilcox_p(mean_transition_cm),
    transect_sign_p = safe_sign_p(mean_transition_cm),
    .by = c(date, survey, survey_n, previous_survey, transition_label)
  ) %>%
  arrange(survey_n) %>%
  mutate(
    transect_wilcoxon_fdr = p.adjust(transect_wilcoxon_p, method = "BH"),
    transect_sign_fdr = p.adjust(transect_sign_p, method = "BH")
  )

regional_transition_tests
```

```{r transition-heatmap, fig.width=12, fig.height=7}
transition_color_limit <- max(abs(quantile(
  transect_transitions$mean_transition_cm,
  probs = c(0.02, 0.98),
  na.rm = TRUE
)))

ggplot(
  transect_transitions,
  aes(
    x = fct_reorder(survey, survey_n),
    y = fct_rev(factor(transect)),
    fill = mean_transition_cm
  )
) +
  geom_tile(color = "white", linewidth = 0.15) +
  geom_point(
    data = transect_transitions %>%
      filter(meter_wilcoxon_fdr < 0.05),
    shape = 21,
    fill = NA,
    color = "grey15",
    size = 1.2,
    stroke = 0.35
  ) +
  scale_fill_gradient2(
    low = "#2166ac",
    mid = "white",
    high = "#b2182b",
    midpoint = 0,
    limits = c(-transition_color_limit, transition_color_limit),
    oob = scales::squish
  ) +
  labs(
    title = "Survey-to-survey shifts in transect mean sediment depth",
    subtitle = "Blue indicates decline and red indicates increase; outlines mark meter-level FDR < 0.05",
    x = "Ending survey",
    y = "Transect",
    fill = "Mean shift (cm)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Deviations from the robust three-survey early state

The meter-specific Baseline-Con1-Con2 median is used here only as a diagnostic
anchor. A project-related pattern would be more plausible if positive deviations
emerge preferentially or earlier in transects with modeled exposure, rather
than appearing synchronously across all transects.

```{r deviation-by-impact-class, fig.width=12, fig.height=8}
robust_early_reference <- sediment_analysis %>%
  filter(survey_n %in% 0:2) %>%
  summarize(
    robust_early_reference_cm = median(meter_mean_sediment_cm),
    .by = c(transect, meter)
  )

survey_deviations <- sediment_analysis %>%
  filter(survey_n >= 3) %>%
  left_join(robust_early_reference, by = c("transect", "meter")) %>%
  mutate(
    deviation_cm = meter_mean_sediment_cm - robust_early_reference_cm
  ) %>%
  summarize(
    mean_deviation_cm = mean(deviation_cm),
    median_deviation_cm = median(deviation_cm),
    pct_meter_positions_increasing = mean(deviation_cm > 0),
    .by = c(
      transect,
      predicted_impact_class,
      date,
      survey,
      survey_n
    )
  )

ggplot(
  survey_deviations,
  aes(
    x = survey_n,
    y = mean_deviation_cm,
    color = predicted_impact_class,
    group = transect
  )
) +
  geom_hline(yintercept = 0, color = "grey30", linewidth = 0.4) +
  geom_line(linewidth = 0.65, alpha = 0.75) +
  geom_point(size = 1.3) +
  facet_wrap(~ predicted_impact_class, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c(
    "1 cm predicted" = "#b2182b",
    "1.5 mm predicted" = "#ef8a62",
    "0 cm predicted" = "#2166ac"
  )) +
  scale_x_continuous(breaks = sort(unique(survey_deviations$survey_n))) +
  labs(
    title = "Deviation from the robust early state by modeled impact class",
    subtitle = "Each line is one transect; selective positive departures are more compatible with project-related flux",
    x = "Construction survey number",
    y = "Mean deviation from early median reference (cm)",
    color = "Modeled sedimentation"
  ) +
  guides(color = "none") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

## Sequential reference-window evaluation

Candidate windows are extended one survey at a time from Con1 through Con6.
For each endpoint, both a mean and a median meter-specific reference are used to
predict the next survey. A useful additional survey should reduce absolute
next-survey error without causing a large systematic drift in the reference.

```{r sequential-reference-windows}
reference_window_ends <- 1:6

sequential_references <- crossing(
  sediment_analysis %>% distinct(transect, meter),
  window_end = reference_window_ends,
  estimator = c("Mean", "Median")
) %>%
  left_join(
    sediment_analysis %>%
      select(transect, meter, survey_n, meter_mean_sediment_cm),
    by = c("transect", "meter"),
    relationship = "many-to-many"
  ) %>%
  filter(survey_n <= window_end) %>%
  summarize(
    reference_cm = if_else(
      first(estimator) == "Mean",
      mean(meter_mean_sediment_cm),
      median(meter_mean_sediment_cm)
    ),
    n_reference_surveys = n(),
    .by = c(transect, meter, window_end, estimator)
  )

sequential_evaluation <- sequential_references %>%
  left_join(
    sediment_analysis %>%
      select(
        transect,
        meter,
        next_survey_n = survey_n,
        next_depth_cm = meter_mean_sediment_cm
      ),
    by = c("transect", "meter"),
    relationship = "many-to-many"
  ) %>%
  filter(next_survey_n == window_end + 1) %>%
  mutate(next_delta_cm = next_depth_cm - reference_cm) %>%
  summarize(
    transect_next_bias_cm = mean(next_delta_cm),
    transect_next_mae_cm = mean(abs(next_delta_cm)),
    .by = c(transect, window_end, estimator)
  )

sequential_reference_drift <- sequential_references %>%
  arrange(transect, meter, estimator, window_end) %>%
  mutate(
    reference_drift_cm = reference_cm - lag(reference_cm),
    .by = c(transect, meter, estimator)
  ) %>%
  filter(!is.na(reference_drift_cm)) %>%
  summarize(
    transect_mean_reference_drift_cm = mean(reference_drift_cm),
    transect_absolute_reference_drift_cm = mean(abs(reference_drift_cm)),
    .by = c(transect, window_end, estimator)
  )

sequential_summary <- sequential_evaluation %>%
  summarize(
    median_next_bias_cm = median(transect_next_bias_cm),
    mean_absolute_transect_bias_cm = mean(abs(transect_next_bias_cm)),
    median_next_mae_cm = median(transect_next_mae_cm),
    .by = c(window_end, estimator)
  ) %>%
  left_join(
    sequential_reference_drift %>%
      summarize(
        median_absolute_reference_drift_cm = median(
          transect_absolute_reference_drift_cm
        ),
        .by = c(window_end, estimator)
      ),
    by = c("window_end", "estimator")
  )

sequential_summary
```

```{r sequential-window-plot, fig.width=11, fig.height=8}
window_bias_plot <- ggplot(
  sequential_summary,
  aes(
    x = window_end,
    y = mean_absolute_transect_bias_cm,
    color = estimator
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_x_continuous(
    breaks = reference_window_ends,
    labels = paste0("Through Con", reference_window_ends)
  ) +
  scale_color_manual(values = c("Mean" = "#b2182b", "Median" = "#2166ac")) +
  labs(
    title = "Prediction of the next construction survey",
    x = NULL,
    y = "Mean absolute transect bias (cm)",
    color = "Reference estimator"
  ) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

window_drift_plot <- ggplot(
  sequential_summary,
  aes(
    x = window_end,
    y = median_absolute_reference_drift_cm,
    color = estimator
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_x_continuous(
    breaks = reference_window_ends,
    labels = paste0("Add Con", reference_window_ends)
  ) +
  scale_color_manual(values = c("Mean" = "#b2182b", "Median" = "#2166ac")) +
  labs(
    title = "Reference movement when each survey is added",
    x = NULL,
    y = "Median absolute reference drift (cm)",
    color = "Reference estimator"
  ) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

window_bias_plot / window_drift_plot +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

The sequential analysis is intended to identify plausible reference windows,
not mechanically choose the window with the smallest prediction error. The
preferred endpoint should precede any sustained or impact-class-specific
departure in the transition and deviation plots. Surveys after that point are
outcomes to be detected, not observations to be absorbed into the reference.

# Candidate alternative references

Four references are evaluated at each transect and meter:

1. **Official Baseline**: the regulatory pre-construction observation.
2. **Baseline-Con1-Con2 mean**: the current localized-analysis sensitivity
   reference; uses all three surveys but is sensitive to one high value.
3. **Baseline-Con1-Con2 median**: robust to one unusually high or low survey.
   If Baseline is the highest of the three, this selects the higher of Con1 and
   Con2 and is therefore conservative relative to their mean.
4. **Con1-Con2 mean**: excludes the official baseline and provides the clearest
   estimate of the early construction state, but has the greatest risk of
   absorbing genuine early project-related deposition.

```{r candidate-references}
candidate_references <- early_meter_wide %>%
  transmute(
    transect,
    meter,
    `Official Baseline` = baseline_cm,
    `Baseline-Con1-Con2 mean` = (baseline_cm + con1_cm + con2_cm) / 3,
    `Baseline-Con1-Con2 median` = pmap_dbl(
      list(baseline_cm, con1_cm, con2_cm),
      ~ median(c(...))
    ),
    `Con1-Con2 mean` = (con1_cm + con2_cm) / 2
  ) %>%
  pivot_longer(
    -c(transect, meter),
    names_to = "reference_method",
    values_to = "reference_cm"
  ) %>%
  mutate(
    reference_method = factor(
      reference_method,
      levels = c(
        "Official Baseline",
        "Baseline-Con1-Con2 mean",
        "Baseline-Con1-Con2 median",
        "Con1-Con2 mean"
      )
    )
  )

candidate_contrasts <- candidate_references %>%
  select(transect, meter, reference_method, reference_cm) %>%
  pivot_wider(names_from = reference_method, values_from = reference_cm) %>%
  mutate(
    mean3_reduction_cm =
      `Official Baseline` - `Baseline-Con1-Con2 mean`,
    median3_reduction_cm =
      `Official Baseline` - `Baseline-Con1-Con2 median`,
    early2_reduction_cm =
      `Official Baseline` - `Con1-Con2 mean`
  )

candidate_contrast_summary <- candidate_contrasts %>%
  summarize(
    mean3_reduction_cm = mean(mean3_reduction_cm),
    median3_reduction_cm = mean(median3_reduction_cm),
    early2_reduction_cm = mean(early2_reduction_cm),
    pct_median3_below_official = mean(median3_reduction_cm > 0),
    .by = transect
  )

candidate_contrast_summary
```

## How often is each early survey selected by the median?

```{r median-source}
median_source <- early_meter_wide %>%
  mutate(
    median_reference_cm = pmap_dbl(
      list(baseline_cm, con1_cm, con2_cm),
      ~ median(c(...))
    ),
    median_source = case_when(
      near(median_reference_cm, baseline_cm) ~ "Official Baseline",
      near(median_reference_cm, con1_cm) ~ "Con1",
      near(median_reference_cm, con2_cm) ~ "Con2",
      TRUE ~ "Tie"
    )
  ) %>%
  count(transect, median_source) %>%
  group_by(transect) %>%
  mutate(fraction = n / sum(n)) %>%
  ungroup()

ggplot(
  median_source,
  aes(x = transect, y = fraction, fill = median_source)
) +
  geom_col(width = 0.75) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_fill_manual(values = c(
    "Official Baseline" = "#555555",
    "Con1" = "#d6604d",
    "Con2" = "#4393c3",
    "Tie" = "#cccccc"
  )) +
  labs(
    title = "Source survey selected by the three-survey median",
    subtitle = "The median retains the middle observation independently at each meter",
    x = "Transect",
    y = "Meter positions",
    fill = "Selected survey"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Holdout evaluation using Con3-Con6

Con3-Con6 are not used to construct any candidate. Their deltas provide a common
early holdout period for evaluating how strongly each reference systematically
centers later observations above or below zero. This is a stability diagnostic,
not proof that the holdout surveys are impact-free.

```{r holdout-evaluation}
holdout_deltas <- sediment_analysis %>%
  filter(survey_n %in% 3:6) %>%
  select(transect, meter, date, survey, survey_n, meter_mean_sediment_cm) %>%
  left_join(
    candidate_references,
    by = c("transect", "meter"),
    relationship = "many-to-many"
  ) %>%
  mutate(delta_cm = meter_mean_sediment_cm - reference_cm)

holdout_by_transect <- holdout_deltas %>%
  summarize(
    mean_bias_cm = mean(delta_cm),
    median_bias_cm = median(delta_cm),
    mean_absolute_delta_cm = mean(abs(delta_cm)),
    pct_below_reference = mean(delta_cm < 0),
    .by = c(transect, reference_method)
  )

holdout_overall <- holdout_by_transect %>%
  summarize(
    median_transect_bias_cm = median(mean_bias_cm),
    mean_absolute_transect_bias_cm = mean(abs(mean_bias_cm)),
    median_meter_absolute_delta_cm = median(mean_absolute_delta_cm),
    median_pct_below_reference = median(pct_below_reference),
    .by = reference_method
  ) %>%
  arrange(mean_absolute_transect_bias_cm)

holdout_overall
```

```{r holdout-bias-plot, fig.width=11, fig.height=7}
ggplot(
  holdout_by_transect,
  aes(x = reference_method, y = mean_bias_cm, color = reference_method)
) +
  geom_hline(yintercept = 0, color = "grey30", linewidth = 0.4) +
  geom_boxplot(width = 0.55, outlier.shape = NA, color = "grey45") +
  geom_point(
    aes(group = transect),
    position = position_jitter(width = 0.08, height = 0),
    size = 2,
    alpha = 0.85
  ) +
  scale_color_manual(values = c(
    "Official Baseline" = "#555555",
    "Baseline-Con1-Con2 mean" = "#b2182b",
    "Baseline-Con1-Con2 median" = "#ef8a62",
    "Con1-Con2 mean" = "#2166ac"
  )) +
  labs(
    title = "Reference bias during the Con3-Con6 holdout period",
    subtitle = "Each point is one transect; values below zero indicate that holdout depths are below the reference",
    x = NULL,
    y = "Mean holdout delta (cm)",
    color = "Reference"
  ) +
  guides(color = guide_legend(nrow = 2)) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )
```

# Bootstrap uncertainty for reference differences

Meters are resampled within each transect. The intervals quantify uncertainty
in the average difference between the official baseline and each alternative;
they do not account for temporal autocorrelation or prove that one reference is
causally correct.

```{r bootstrap-reference-differences}
set.seed(240917)

bootstrap_mean_ci <- function(x, n_boot = 4000) {
  boot_means <- replicate(
    n_boot,
    mean(sample(x, size = length(x), replace = TRUE))
  )
  tibble(
    estimate = mean(x),
    lower_95 = quantile(boot_means, 0.025),
    upper_95 = quantile(boot_means, 0.975)
  )
}

bootstrap_differences <- candidate_contrasts %>%
  select(
    transect,
    `Baseline-Con1-Con2 mean` = mean3_reduction_cm,
    `Baseline-Con1-Con2 median` = median3_reduction_cm,
    `Con1-Con2 mean` = early2_reduction_cm
  ) %>%
  pivot_longer(
    -transect,
    names_to = "alternative_reference",
    values_to = "official_minus_alternative_cm"
  ) %>%
  nest(.by = c(transect, alternative_reference)) %>%
  mutate(
    ci = map(
      data,
      ~ bootstrap_mean_ci(.x$official_minus_alternative_cm)
    )
  ) %>%
  select(-data) %>%
  unnest(ci)

bootstrap_differences
```

```{r bootstrap-plot, fig.width=11, fig.height=8}
ggplot(
  bootstrap_differences,
  aes(
    x = estimate,
    y = fct_rev(factor(transect)),
    xmin = lower_95,
    xmax = upper_95,
    color = alternative_reference
  )
) +
  geom_vline(xintercept = 0, color = "grey30", linewidth = 0.4) +
  geom_errorbarh(
    position = position_dodge(width = 0.6),
    height = 0.15,
    linewidth = 0.5
  ) +
  geom_point(position = position_dodge(width = 0.6), size = 1.8) +
  scale_color_manual(values = c(
    "Baseline-Con1-Con2 mean" = "#b2182b",
    "Baseline-Con1-Con2 median" = "#ef8a62",
    "Con1-Con2 mean" = "#2166ac"
  )) +
  labs(
    title = "Official baseline elevation above candidate alternatives",
    subtitle = "Mean meter-matched difference with within-transect bootstrap 95% intervals",
    x = "Official baseline minus alternative reference (cm)",
    y = "Transect",
    color = "Alternative reference"
  ) +
  theme(legend.position = "bottom")
```

# Regulatory-threshold sensitivity

The practical consequence of reference choice is evaluated using cumulative net
exceedance at 1.5 mm, 5 mm, and 1 cm. Positive values mean that accumulation
prevalence exceeds equivalent-loss prevalence.

```{r threshold-sensitivity}
regulatory_thresholds <- tibble(
  threshold_cm = c(0.15, 0.5, 1),
  threshold_label = factor(
    c("1.5 mm", "5 mm", "1 cm"),
    levels = c("1.5 mm", "5 mm", "1 cm")
  )
)

candidate_all_deltas <- sediment_analysis %>%
  filter(survey_n >= 3) %>%
  select(transect, meter, date, survey, survey_n, meter_mean_sediment_cm) %>%
  left_join(
    candidate_references,
    by = c("transect", "meter"),
    relationship = "many-to-many"
  ) %>%
  mutate(delta_cm = meter_mean_sediment_cm - reference_cm)

threshold_sensitivity <- crossing(
  candidate_all_deltas,
  regulatory_thresholds
) %>%
  summarize(
    accumulation_fraction = mean(delta_cm >= threshold_cm),
    loss_fraction = mean(delta_cm <= -threshold_cm),
    net_exceedance = accumulation_fraction - loss_fraction,
    .by = c(
      transect,
      date,
      survey,
      survey_n,
      reference_method,
      threshold_cm,
      threshold_label
    )
  )

threshold_sensitivity_summary <- threshold_sensitivity %>%
  summarize(
    median_net_exceedance = median(net_exceedance),
    maximum_net_exceedance = max(net_exceedance),
    pct_surveys_positive = mean(net_exceedance > 0),
    .by = c(reference_method, threshold_label)
  )

threshold_sensitivity_summary
```

```{r threshold-sensitivity-plot, fig.width=12, fig.height=7}
ggplot(
  threshold_sensitivity,
  aes(x = reference_method, y = net_exceedance, color = reference_method)
) +
  geom_hline(yintercept = 0, color = "grey30", linewidth = 0.4) +
  geom_boxplot(outlier.shape = NA, width = 0.6, color = "grey50") +
  facet_wrap(~ threshold_label, nrow = 1) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_color_manual(values = c(
    "Official Baseline" = "#555555",
    "Baseline-Con1-Con2 mean" = "#b2182b",
    "Baseline-Con1-Con2 median" = "#ef8a62",
    "Con1-Con2 mean" = "#2166ac"
  )) +
  labs(
    title = "Sensitivity of regulatory-threshold net exceedance to reference choice",
    subtitle = "Distributions include every construction survey from Con3 onward across all monitored transects",
    x = NULL,
    y = "Accumulation minus equivalent-loss prevalence",
    color = "Reference"
  ) +
  guides(color = guide_legend(nrow = 2)) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom",
    strip.text = element_text(face = "bold")
  )
```

# Decision matrix

```{r decision-matrix}
decision_matrix <- tribble(
  ~reference_method, ~official_status, ~outlier_resistance, ~early_deposition_risk, ~interpretive_note,
  "Official Baseline", "Regulatory baseline", "None", "None", "Direct BMP comparison, but may represent a transient high state",
  "Baseline-Con1-Con2 mean", "Sensitivity reference", "Low", "Moderate", "Uses all three surveys but a high baseline still pulls the reference upward",
  "Baseline-Con1-Con2 median", "Sensitivity reference", "High for one anomalous survey", "Moderate", "Retains official data and selects the middle early observation at each meter",
  "Con1-Con2 mean", "Sensitivity reference", "Not affected by official baseline", "Highest", "Best description of the early construction state, but most vulnerable to absorbing early effects"
)

decision_matrix
```

# Provisional interpretation

```{r provisional-metrics, include=FALSE}
n_baseline_higher_con1 <- paired_tests %>%
  filter(
    comparison == "Baseline - Con1",
    median_excess_cm > 0,
    wilcoxon_fdr < 0.05
  ) %>%
  nrow()

n_baseline_higher_con2 <- paired_tests %>%
  filter(
    comparison == "Baseline - Con2",
    median_excess_cm > 0,
    wilcoxon_fdr < 0.05
  ) %>%
  nrow()

overall_baseline_excess <- early_meter_wide %>%
  summarize(mean(baseline_minus_early_mean_cm)) %>%
  pull()

best_holdout_method <- holdout_overall %>%
  slice_min(mean_absolute_transect_bias_cm, n = 1, with_ties = FALSE) %>%
  pull(reference_method) %>%
  as.character()

best_sequential_median_window <- sequential_summary %>%
  filter(estimator == "Median") %>%
  slice_min(mean_absolute_transect_bias_cm, n = 1, with_ties = FALSE) %>%
  pull(window_end)

baseline_transition_p <- transition_test$paired_absolute_shift_p
```

Across the 13 monitored transects, the official baseline is significantly
higher than Con1 at `r n_baseline_higher_con1` transects and significantly
higher than Con2 at `r n_baseline_higher_con2` transects after false-discovery
rate adjustment. Averaged across all paired meter positions, the official
baseline exceeds the Con1-Con2 mean by `r round(overall_baseline_excess, 2)` cm.
At the transect level, the Baseline-to-Con1 shift is larger than each
transect's typical early absolute shift (`r paste0("paired p = ",
format.pval(baseline_transition_p, digits = 2))`).

The candidate with the smallest mean absolute transect-level bias during the
Con3-Con6 holdout is **`r best_holdout_method`**. That result should be weighed
against regulatory continuity and the possibility that early construction
surveys already include project effects.

When reference windows are extended sequentially, the median estimator has its
smallest next-survey transect bias for a window ending at **Con`r
best_sequential_median_window`**. This makes surveys through that point worth
considering in the reference-window discussion, but only if the transition and
modeled-impact-class diagnostics show no selective or sustained project-like
departure before then. The sequential result is evidence about temporal
stability, not a license to absorb all surveys through Con`r
best_sequential_median_window` into baseline.

For a single robust sensitivity reference applied consistently across
transects, the **meter-specific Baseline-Con1-Con2 median** is a strong default
candidate because it:

- uses the official baseline rather than discarding it;
- resists one unusually high or low early survey;
- remains meter-matched and therefore preserves localized spatial structure;
- is conservative relative to the Con1-Con2 mean whenever Baseline is the
  highest of the three observations.

The official baseline should still be retained in regulatory reporting, with
the robust alternative presented as a clearly labeled sensitivity analysis.
Before changing the localized-analysis workflow, the statistical evidence in
this report should be considered alongside field notes, sampling conditions,
seasonality, construction timing, and any documented method changes.
