---
title: "Sedimentation Monitoring"
author: "Ross Cunning"
date: "2026-04-29"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
```

## Import

```{r paths}
sediment_dir <- "data/sedimentation_monitoring"
processed_dir <- "data/processed"
audit_dir <- "analysis/sedimentation_raw_data_audit"

dir.create(processed_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(audit_dir, showWarnings = FALSE, recursive = TRUE)

gps_file <- file.path(sediment_dir, "PE_SB_Station_Transect_GPS.xlsx")

sediment_files <- list.files(
  sediment_dir,
  pattern = "xlsx$",
  recursive = TRUE,
  full.names = TRUE
) %>%
  discard(~ basename(.x) == basename(gps_file))
```

```{r helpers}
excel_date <- function(x) {
  if (inherits(x, "Date")) return(x)
  if (inherits(x, "POSIXt")) return(as.Date(x))

  suppressWarnings(as.Date(as.numeric(x), origin = "1899-12-30"))
}

survey_number <- function(x) {
  if_else(
    x == "Baseline",
    0L,
    suppressWarnings(as.integer(str_extract(x, "\\d+")))
  )
}

normalize_transect_id <- function(x) {
  x %>%
    str_squish() %>%
    str_replace_all("\\s+", "_")
}

model_sedimentation_class <- function(transect) {
  case_when(
    transect %in% c("ST22", "ST31") ~ "1.0 cm predicted (dashed red)",
    transect %in% c("ST28", "ST5") ~ "1.5 mm predicted (solid red)",
    TRUE ~ "0 cm predicted (outside)"
  ) %>%
    factor(
      levels = c(
        "1.0 cm predicted (dashed red)",
        "1.5 mm predicted (solid red)",
        "0 cm predicted (outside)"
      ),
      ordered = TRUE
    )
}

model_sedimentation_predictions <- tibble(
  model_sedimentation_class = factor(
    c(
      "1.0 cm predicted (dashed red)",
      "1.5 mm predicted (solid red)",
      "0 cm predicted (outside)"
    ),
    levels = c(
      "1.0 cm predicted (dashed red)",
      "1.5 mm predicted (solid red)",
      "0 cm predicted (outside)"
    ),
    ordered = TRUE
  ),
  predicted_total_sedimentation_cm = c(1, 0.15, 0)
)

threshold_class <- function(x) {
  case_when(
    is.na(x) ~ NA_character_,
    x > 10 ~ ">10 cm",
    x > 5 ~ "5-10 cm",
    x > 1 ~ "1-5 cm",
    x > 0.5 ~ ">5 mm to 1 cm",
    x > 0.15 ~ ">1.5 to 5 mm",
    x >= 0 ~ "0 to 1.5 mm",
    TRUE ~ "below baseline"
  ) %>%
    factor(
      levels = c(
        "below baseline",
        "0 to 1.5 mm",
        ">1.5 to 5 mm",
        ">5 mm to 1 cm",
        "1-5 cm",
        "5-10 cm",
        ">10 cm"
      )
    )
}

read_sediment_raw <- function(path) {
  read_excel(path, sheet = "SedimentDepth_RawData", guess_max = 20000) %>%
    clean_names() %>%
    mutate(
      source_file = path,
      source_file_name = basename(path),
      source_modified = file.info(path)$mtime,
      monitoring_phase = case_when(
        str_detect(path, "1_Pre-Construction") ~ "pre_construction",
        str_detect(path, "2_During-Construction") ~ "during_construction",
        TRUE ~ "unknown"
      ),
      monitoring_type = case_when(
        str_detect(path, "Weekly Monitoring") ~ "weekly",
        str_detect(path, "Exceedance Triggered Monitoring") ~ "exceedance_triggered",
        TRUE ~ "compiled"
      ),
      date = excel_date(date),
      survey = str_squish(survey),
      survey_n = survey_number(survey),
      transect_type = str_squish(transect_type),
      transect = normalize_transect_id(transect),
      model_sedimentation_class = model_sedimentation_class(transect),
      meter = as.integer(meter),
      expected_permanent_length_m = case_when(
        transect_type == "Permanent" & transect %in% c("ST1", "ST38") ~ 150L,
        transect_type == "Permanent" ~ 50L,
        TRUE ~ NA_integer_
      ),
      across(starts_with("sed_depth_cm_"), as.numeric)
    )
}
```

## GPS Lookup

```{r gps}
hardbottom_gps <- read_excel(gps_file, sheet = "Hardbottom") %>%
  clean_names() %>%
  transmute(
    habitat = "hardbottom",
    transect = normalize_transect_id(str_remove(station_transect_id, "\\s+\\d+m$")),
    meter = as.integer(str_extract(station_transect_id, "\\d+(?=m$)")),
    gps_note = notes,
    lat_dd,
    lon_dd
  ) %>%
  filter(!is.na(transect), !is.na(meter))

seagrass_gps <- read_excel(gps_file, sheet = "Seagrass") %>%
  clean_names() %>%
  transmute(
    habitat = "seagrass",
    transect = normalize_transect_id(name),
    meter = NA_integer_,
    gps_note = location,
    lat_dd,
    lon_dd
  )

transect_meter_gps <- hardbottom_gps %>%
  filter(n() > 1, .by = transect) %>%
  arrange(transect, meter) %>%
  group_by(transect) %>%
  group_modify(~ {
    meter_seq <- seq(min(.x$meter), max(.x$meter))

    tibble(
      meter = meter_seq,
      lat_dd = approx(.x$meter, .x$lat_dd, xout = meter_seq, rule = 2)$y,
      lon_dd = approx(.x$meter, .x$lon_dd, xout = meter_seq, rule = 2)$y
    )
  }) %>%
  ungroup()
```

## Raw Sediment Data

The monitoring workbooks are cumulative, so later files repeat earlier surveys.
After reading all workbooks, this keeps the most recent source for each
survey/transect/meter record while retaining source metadata for traceability.

```{r sediment-raw}
sediment_raw_sources <- map_dfr(sediment_files, read_sediment_raw) %>%
  mutate(
    source_max_survey_n = max(survey_n, na.rm = TRUE),
    n_replicates_present = rowSums(!is.na(pick(starts_with("sed_depth_cm_")))),
    .by = source_file
  )

sediment_raw <- sediment_raw_sources %>%
  arrange(desc(source_max_survey_n), desc(source_modified), source_file_name) %>%
  distinct(
    date,
    survey,
    survey_n,
    transect_type,
    transect,
    meter,
    sed_depth_cm_1,
    sed_depth_cm_2,
    sed_depth_cm_3,
    .keep_all = TRUE
  ) %>%
  arrange(survey_n, transect, meter)

sediment_replicates <- sediment_raw %>%
  pivot_longer(
    starts_with("sed_depth_cm_"),
    names_to = "replicate",
    values_to = "sediment_depth_cm",
    values_drop_na = TRUE
  ) %>%
  mutate(
    replicate = as.integer(str_extract(replicate, "\\d+$")),
    sediment_depth_class = threshold_class(sediment_depth_cm)
  ) %>%
  left_join(transect_meter_gps, by = c("transect", "meter"))
```

## Raw Data Audit

These summaries scope only the `SedimentDepth_RawData` sheets. The processed
mean/difference sheets in the workbooks are not used.

```{r raw-data-audit}
raw_file_audit <- sediment_raw_sources %>%
  summarize(
    raw_rows = n(),
    unique_surveys = n_distinct(survey),
    surveys = paste(unique(survey[order(survey_n)]), collapse = ", "),
    n_transects = n_distinct(transect),
    n_transect_meter_positions = n_distinct(paste(transect, meter)),
    min_date = min(date, na.rm = TRUE),
    max_date = max(date, na.rm = TRUE),
    replicate_count_values = paste(sort(unique(n_replicates_present)), collapse = ", "),
    .by = c(monitoring_phase, monitoring_type, source_file_name)
  ) %>%
  arrange(monitoring_phase, monitoring_type, source_file_name)

raw_overall_audit <- sediment_raw %>%
  summarize(
    unique_raw_rows = n(),
    survey_events = n_distinct(survey),
    survey_date_combinations = n_distinct(paste(survey, date)),
    unique_dates = n_distinct(date),
    transects = n_distinct(transect),
    permanent_transects = n_distinct(transect[transect_type == "Permanent"]),
    adjacent_transects = n_distinct(transect[transect_type == "Adjacent"]),
    transect_meter_positions = n_distinct(paste(transect, meter)),
    replicate_columns = paste(
      names(select(sediment_raw, starts_with("sed_depth_cm_"))),
      collapse = ", "
    ),
    replicate_count_values = paste(sort(unique(n_replicates_present)), collapse = ", "),
    total_replicate_measurements = sum(n_replicates_present),
    min_date = min(date, na.rm = TRUE),
    max_date = max(date, na.rm = TRUE)
  )

raw_transect_audit <- sediment_raw %>%
  summarize(
    transect_type = first(transect_type),
    model_sedimentation_class = first(model_sedimentation_class),
    min_meter = min(meter),
    max_meter = max(meter),
    expected_permanent_length_m = if_else(
      all(is.na(expected_permanent_length_m)),
      NA_integer_,
      first(na.omit(expected_permanent_length_m))
    ),
    observed_length_m = max_meter - min_meter,
    length_matches_expected = if_else(
      transect_type == "Permanent",
      observed_length_m == expected_permanent_length_m,
      NA
    ),
    n_meters = n_distinct(meter),
    n_survey_events = n_distinct(survey),
    surveys = paste(unique(survey[order(survey_n)]), collapse = ", "),
    n_raw_meter_rows = n(),
    n_rep_measurements = sum(n_replicates_present),
    .by = transect
  ) %>%
  arrange(transect_type, parse_number(transect), transect)

raw_survey_audit <- sediment_raw %>%
  summarize(
    dates = paste(sort(unique(date)), collapse = ", "),
    n_dates = n_distinct(date),
    n_transects = n_distinct(transect),
    transects = paste(sort(unique(transect)), collapse = ", "),
    n_meter_rows = n(),
    n_permanent_transects = n_distinct(transect[transect_type == "Permanent"]),
    n_adjacent_transects = n_distinct(transect[transect_type == "Adjacent"]),
    total_rep_measurements = sum(n_replicates_present),
    .by = c(survey_n, survey)
  ) %>%
  arrange(survey_n)

raw_replicate_audit <- sediment_raw %>%
  summarize(
    n_meter_rows = n(),
    n_rep_1_present = sum(!is.na(sed_depth_cm_1)),
    n_rep_2_present = sum(!is.na(sed_depth_cm_2)),
    n_rep_3_present = sum(!is.na(sed_depth_cm_3)),
    min_reps_per_meter = min(n_replicates_present),
    max_reps_per_meter = max(n_replicates_present),
    rows_with_0_reps = sum(n_replicates_present == 0),
    rows_with_1_rep = sum(n_replicates_present == 1),
    rows_with_2_reps = sum(n_replicates_present == 2),
    rows_with_3_reps = sum(n_replicates_present == 3),
    .by = c(survey_n, survey)
  ) %>%
  arrange(survey_n)

raw_workbook_duplication_audit <- sediment_raw_sources %>%
  summarize(
    source_file_count = n_distinct(source_file),
    .by = c(
      date,
      survey,
      survey_n,
      transect_type,
      transect,
      meter,
      sed_depth_cm_1,
      sed_depth_cm_2,
      sed_depth_cm_3
    )
  ) %>%
  count(source_file_count, name = "unique_raw_rows") %>%
  arrange(source_file_count)

raw_value_conflict_audit <- sediment_raw_sources %>%
  summarize(
    n_value_versions = n_distinct(
      paste(sed_depth_cm_1, sed_depth_cm_2, sed_depth_cm_3)
    ),
    values = paste(
      unique(paste(sed_depth_cm_1, sed_depth_cm_2, sed_depth_cm_3, sep = "/")),
      collapse = " | "
    ),
    source_files = paste(sort(unique(source_file_name)), collapse = " | "),
    .by = c(date, survey, survey_n, transect_type, transect, meter)
  ) %>%
  filter(n_value_versions > 1) %>%
  arrange(survey_n, transect, meter)

raw_gps_audit <- sediment_raw %>%
  distinct(transect_type, transect) %>%
  left_join(
    hardbottom_gps %>%
      summarize(
        gps_points = n(),
        gps_meters = paste(sort(meter), collapse = ", "),
        .by = transect
      ),
    by = "transect"
  ) %>%
  arrange(transect_type, parse_number(transect), transect)

raw_file_audit
raw_overall_audit
raw_transect_audit
raw_survey_audit
raw_replicate_audit
raw_workbook_duplication_audit
raw_value_conflict_audit
raw_gps_audit
```

```{r export-raw-audit}
write_csv(raw_file_audit, file.path(audit_dir, "raw_file_audit.csv"))
write_csv(raw_overall_audit, file.path(audit_dir, "raw_overall_audit.csv"))
write_csv(raw_transect_audit, file.path(audit_dir, "raw_transect_audit.csv"))
write_csv(raw_survey_audit, file.path(audit_dir, "raw_survey_audit.csv"))
write_csv(raw_replicate_audit, file.path(audit_dir, "raw_replicate_audit.csv"))
write_csv(
  raw_workbook_duplication_audit,
  file.path(audit_dir, "raw_workbook_duplication_audit.csv")
)
write_csv(
  raw_value_conflict_audit,
  file.path(audit_dir, "raw_value_conflict_audit.csv")
)
write_csv(raw_gps_audit, file.path(audit_dir, "raw_gps_audit.csv"))
```

## Baseline Differences

Sedimentation impact is defined in the Biological Monitoring Plan as an increase
in mean sediment depth of 1 cm or more above mean baseline sediment depth. This
chunk creates both raw replicate-level deltas and meter-level mean deltas so
individual high measurements remain visible before any transect averaging.

```{r baseline-differences}
baseline_by_meter <- sediment_raw %>%
  filter(survey == "Baseline") %>%
  transmute(
    transect_type,
    transect,
    meter,
    baseline_mean_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    ),
    baseline_max_cm = do.call(
      pmax,
      c(pick(starts_with("sed_depth_cm_")), list(na.rm = TRUE))
    )
  )

sediment_replicates_processed <- sediment_replicates %>%
  left_join(
    baseline_by_meter,
    by = c("transect_type", "transect", "meter")
  ) %>%
  mutate(
    sediment_delta_from_baseline_cm = sediment_depth_cm - baseline_mean_cm,
    sediment_delta_class = threshold_class(sediment_delta_from_baseline_cm),
    raw_gt_1cm_above_baseline = sediment_delta_from_baseline_cm >= 1
  )

sediment_meter_processed <- sediment_raw %>%
  mutate(
    mean_sediment_depth_cm = rowMeans(
      pick(starts_with("sed_depth_cm_")),
      na.rm = TRUE
    ),
    max_sediment_depth_cm = do.call(
      pmax,
      c(pick(starts_with("sed_depth_cm_")), list(na.rm = TRUE))
    ),
    n_replicates = rowSums(!is.na(pick(starts_with("sed_depth_cm_"))))
  ) %>%
  select(
    date,
    survey,
    survey_n,
    transect_type,
    transect,
    model_sedimentation_class,
    meter,
    expected_permanent_length_m,
    mean_sediment_depth_cm,
    max_sediment_depth_cm,
    n_replicates,
    monitoring_phase,
    monitoring_type,
    source_file,
    source_file_name
  ) %>%
  left_join(baseline_by_meter, by = c("transect_type", "transect", "meter")) %>%
  left_join(transect_meter_gps, by = c("transect", "meter")) %>%
  mutate(
    mean_delta_from_baseline_cm = mean_sediment_depth_cm - baseline_mean_cm,
    max_delta_from_baseline_cm = max_sediment_depth_cm - baseline_mean_cm,
    mean_delta_class = threshold_class(mean_delta_from_baseline_cm),
    max_delta_class = threshold_class(max_delta_from_baseline_cm),
    sedimentation_impact = mean_delta_from_baseline_cm >= 1,
    main_50m_segment = transect %in% c("ST1", "ST38") & between(meter, 0, 50)
  )
```

## Transect Summaries

```{r transect-summaries}
sediment_transect_processed <- sediment_meter_processed %>%
  summarize(
    n_meters = n(),
    transect_mean_depth_cm = mean(mean_sediment_depth_cm, na.rm = TRUE),
    transect_mean_baseline_cm = mean(baseline_mean_cm, na.rm = TRUE),
    transect_mean_delta_cm = transect_mean_depth_cm - transect_mean_baseline_cm,
    max_meter_mean_delta_cm = max(mean_delta_from_baseline_cm, na.rm = TRUE),
    max_raw_delta_cm = max(max_delta_from_baseline_cm, na.rm = TRUE),
    n_meters_mean_gt_1cm = sum(mean_delta_from_baseline_cm >= 1, na.rm = TRUE),
    pct_meters_mean_gt_1cm = n_meters_mean_gt_1cm / n_meters,
    sedimentation_impact = transect_mean_delta_cm >= 1,
    main_50m_sedimentation_impact = any(
      main_50m_segment & mean_delta_from_baseline_cm >= 1,
      na.rm = TRUE
    ),
    .by = c(
      date,
      survey,
      survey_n,
      transect_type,
      transect,
      model_sedimentation_class
    )
  ) %>%
  arrange(survey_n, model_sedimentation_class, transect)

sediment_transect_processed %>%
  filter(survey != "Baseline") %>%
  arrange(desc(max_raw_delta_cm))
```

## Export Processed Tables

```{r export}
write_csv(
  sediment_replicates_processed,
  file.path(processed_dir, "sedimentation_replicates_processed.csv")
)

write_csv(
  sediment_meter_processed,
  file.path(processed_dir, "sedimentation_meter_processed.csv")
)

write_csv(
  sediment_transect_processed,
  file.path(processed_dir, "sedimentation_transect_processed.csv")
)
```

## Quick Visual Checks

The raw sediment sheets contain 19 transects total. Six adjacent transects are
Baseline-only, so construction-only plots show the 13 permanent transects.

```{r raw-coverage-plot, fig.width=11, fig.height=6}
sediment_raw %>%
  distinct(transect_type, transect, survey_n, survey) %>%
  mutate(
    transect = fct_reorder(transect, survey_n, .fun = max),
    survey = fct_reorder(survey, survey_n)
  ) %>%
  ggplot(aes(x = survey, y = transect, fill = transect_type)) +
  geom_tile(color = "white", linewidth = 0.2) +
  scale_fill_manual(
    values = c("Adjacent" = "#7f8c8d", "Permanent" = "#2a9d8f")
  ) +
  labs(
    x = "Survey event",
    y = "Transect",
    fill = "Transect type"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r raw-baseline-depth-plot, fig.width=11, fig.height=7}
sediment_replicates %>%
  filter(survey == "Baseline") %>%
  ggplot(aes(x = meter, y = sediment_depth_cm)) +
  geom_point(aes(color = sediment_depth_class), alpha = 0.55, size = 0.7) +
  facet_wrap(~ transect, scales = "free_x") +
  scale_color_manual(
    values = c(
      "below baseline" = "grey90",
      "0 to 1.5 mm" = "#8ecae6",
      ">1.5 to 5 mm" = "#ffb703",
      ">5 mm to 1 cm" = "#fb8500",
      "1-5 cm" = "#d00000",
      "5-10 cm" = "#9d0208",
      ">10 cm" = "#370617"
    ),
    drop = FALSE
  ) +
  labs(
    x = "Transect meter",
    y = "Raw sediment depth (cm)",
    color = "Raw depth class"
  ) +
  theme_classic()
```

```{r construction-delta-plot, fig.width=11, fig.height=7}
sediment_meter_processed %>%
  filter(survey != "Baseline", transect_type == "Permanent") %>%
  {
    meter_grid <- distinct(
      .,
      date,
      survey,
      survey_n,
      transect_type,
      transect,
      model_sedimentation_class,
      expected_permanent_length_m
    ) %>%
      mutate(meter = map(expected_permanent_length_m, ~ seq(0L, .x))) %>%
      unnest(meter)

    right_join(
      .,
      meter_grid,
      by = c(
        "date",
        "survey",
        "survey_n",
        "transect_type",
        "transect",
        "model_sedimentation_class",
        "expected_permanent_length_m",
        "meter"
      )
    )
  } %>%
  mutate(
    transect = fct_reorder(
      transect,
      as.integer(model_sedimentation_class) * 1000 + parse_number(transect),
      .fun = min
    )
  ) %>%
  ggplot(aes(x = meter, y = survey_n, fill = mean_delta_class)) +
  geom_tile() +
  facet_wrap(~ model_sedimentation_class + transect, scales = "free_x") +
  scale_fill_manual(
    values = c(
      "below baseline" = "grey90",
      "0 to 1.5 mm" = "#8ecae6",
      ">1.5 to 5 mm" = "#ffb703",
      ">5 mm to 1 cm" = "#fb8500",
      "1-5 cm" = "#d00000",
      "5-10 cm" = "#9d0208",
      ">10 cm" = "#370617"
    ),
    na.value = "grey95",
    drop = FALSE
  ) +
  labs(
    x = "Transect meter",
    y = "Construction survey",
    fill = "Mean delta from baseline"
  ) +
  theme_classic()
```

## Reef-scale Construction Delta Summaries

This summarizes construction-period sediment deltas by survey and model
sedimentation class. Percentages are calculated among measured meter positions;
unsampled expected meter positions are retained in the tables for coverage
tracking but omitted from the threshold-composition figure.

```{r reef-delta-summary, fig.width=10, fig.height=6}
reef_summary_dir <- file.path("analysis", "sedimentation_reef_delta_summary")
dir.create(reef_summary_dir, showWarnings = FALSE, recursive = TRUE)

threshold_levels <- levels(threshold_class(0))
threshold_colors <- c(
  "below baseline" = "grey90",
  "0 to 1.5 mm" = "#8ecae6",
  ">1.5 to 5 mm" = "#ffb703",
  ">5 mm to 1 cm" = "#fb8500",
  "1-5 cm" = "#d00000",
  "5-10 cm" = "#9d0208",
  ">10 cm" = "#370617",
  "unmeasured" = "grey95"
)

construction_reef_delta_grid <- sediment_meter_processed %>%
  filter(survey != "Baseline", transect_type == "Permanent") %>%
  {
    meter_grid <- distinct(
      .,
      date,
      survey,
      survey_n,
      transect_type,
      transect,
      model_sedimentation_class,
      expected_permanent_length_m
    ) %>%
      mutate(meter = map(expected_permanent_length_m, ~ seq(0L, .x))) %>%
      unnest(meter)

    right_join(
      .,
      meter_grid,
      by = c(
        "date",
        "survey",
        "survey_n",
        "transect_type",
        "transect",
        "model_sedimentation_class",
        "expected_permanent_length_m",
        "meter"
      )
    )
  } %>%
  mutate(
    mean_delta_class = factor(
      if_else(
        is.na(mean_delta_class),
        "unmeasured",
        as.character(mean_delta_class)
      ),
      levels = c(threshold_levels, "unmeasured")
    )
  )

reef_delta_denominators <- construction_reef_delta_grid %>%
  summarize(
    survey_start_date = min(date, na.rm = TRUE),
    survey_end_date = max(date, na.rm = TRUE),
    survey_dates = paste(sort(unique(date)), collapse = ", "),
    expected_meter_positions = n(),
    measured_meter_positions = sum(!is.na(mean_delta_from_baseline_cm)),
    pct_meter_positions_measured = measured_meter_positions / expected_meter_positions,
    .by = c(survey, survey_n, model_sedimentation_class)
  )

reef_delta_summary_by_week_class <- construction_reef_delta_grid %>%
  filter(!is.na(mean_delta_from_baseline_cm)) %>%
  summarize(
    mean_delta_cm = mean(mean_delta_from_baseline_cm),
    p01_delta_cm = quantile(mean_delta_from_baseline_cm, 0.01),
    p10_delta_cm = quantile(mean_delta_from_baseline_cm, 0.10),
    p25_delta_cm = quantile(mean_delta_from_baseline_cm, 0.25),
    median_delta_cm = median(mean_delta_from_baseline_cm),
    p75_delta_cm = quantile(mean_delta_from_baseline_cm, 0.75),
    p90_delta_cm = quantile(mean_delta_from_baseline_cm, 0.90),
    p99_delta_cm = quantile(mean_delta_from_baseline_cm, 0.99),
    min_delta_cm = min(mean_delta_from_baseline_cm),
    max_delta_cm = max(mean_delta_from_baseline_cm),
    range_delta_cm = max_delta_cm - min_delta_cm,
    .by = c(survey, survey_n, model_sedimentation_class)
  ) %>%
  left_join(
    reef_delta_denominators,
    by = c("survey", "survey_n", "model_sedimentation_class")
  ) %>%
  arrange(survey_n, model_sedimentation_class)

reef_delta_class_pct_by_week_class <- construction_reef_delta_grid %>%
  count(
    survey,
    survey_n,
    model_sedimentation_class,
    mean_delta_class,
    name = "n_meter_positions"
  ) %>%
  left_join(
    reef_delta_denominators,
    by = c("survey", "survey_n", "model_sedimentation_class")
  ) %>%
  mutate(
    pct_expected_reef = n_meter_positions / expected_meter_positions,
    pct_measured_reef = if_else(
      mean_delta_class == "unmeasured" | measured_meter_positions == 0,
      NA_real_,
      n_meter_positions / measured_meter_positions
    )
  ) %>%
  arrange(survey_n, model_sedimentation_class, mean_delta_class)

reef_delta_class_pct_wide <- reef_delta_class_pct_by_week_class %>%
  filter(mean_delta_class != "unmeasured") %>%
  transmute(
    survey,
    survey_n,
    survey_start_date,
    survey_end_date,
    survey_dates,
    model_sedimentation_class,
    mean_delta_class,
    pct_measured_reef = 100 * pct_measured_reef
  ) %>%
  pivot_wider(
    names_from = mean_delta_class,
    values_from = pct_measured_reef,
    values_fill = 0,
    names_prefix = "pct_measured_"
  ) %>%
  arrange(survey_n, model_sedimentation_class)

write_csv(
  reef_delta_summary_by_week_class,
  file.path(reef_summary_dir, "reef_delta_summary_by_week_class.csv")
)

write_csv(
  reef_delta_class_pct_by_week_class,
  file.path(reef_summary_dir, "reef_delta_class_pct_by_week_class.csv")
)

write_csv(
  reef_delta_class_pct_wide,
  file.path(reef_summary_dir, "reef_delta_class_pct_wide.csv")
)

reef_delta_summary_by_week_class
reef_delta_class_pct_wide
```

```{r reef-delta-summary-figures, fig.width=11, fig.height=7}
reef_delta_summary_plot <- reef_delta_summary_by_week_class %>%
  select(
    survey_start_date,
    survey_n,
    model_sedimentation_class,
    p01_delta_cm,
    p10_delta_cm,
    p25_delta_cm,
    median_delta_cm,
    p75_delta_cm,
    p90_delta_cm,
    p99_delta_cm
  ) %>%
  pivot_longer(
    ends_with("_delta_cm"),
    names_to = "delta_percentile",
    values_to = "delta_cm"
  ) %>%
  mutate(
    delta_percentile = recode(
      delta_percentile,
      p01_delta_cm = "1st percentile",
      p10_delta_cm = "10th percentile",
      p25_delta_cm = "25th percentile (25% at or below)",
      median_delta_cm = "Median",
      p75_delta_cm = "75th percentile (25% at or above)",
      p90_delta_cm = "90th percentile",
      p99_delta_cm = "99th percentile"
    ),
    delta_percentile = factor(
      delta_percentile,
      levels = c(
        "1st percentile",
        "10th percentile",
        "25th percentile (25% at or below)",
        "Median",
        "75th percentile (25% at or above)",
        "90th percentile",
        "99th percentile"
      )
    )
  ) %>%
  arrange(model_sedimentation_class, delta_percentile, survey_start_date) %>%
  mutate(
    monitoring_gap_days = as.numeric(survey_start_date - lag(survey_start_date)),
    plot_segment = cumsum(coalesce(monitoring_gap_days > 14, FALSE)),
    .by = c(model_sedimentation_class, delta_percentile)
  ) %>%
  ggplot(aes(x = survey_n, y = delta_cm, color = delta_percentile)) +
  geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
  geom_hline(
    data = model_sedimentation_predictions,
    aes(yintercept = predicted_total_sedimentation_cm),
    linetype = "dashed",
    color = "black",
    linewidth = 0.4
  ) +
  geom_line(aes(group = interaction(delta_percentile, plot_segment)), linewidth = 0.7) +
  facet_wrap(~ model_sedimentation_class) +
  scale_color_manual(
    values = c(
      "1st percentile" = "#2166ac",
      "10th percentile" = "#67a9cf",
      "25th percentile (25% at or below)" = "#d1e5f0",
      "Median" = "#2a9d8f",
      "75th percentile (25% at or above)" = "#fb8500",
      "90th percentile" = "#d00000",
      "99th percentile" = "#370617"
    )
  ) +
  labs(
    x = "Construction survey",
    y = "Sediment delta from baseline (cm)",
    color = "Measured reef percentile",
    title = "Construction sediment delta percentiles by model sedimentation class"
  ) +
  theme_classic()

reef_delta_class_pct_plot <- reef_delta_class_pct_by_week_class %>%
  filter(mean_delta_class != "unmeasured") %>%
  ggplot(aes(x = survey_n, y = pct_measured_reef, fill = mean_delta_class)) +
  geom_col(width = 0.85) +
  facet_wrap(~ model_sedimentation_class) +
  scale_fill_manual(values = threshold_colors, drop = FALSE) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Construction survey",
    y = "Percent of measured reef meter positions",
    fill = "Mean delta class",
    title = "Threshold class composition by model sedimentation class"
  ) +
  theme_classic()

ggsave(
  file.path(reef_summary_dir, "reef_delta_summary_by_week_class.png"),
  reef_delta_summary_plot,
  width = 11,
  height = 7,
  dpi = 300
)

ggsave(
  file.path(reef_summary_dir, "reef_delta_class_pct_by_week_class.png"),
  reef_delta_class_pct_plot,
  width = 11,
  height = 7,
  dpi = 300
)

reef_delta_summary_plot
reef_delta_class_pct_plot
```

## Net Positive Exceedance Tests

These tests compare the measured area with positive sedimentation above a
threshold to the measured area with negative sedimentation below the equivalent
threshold. The `outside` class is outside the model-predicted sedimentation
footprint, but it is not treated as an unimpacted reference because it may still
receive dredging-related sediment. The bootstrap resamples transects within
each survey and model sedimentation class, so confidence intervals reflect
transect-level clustering rather than treating every meter as fully independent.

```{r net-positive-exceedance-tests, fig.width=11, fig.height=7}
set.seed(20260429)

net_exceedance_bootstrap_n <- 1000

net_exceedance_thresholds <- tibble(
  threshold_cm = c(0.15, 0.5, 1),
  threshold_label = factor(
    c("0.15 cm (1.5 mm)", "0.5 cm", "1.0 cm"),
    levels = c("0.15 cm (1.5 mm)", "0.5 cm", "1.0 cm")
  )
)

bootstrap_net_exceedance <- function(data, threshold_cm, n_boot) {
  transects <- sort(unique(data$transect))

  map_dfr(seq_len(n_boot), function(boot_i) {
    boot_delta <- sample(transects, length(transects), replace = TRUE) %>%
      map(~ data$mean_delta_from_baseline_cm[data$transect == .x]) %>%
      list_c()

    tibble(
      boot_i = boot_i,
      positive_area = mean(boot_delta >= threshold_cm),
      negative_area = mean(boot_delta <= -threshold_cm),
      net_positive_area = positive_area - negative_area
    )
  })
}

construction_measured_deltas <- construction_reef_delta_grid %>%
  filter(!is.na(mean_delta_from_baseline_cm))

net_exceedance_observed <- construction_measured_deltas %>%
  cross_join(net_exceedance_thresholds) %>%
  summarize(
    survey_start_date = min(date, na.rm = TRUE),
    survey_end_date = max(date, na.rm = TRUE),
    survey_dates = paste(sort(unique(date)), collapse = ", "),
    measured_meter_positions = n(),
    n_transects = n_distinct(transect),
    positive_meter_positions = sum(mean_delta_from_baseline_cm >= threshold_cm),
    negative_meter_positions = sum(mean_delta_from_baseline_cm <= -threshold_cm),
    positive_area = positive_meter_positions / measured_meter_positions,
    negative_area = negative_meter_positions / measured_meter_positions,
    net_positive_area = positive_area - negative_area,
    .by = c(
      survey,
      survey_n,
      model_sedimentation_class,
      threshold_cm,
      threshold_label
    )
  )

net_exceedance_bootstrap <- construction_measured_deltas %>%
  cross_join(net_exceedance_thresholds) %>%
  group_by(survey, survey_n, model_sedimentation_class, threshold_cm, threshold_label) %>%
  group_modify(
    ~ bootstrap_net_exceedance(
      .x,
      threshold_cm = .y$threshold_cm,
      n_boot = net_exceedance_bootstrap_n
    )
  ) %>%
  ungroup()

net_exceedance_ci <- net_exceedance_bootstrap %>%
  summarize(
    net_positive_area_lcl = quantile(net_positive_area, 0.025),
    net_positive_area_ucl = quantile(net_positive_area, 0.975),
    .by = c(
      survey,
      survey_n,
      model_sedimentation_class,
      threshold_cm,
      threshold_label
    )
  )

net_exceedance_by_week_class <- net_exceedance_observed %>%
  left_join(
    net_exceedance_ci,
    by = c(
      "survey",
      "survey_n",
      "model_sedimentation_class",
      "threshold_cm",
      "threshold_label"
    )
  ) %>%
  mutate(
    ci_note = if_else(
      n_transects < 3,
      "CI not estimated: fewer than 3 transects",
      "Transect bootstrap CI"
    ),
    across(
      c(net_positive_area_lcl, net_positive_area_ucl),
      ~ if_else(n_transects < 3, NA_real_, .x)
    ),
    evidence_positive_asymmetry = net_positive_area_lcl > 0
  ) %>%
  arrange(survey_n, model_sedimentation_class, threshold_cm)

write_csv(
  net_exceedance_by_week_class,
  file.path(reef_summary_dir, "net_positive_exceedance_by_week_class.csv")
)

net_exceedance_by_week_class
```

```{r net-positive-exceedance-figures, fig.width=11, fig.height=7}
net_exceedance_threshold_plot_data <- net_exceedance_by_week_class %>%
  filter(threshold_cm %in% c(0.15, 1.0)) %>%
  arrange(model_sedimentation_class, threshold_label, survey_start_date) %>%
  mutate(
    monitoring_gap_days = as.numeric(survey_start_date - lag(survey_start_date)),
    plot_segment = cumsum(coalesce(monitoring_gap_days > 14, FALSE)),
    .by = c(model_sedimentation_class, threshold_label)
  )

positive_asymmetry_annotations <- net_exceedance_threshold_plot_data %>%
  filter(evidence_positive_asymmetry) %>%
  mutate(annotation_y = net_positive_area_ucl)

net_exceedance_threshold_plot <- net_exceedance_threshold_plot_data %>%
  ggplot(aes(x = survey_n, y = net_positive_area, color = threshold_label)) +
  geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
  geom_ribbon(
    data = ~ filter(.x, !is.na(net_positive_area_lcl), !is.na(net_positive_area_ucl)),
    aes(
      ymin = net_positive_area_lcl,
      ymax = net_positive_area_ucl,
      fill = threshold_label,
      group = interaction(threshold_label, plot_segment)
    ),
    alpha = 0.14,
    color = NA
  ) +
  geom_line(
    aes(group = interaction(threshold_label, plot_segment)),
    linewidth = 0.8
  ) +
  geom_text(
    data = positive_asymmetry_annotations,
    aes(x = survey_n, y = annotation_y, label = "*"),
    color = "black",
    vjust = -0.35,
    size = 4,
    show.legend = FALSE
  ) +
  facet_wrap(~ model_sedimentation_class) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(
    values = c(
      "0.15 cm (1.5 mm)" = "#fb8500",
      "1.0 cm" = "#d00000"
    )
  ) +
  scale_fill_manual(
    values = c(
      "0.15 cm (1.5 mm)" = "#fb8500",
      "1.0 cm" = "#d00000"
    )
  ) +
  labs(
    x = "Construction survey",
    y = "Net measured reef area",
    color = "Delta threshold",
    fill = "Delta threshold",
    title = "Net positive exceedance area by predicted sedimentation class",
    caption = "* Bootstrap CI for net positive area is entirely above zero"
  ) +
  theme_classic()

ggsave(
  file.path(reef_summary_dir, "net_positive_exceedance_by_week_class.png"),
  net_exceedance_threshold_plot,
  width = 11,
  height = 7,
  dpi = 300
)

net_exceedance_threshold_plot

net_exceedance_by_week_class %>% filter(evidence_positive_asymmetry)
```

## Exploratory Attribution Analyses

The net-positive exceedance summaries are useful screening tools, but they are
not a complete attribution analysis. The following sections explore complementary
ways to ask whether construction-period deltas show directional sediment
accumulation. These are deliberately exploratory: each method answers a slightly
different question and has different assumptions.

### Transect-level Net Exceedance Tests

This repeats the threshold asymmetry analysis using transects as the analysis
unit. For each survey, predicted sedimentation class, threshold, and transect,
it calculates the transect-level net positive area. It then tests whether the
transect-level values are centered above zero. This is statistically cleaner
than treating meters as independent, but power is limited when a class has only
one or two transects.

```{r transect-level-net-exceedance-tests}
transect_net_exceedance <- construction_measured_deltas %>%
  cross_join(net_exceedance_thresholds) %>%
  summarize(
    n_meters = n(),
    positive_meter_positions = sum(mean_delta_from_baseline_cm >= threshold_cm),
    negative_meter_positions = sum(mean_delta_from_baseline_cm <= -threshold_cm),
    positive_area = positive_meter_positions / n_meters,
    negative_area = negative_meter_positions / n_meters,
    net_positive_area = positive_area - negative_area,
    .by = c(
      survey,
      survey_n,
      model_sedimentation_class,
      threshold_cm,
      threshold_label,
      transect
    )
  )

transect_net_test <- function(data) {
  nonzero_net <- data$net_positive_area[data$net_positive_area != 0]
  n_positive <- sum(data$net_positive_area > 0)
  n_negative <- sum(data$net_positive_area < 0)

  tibble(
    n_transects = n_distinct(data$transect),
    mean_transect_net_positive_area = mean(data$net_positive_area),
    median_transect_net_positive_area = median(data$net_positive_area),
    n_transects_positive = n_positive,
    n_transects_negative = n_negative,
    sign_test_p_value = if (length(nonzero_net) >= 3) {
      binom.test(
        n_positive,
        n_positive + n_negative,
        p = 0.5,
        alternative = "greater"
      )$p.value
    } else {
      NA_real_
    },
    wilcox_p_value = if (length(nonzero_net) >= 3) {
      wilcox.test(
        data$net_positive_area,
        mu = 0,
        alternative = "greater",
        exact = FALSE
      )$p.value
    } else {
      NA_real_
    },
    test_note = if_else(
      n_transects < 3,
      "Not enough transects for a useful transect-level test",
      "Exploratory one-sided tests of transect-level net positive area > 0"
    )
  )
}

transect_net_exceedance_tests <- transect_net_exceedance %>%
  group_by(survey, survey_n, model_sedimentation_class, threshold_cm, threshold_label) %>%
  group_modify(~ transect_net_test(.x)) %>%
  ungroup() %>%
  arrange(survey_n, model_sedimentation_class, threshold_cm)

write_csv(
  transect_net_exceedance,
  file.path(reef_summary_dir, "transect_net_exceedance.csv")
)

write_csv(
  transect_net_exceedance_tests,
  file.path(reef_summary_dir, "transect_net_exceedance_tests.csv")
)

transect_net_exceedance_tests
```

### Week-specific Delta Percentile Tests

This approach keeps construction survey as a time series rather than fitting a
single linear trend. For each survey and predicted sedimentation class, it
estimates selected meter-level delta percentiles and bootstraps confidence
intervals by resampling transects. The one-sided evidence flag asks whether a
given percentile is above zero for that specific survey. For example, if the
90th percentile is above zero, then the upper 10% of measured reef meters have
positive deltas at that time point. Confidence intervals are not estimated when
fewer than three transects are available in a survey/class group.

```{r week-specific-delta-percentile-tests, fig.width=11, fig.height=7}
week_delta_percentile_bootstrap_n <- 1000

week_delta_percentile_probs <- tibble(
  percentile_prob = c(0.50, 0.75, 0.90, 0.95),
  percentile_label = factor(
    c("Median", "75th percentile", "90th percentile", "95th percentile"),
    levels = c("Median", "75th percentile", "90th percentile", "95th percentile")
  )
)

summarize_delta_percentiles <- function(delta) {
  tibble(
    percentile_prob = week_delta_percentile_probs$percentile_prob,
    percentile_label = week_delta_percentile_probs$percentile_label,
    delta_cm = as.numeric(
      quantile(
        delta,
        probs = week_delta_percentile_probs$percentile_prob,
        names = FALSE
      )
    )
  )
}

bootstrap_delta_percentiles <- function(data, n_boot) {
  transects <- sort(unique(data$transect))

  map_dfr(seq_len(n_boot), function(boot_i) {
    boot_delta <- sample(transects, length(transects), replace = TRUE) %>%
      map(~ data$mean_delta_from_baseline_cm[data$transect == .x]) %>%
      list_c()

    summarize_delta_percentiles(boot_delta) %>%
      mutate(boot_i = boot_i, .before = 1)
  })
}

week_delta_percentile_observed <- construction_measured_deltas %>%
  summarize(
    survey_start_date = min(date, na.rm = TRUE),
    survey_end_date = max(date, na.rm = TRUE),
    survey_dates = paste(sort(unique(date)), collapse = ", "),
    measured_meter_positions = n(),
    n_transects = n_distinct(transect),
    delta_percentiles = list(summarize_delta_percentiles(mean_delta_from_baseline_cm)),
    .by = c(survey, survey_n, model_sedimentation_class)
  ) %>%
  unnest(delta_percentiles)

week_delta_percentile_bootstrap <- construction_measured_deltas %>%
  group_by(survey, survey_n, model_sedimentation_class) %>%
  group_modify(~ bootstrap_delta_percentiles(.x, week_delta_percentile_bootstrap_n)) %>%
  ungroup()

week_delta_percentile_ci <- week_delta_percentile_bootstrap %>%
  summarize(
    delta_cm_lcl = quantile(delta_cm, 0.025),
    delta_cm_ucl = quantile(delta_cm, 0.975),
    .by = c(survey, survey_n, model_sedimentation_class, percentile_prob, percentile_label)
  )

week_delta_percentile_tests <- week_delta_percentile_observed %>%
  left_join(
    week_delta_percentile_ci,
    by = c(
      "survey",
      "survey_n",
      "model_sedimentation_class",
      "percentile_prob",
      "percentile_label"
    )
  ) %>%
  mutate(
    ci_note = if_else(
      n_transects < 3,
      "CI not estimated: fewer than 3 transects",
      "Transect bootstrap CI"
    ),
    across(
      c(delta_cm_lcl, delta_cm_ucl),
      ~ if_else(n_transects < 3, NA_real_, .x)
    ),
    evidence_delta_above_zero = delta_cm_lcl > 0
  ) %>%
  arrange(survey_n, model_sedimentation_class, percentile_prob)

week_delta_percentile_plot_data <- week_delta_percentile_tests %>%
  arrange(model_sedimentation_class, percentile_label, survey_start_date) %>%
  mutate(
    monitoring_gap_days = as.numeric(survey_start_date - lag(survey_start_date)),
    plot_segment = cumsum(coalesce(monitoring_gap_days > 14, FALSE)),
    .by = c(model_sedimentation_class, percentile_label)
  )

week_delta_percentile_annotations <- week_delta_percentile_plot_data %>%
  filter(evidence_delta_above_zero) %>%
  mutate(annotation_y = delta_cm_ucl)

week_delta_percentile_plot <- week_delta_percentile_plot_data %>%
  ggplot(aes(x = survey_n, y = delta_cm, color = percentile_label)) +
  geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
  geom_ribbon(
    data = ~ filter(.x, !is.na(delta_cm_lcl), !is.na(delta_cm_ucl)),
    aes(
      ymin = delta_cm_lcl,
      ymax = delta_cm_ucl,
      fill = percentile_label,
      group = interaction(percentile_label, plot_segment)
    ),
    alpha = 0.12,
    color = NA
  ) +
  geom_line(
    aes(group = interaction(percentile_label, plot_segment)),
    linewidth = 0.75
  ) +
  geom_text(
    data = week_delta_percentile_annotations,
    aes(x = survey_n, y = annotation_y, label = "*"),
    color = "black",
    vjust = -0.35,
    size = 4,
    show.legend = FALSE
  ) +
  facet_wrap(~ model_sedimentation_class) +
  scale_color_manual(
    values = c(
      "Median" = "#2a9d8f",
      "75th percentile" = "#fb8500",
      "90th percentile" = "#d00000",
      "95th percentile" = "#370617"
    )
  ) +
  scale_fill_manual(
    values = c(
      "Median" = "#2a9d8f",
      "75th percentile" = "#fb8500",
      "90th percentile" = "#d00000",
      "95th percentile" = "#370617"
    )
  ) +
  labs(
    x = "Construction survey",
    y = "Sediment delta from baseline (cm)",
    color = "Measured reef percentile",
    fill = "Measured reef percentile",
    title = "Week-specific sediment delta percentiles",
    caption = "* Transect bootstrap CI for percentile is entirely above zero"
  ) +
  theme_classic()

write_csv(
  week_delta_percentile_tests,
  file.path(reef_summary_dir, "week_delta_percentile_tests.csv")
)

ggsave(
  file.path(reef_summary_dir, "week_delta_percentile_tests.png"),
  week_delta_percentile_plot,
  width = 11,
  height = 7,
  dpi = 300
)

week_delta_percentile_plot
week_delta_percentile_tests %>% filter(evidence_delta_above_zero)
```

### Tail Asymmetry Diagnostics

This diagnostic compares matched lower and upper tails in the observed
distribution. Positive values mean the upper sedimentation tail is larger than
the corresponding lower sediment-loss tail. For example, `p90_minus_abs_p10_cm`
is positive when the 90th percentile exceeds the magnitude of the 10th
percentile.

```{r tail-asymmetry-diagnostics, fig.width=11, fig.height=7}
tail_asymmetry_by_week_class <- reef_delta_summary_by_week_class %>%
  transmute(
    survey,
    survey_n,
    survey_start_date,
    survey_end_date,
    survey_dates,
    model_sedimentation_class,
    p75_minus_abs_p25_cm = p75_delta_cm - abs(p25_delta_cm),
    p90_minus_abs_p10_cm = p90_delta_cm - abs(p10_delta_cm),
    p99_minus_abs_p01_cm = p99_delta_cm - abs(p01_delta_cm)
  ) %>%
  pivot_longer(
    ends_with("_cm"),
    names_to = "tail_contrast",
    values_to = "upper_tail_excess_cm"
  ) %>%
  mutate(
    tail_contrast = recode(
      tail_contrast,
      p75_minus_abs_p25_cm = "75th minus |25th|",
      p90_minus_abs_p10_cm = "90th minus |10th|",
      p99_minus_abs_p01_cm = "99th minus |1st|"
    ),
    tail_contrast = factor(
      tail_contrast,
      levels = c("75th minus |25th|", "90th minus |10th|", "99th minus |1st|")
    )
  ) %>%
  arrange(model_sedimentation_class, tail_contrast, survey_start_date) %>%
  mutate(
    monitoring_gap_days = as.numeric(survey_start_date - lag(survey_start_date)),
    plot_segment = cumsum(coalesce(monitoring_gap_days > 14, FALSE)),
    .by = c(model_sedimentation_class, tail_contrast)
  )

tail_asymmetry_plot <- tail_asymmetry_by_week_class %>%
  ggplot(aes(x = survey_n, y = upper_tail_excess_cm, color = tail_contrast)) +
  geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
  geom_line(aes(group = interaction(tail_contrast, plot_segment)), linewidth = 0.7) +
  facet_wrap(~ model_sedimentation_class) +
  labs(
    x = "Construction survey",
    y = "Upper-tail excess over matched lower tail (cm)",
    color = "Tail contrast",
    title = "Sediment-gain tail compared with sediment-loss tail"
  ) +
  theme_classic()

write_csv(
  tail_asymmetry_by_week_class,
  file.path(reef_summary_dir, "tail_asymmetry_by_week_class.csv")
)

ggsave(
  file.path(reef_summary_dir, "tail_asymmetry_by_week_class.png"),
  tail_asymmetry_plot,
  width = 11,
  height = 7,
  dpi = 300
)

tail_asymmetry_plot
```
